### 2017-11-06 Paper.tex
Create folder img to store images. Create folder v1 to contain older version. Paper and bib file renamed.

\documentclass[10pt,a4paper]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}

% for institute information	
\usepackage{authblk}

% definition of keywords
\def\keywords{\vspace{.5em}
	{\bfseries\textit{Keywords}---\,\relax%
}}
\def\endkeywords{\par}

% Packages for table
\usepackage{longtable}

\usepackage{fancyhdr}
\usepackage{booktabs}
\usepackage{cite}
\usepackage{graphicx}
\usepackage{array}
\usepackage{multirow}
\usepackage{amssymb}
\usepackage{fancyhdr}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{float}
\usepackage[tight,footnotesize]{subfigure}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

\usepackage{filecontents}
\usepackage{cite}

\begin{document}
\title{Recent advances of Malaria parasites detection systems based on mathematical morphology techniques}

\author[1]{Andrea Loddo}
\author[1]{Cecilia Di Ruberto}
\author[2]{Michel Kocher}

\affil[1]{Department of Mathematics and Computer Science\\	Via Ospedale 72, 09124, Cagliari, Italy}
\affil[ ]{\textit {\{andrea.loddo, dirubert\}@unica.it}}
\affil[2]{Biomedical Imaging Group, École Polytechnique
	Fédérale de Lausanne (EPFL), CH-1015 Lausanne, Switzerland}
\affil[ ]{\textit {michel.kocher@heig-vd.ch}}
\date{}                     %% if you don't need date to appear
\setcounter{Maxaffil}{0}
\renewcommand\Affilfont{\itshape\small}

\maketitle

\begin{abstract}
	This paper investigates existing mathematical morphology based techniques applied for performing malaria parasites detection and identification in both Giemsa and Leishman stained blood smears images.
	Malaria is an epidemic health disease and a rapid, accurate diagnosis is necessary for proper intervention. Generally, pathologists visually examine blood stained slides for malaria diagnosis; this kind of visual inspection is subjective, error-prone and time consuming. In order to cope with such issues, computer-aided methods have been increasingly evolved for abnormal erythrocyte detection, segmentation and semi/fully automated classification. The aim of this paper is to present a review on the methods that are based on mathematical morphology techniques for malaria parasite detection.

	\keywords{malaria, red blood cells segmentation, mathematical morphology, medical image analysis}
\end{abstract}

\section{Introduction}
Haematology is the branch of medicine concerned with the study, diagnosis, monitoring, treatment, and prevention of diseases related to the blood and blood-forming organs. Haematology studies the blood in health and pathological conditions, firstly to identify the patient's health condition and, secondly, to predict how the bone marrow may have contributed to reach that condition.

Thus, haematology studies the relationship between the bone marrow and the systemic circulation. In fact, there are many diseases, disorders, and deficiencies that can affect the number and type of blood cells produced, their function and their lifespan. Usually, only normal, mature or nearly mature cells are released into the bloodstream but certain circumstances can induce the bone marrow to release immature and/or abnormal cells into the circulation. One of the most frequently ordered test to monitor the proportion of the cell components into the blood stream is the Complete Blood Count (CBC), that offers various haematologic data represented by the numbers and types of cells in the peripheral circulation. The cells percentage is compared with the reference ranges in order to determine if the cells are present in their expected percentage, if one cell type is increased, decreased or if immature cells exist. Reference ranges for blood tests are sets of values used to interpret a set of diagnostic test results from blood samples. Since it is difficult to prove that healthy-considered subjects may not have infections, parasitic infection and nutritional deficiency, it is more feasible to talk about reference ranges rather than normal ranges. A reference range is usually defined as the set of values in which 95\% of the normal population falls within. It is determined by collecting data from vast numbers of laboratory tests result from a large number of subjects who are assumed to be representative of the population. With automatic counters or the flow cytometry an automated CBC can be performed quickly. However, if the results from an automated cell count indicate the presence of abnormal cells or if there is a reason to suspect that abnormal cells are present, then a blood smear will be collected \cite{Loddo2016}. A blood smear is often used to categorize and/or identify conditions that affect one or more types of blood cells and to monitor individuals undergoing treatment for these conditions. The results of a blood smear typically include a description of the cells appearance, as well as any abnormalities that may be seen on the slide. The manual analysis of blood smears is tedious, lengthy, repetitive and it suffers from the presence of a non-standard precision because it depends on the operator's skill. The use of image processing techniques can help to analyse, count the cells in human blood and, at the same time, to provide useful and precise information about cells morphology.

Peripheral blood smears analysis is a common and economical diagnosis technique by which expert pathologists may obtain health information about the patients. Although this procedure requires highly trained experts, it is error-prone and could be affected by inter-observer variations. Moreover, blood cells images taken from microscope could vary in their illumination and colouration conditions. Typical blood cells images contain three main components of interest: the platelets (or thrombocytes), the red blood cells (or erythrocytes) and the white blood cells (or leukocytes). It is worth considering that blood cells exist with different shapes, characteristics and colourations, according to their types.

Many tests are designed to determine the number of erythrocytes and leukocytes in the blood, together with the volume, sedimentation rate, and haemoglobin concentration of the red blood cells (blood count). In addition, certain tests are used to classify blood according to specific red blood cell antigens, or blood groups. Other tests elucidate the shape and structural details of blood cells and haemoglobin and other blood proteins. Blood can be analysed to determine the activity of various enzymes, or protein catalysts, that either are associated with the blood cells or are found free in the blood plasma.
Blood also may be analysed on the basis of properties such as total volume, circulation time, viscosity, clotting time and clotting abnormalities, acidity (pH), levels of oxygen and carbon dioxide, and the clearance rate of various substances. There are special tests based on the presence in the blood of substances characteristic of specific infections, such as the serological tests for syphilis, hepatitis, and human immunodeficiency virus (HIV, the AIDS virus) \footnote{https://www.britannica.com/topic/blood-analysis}.

Among the several available blood tests, the most common are certainly the blood cells counts, e.g., a CBC is a measure of the haematologic parameters of the blood. Included in the CBC is the calculation of the number of red blood cells (red blood cell count) or white blood cells (white blood cell count) in a cubic millimetre $(mm^{3})$ of blood, a differential white blood cell count, a haemoglobin assay, a hematocrit, calculations of red cell volume, and a platelet count. The differential white blood cell count includes measurements of the different types of white blood cells that constitute the total white blood cell count: the band neutrophils, segmented neutrophils, lymphocytes, monocytes, eosinophils, and basophils. A specific infection can be suspected on the basis of the type of leukocyte that has an abnormal value.

Human malaria infection is not strongly related to cells count but it needs different tests in order to be identified. It can only be caused by parasitic protozoans belonging to the Plasmodium type. The parasites are spread to people through the bites of infected female Anopheles mosquitoes, called "malaria vectors".
There are five parasite species that cause malaria in humans and two of these species, Plasmodium falciparum and Plasmodium vivax, constitute the greatest threat. Plasmodium ovale, Plasmodium malariae and Plasmodium knowlesi are the three remaining species which are less dangerous in human \cite{WHO_dec_2016}, as shown in fig.\ref{fig:malaria_types}.
All five species may appear in four different life-cycle stages during the infection phase in peripheral blood: ring, trophozoite, schizont and gametocyte. Some examples are shown in fig.\ref{fig:malaria_stages}.
The life-cycle-stage of the parasite is defined by its morphology, size and the presence or absence of malarial pigment.
The species differ in the changes of infected cell's shape, presence of some characteristic dots and the morphology of the parasite in some of the life-cycle-stages \cite{Somasekar2011}.

Computer vision techniques for malaria diagnosis and recognition represent a relatively new area for early malaria detection and, in general, for medical imaging, able to overcome the problems related to manual analysis, that is performed by human visual examination of blood smears. The whole process requires an ability to differentiate between non-parasitic stained components/bodies (e.g. red blood cells, white blood cells, platelets, and artefacts) and the malarial parasites using visual information. If the blood sample is diagnosed as positive (i.e. parasites present) an additional capability of differentiating species and life-stages (i.e. identification) is required to specify the infection. Numerous methods of automatic malaria diagnosis have been proposed so far, in order to overcome the issues before mentioned. The aim of this paper is to review and analyse the works of different researchers who in particular have used mathematical morphology as a tool for computer aided malaria detection and classification.

\begin{figure}[!t]
	\centering
	\includegraphics[width=2.75cm]{img/f3_Pfalciparum}
	\includegraphics[width=2.75cm]{img/f3_Pvivax}
	\includegraphics[width=2.75cm]{img/f3_Povale}
	\includegraphics[width=2.75cm]{img/f3_Pmalariae}
	\caption{\label{fig:malaria_types}Types of human malaria: from left to right, P. Falciparum, P. Vivax, P. Ovale, P. Malariae \cite{Khan2011}.}
\end{figure}

\begin{figure}[!t]
	\centering
	\includegraphics[width=2.75cm, height=2.75cm]{img/1_Pfalciparum_ring}
	\includegraphics[width=2.75cm, height=2.75cm]{img/2_Pfalciparum_trophozoite}
	\includegraphics[width=2.75cm, height=2.75cm]{img/3_Pfalciparum_gametocyte}
	\includegraphics[width=2.75cm, height=2.75cm]{img/4_Pfalciparum_schizont}
	\includegraphics[width=2.75cm, height=2.75cm]{img/5_Pmalariae_ring}
	\includegraphics[width=2.75cm, height=2.75cm]{img/6_Pmalariae_schizont}
	\includegraphics[width=2.75cm, height=2.75cm]{img/7_Povale_trophozoites}
	\includegraphics[width=2.75cm, height=2.75cm]{img/8_Pvivax_trophozoites}
	\includegraphics[width=2.75cm, height=2.75cm]{img/9_Pvivax_ring}
	\includegraphics[width=2.75cm, height=2.75cm]{img/10_Pvivax_gametocyte}
	\includegraphics[width=2.75cm, height=2.75cm]{img/11_Pvivax_ring}
	\caption{\label{fig:malaria_stages}Examples of malaria parasite stages. From top left: P.falciparum ring, trophozoite, gametocyte, schizont; P.malariae ring and schizont; P.ovale and P.vivax trophozoites; P.vivax ring and gametocyte; P.vivax ring \cite{Tek2010}}
\end{figure}

\subsection{Mathematical morphology}
Mathematical morphology (MM) can be defined as a theory for the analysis of spatial structures. It is called morphology because it aims at analysing the shape and form of objects. It is mathematical in the sense that the analysis is based on set theory, integral geometry, and lattice algebra. MM is not only a theory, but also a very powerful image analysis technique \cite{Soille2004}.
It was introduced by Matheron as a technique for analysing geometric structure of metallic and geologic samples. It refers to a branch of non-linear image processing and analysis that concentrates on the geometric structure within an image.
The morphological filter, which can be constructed on the basis of the underlying morphological operations, are more suitable for shape analysis than the standard linear filters since the latter sometimes distort the underlying geometric form of the image. Some of the salient points regarding the morphological approach are as follows \cite{Giardina1988}:
\begin{enumerate}
	\item Morphological operations provide for the systematic alteration of the geometric content of an image while maintaining the stability of the important geometric characteristics.
	\item There exists a well-developed morphological algebra that can be employed for representation and optimization.
	\item It is possible to express digital algorithms in terms of a very small class of primitive morphological operations.
	\item There exist rigorous representations theorems by means of which one can obtain the expression of morphological filters in terms of the primitive morphological operations.
\end{enumerate}
Dilation and erosion are the basic morphological processing operations. They are defined in terms of more elementary set operations, but are employed as the basic elements of many algorithms. Both dilation and erosion are produced by the interaction of a set called structuring element with a set of pixels of interest in the image. The structuring element has both a shape and an origin.

In this paper we present a review of computer-aided methods oriented to malaria parasites detection and segmentation by mathematical morphology based techniques.
The rest of the paper is organised as follows. Section 2 introduces the adopted materials and methods for malaria-infected peripheral blood cells images analysis. It has been structured in preprocessing, segmentation, feature extraction, feature selection and classification subsections, consistently with a typical pipeline of a computer-aided image analysis process. All the considered works make use of morphological operators in at least one of the phases of image analysis. Section 3 contains an overall discussion about the methods and the conclusions are expressed in section 4.

\section{Materials and methods}
This section presents a review of some of the main recent studies existing in literature regarding the analysis of malaria infected blood smears using mathematical morphology.
A typical approach usually comprises five different image processing and analysis tasks, as follows:
\begin{enumerate}
	\item Pre-processing.
	\item Segmentation.
	\item Feature extraction.
	\item Feature selection.
	\item Classification.
\end{enumerate}
Since morphological techniques have been used in the first three phases, the reviewed works have been divided into the following sub-sections: pre-processing, segmentation and feature extraction.
Each sub-section contains description about methods that cope with malaria parasites (MP) stained components analysis, both on thin and thick blood smears, without distinction.

%\subsection{Search of related literatures}
Extensive search of articles has been made in PubMed and Google Scholar search engines based on the keywords: "malaria, mathematical morphology, automated malaria diagnosis" up to October  2017. The search includes papers published in English and titles and abstracts of potentially relevant studies were selected and presented from the most recent ones. Thereafter, the full texts of these studies were evaluated as per the exclusion criteria.

%TODO inserire descrizione tabella "una sintesi dei lavori presi in considerazione è mostrata nella tabella X, bla bla bla ... Some works are missing because...

%\subsection{Blood smear staining techniques}
Two main factors are generally considered if we refer to staining techniques: the type of colouration, in which Giemsa and Leishman are the most common, and the thickness of blood slide, which may be thick or thin.
The majority of studies have been employed on thin blood smear images (over two-third of the total count) while only a few have used thick blood smear images.
Typically, thin smears permit the identification of specific parasitic stage and quantification of malaria parasite; on the other hand, thick smears are better if the target is to perform an initial identification of malaria infection using blood pathology. Some examples are shown in fig. \ref{fig:thick_thin_film}.
Giemsa stained blood smear is considered in most of the analysed literatures whereas Leishman stain is considered in few studies, as showed in table \ref{tab:summary_table}.

It is reported that Leishman stain has bigger sensitivity for parasite detection than Giemsa \cite{Khan2014} and is superior for visualization of red and white blood cell morphology. \cite{Sathpathi2014}. On the contrary, Giemsa stain highlights both malaria parasites and white blood cells and, therefore, it is an additional issue to deal with.
Giemsa stain is much costly and also time-taking procedure than Leishman.
Moreover, magnification of 100X by using an oil immersion objective is used for capturing microscopic images of thin blood smear for identification of specific parasites and their infected stage.

\begin{figure}[!t]
	\centering
	\includegraphics[height=0.15\textheight]{img/f2_ThickGiemsa}
	\includegraphics[height=0.15\textheight]{img/f2_ThickLeishman}
	\includegraphics[height=0.15\textheight]{img/f2_ThinGiemsa}
	\includegraphics[height=0.15\textheight]{img/f2_ThinLeishman}
	\caption{\label{fig:thick_thin_film}Malaria infected blood smear images. From top left, clockwise: thick smear with Giemsa stain, thick smear with Leishman stain, thin smear with Leishman stain, thick smear with Giemsa stain \cite{Das2015}. Arrows in top images indicate chromatin dots while, in the bottom ones, they show the infected erythrocytes.}
\end{figure}

\subsection{Preprocessing}
In image analysis field, especially when we refer to complex computer-aided pipelines, pre-processing methods are particularly used in order to improve the image data by suppressing unwanted noise or enhancing some image features for further processing.
%Typically, pre-processing methods are classified according to the following schema:
%\begin{enumerate}
%	\item pixel brightness transformations;
%	\item geometric transformations;
%	\item pre-processing methods that use a local neighbourhood of the processed pixel;
%	\item image restoration that requires knowledge about the entire image.
%\end{enumerate}
It is worth to mention pre-processing methods because they are an important step regarding image analysis field but, for what concerns the malaria-affected blood image analysis, in our review we particularly found methods which operate for illumination correction and noise filtering purposes.
Generally speaking, digital microscopy images can be acquired in different lighting conditions, with several types of acquisition devices or from blood smears stained with various staining protocol and, consequently, the features of similar images could differ a lot.
Different techniques for illumination correction have been suggested to reduce such variation, e.g., a lot of authors work with grayscale-converted images as an illumination correction method %\cite{Tek2010},\cite{Das2015},\cite{Tek2006}.
On the other hand, noise filtering aims to remove the noise introduced by mishandling the slides and/or the camera settings.

The system proposed in \cite{Oliveira2017} is based on image processing, artificial intelligence techniques and an adapted face detection algorithm to identify Plasmodium parasites. The algorithm uses the integral image and haar-like features concepts, and weak classifiers with adaptive boosting learning. The search scope of the learning algorithm is reduced in the preprocessing step by removing the background around blood cells.
%Morphological operations have been applied to images to enhance the geometric structures on a set of pixels. In particular, an erosion operation has been applied after a Gaussian filtering in order to locate the RBCs and restrict the search space of the algorithm.
%The accuracy of the system was, on average, 91, meaning that for every 100 parasite-infected samples, 91 were identified correctly. The system has been developed for mobile phones and tablets and can represent a useful tool
%in health centers in remote communities without the need for specific expertise.

In \cite{Romero2016} an algorithm that uses morphological operations, the watershed method, the Hough transform and the clustering method of k-means to detect overlapped RBCs is presented. In the pre-processing stage white blood cells and platelets are removed before the segmentation task. During this step, some noise, the WBC cytoplasm and platelets still remain on the image. Therefore, the small objects are removed using a morphological opening and then the image is dilated with a disc-shaped structuring element.

An edge-based segmentation of erythrocytes infected with malaria parasites using microscopic images is proposed in \cite{Somasekar2015}.  A fuzzy C-means clustering is applied to extract infected erythrocytes, which is further processed for the final segmentation. The experimental results showed that the proposed method can gain 98\%, 93.3\%, 98.65\% and 90.33\% of sensitivity, specificity, prediction value positive and prediction value negative, respectively. A morphological erosion is used to erase some small noises and spots before the segmentation and holes inside the infected erythrocytes are filled using a morphological hole filling operation for the final segmentation.
The work \cite{Reni2015} proposes a new algorithm for morphological filtering of the blood images as a pre-processing tool for segmentation. Conventional morphological closing on blood images removes the unwanted components but also useful information. On the opposite the proposed method preserves the necessary information of foreground components while removing noise and artefacts.
The paper in \cite{Kareem2012_1} illustrates a morphological approach for blood cell identification and uses the image features such as intensity, histogram, relative size and geometry for further analysis. Before the identification of blood cells, the authors  propose a novel morphological filtering based on the size of RBC for platelets and/or artifacts elimination. A dilation is performed by a concentric ring structuring element and erosion by disk shaped structuring element. The radius of the structuring element depends on the radius of the RBC, so that all the components smaller than the RBCs are removed.
In \cite{Tek2010} a complete framework to detect and identify malaria parasites in images of Giemsa stained thin blood film specimens is proposed. Also, the system is able to identify the infecting species and life-cycle stages.
The preprocessing step of the proposed method is applied to reduce the variations in the observed size, intensity, and colour of the cells and stained objects before the detection and classification steps. The aim is to correct the non-uniform illumination in the images. The estimation is based on a morphological closing operation using a sufficiently large structuring element. The "sufficiently" large size for an input image is determined automatically with respect to its average cell size computed from the area granulometry distribution.
In the method proposed in \cite{Sheik2013} the first phase is the stain object extraction which extracts candidates objects that can be infected by malaria parasites by using intensity and colour. Before extracting the stained objects the method firstly extracts the foreground and enhanced images. Foreground image is a binary image which is produced after applying morphological hole filling on such pixels which have lower intensity value than average intensity value of green layer. After the stained objects extraction process, a series of morphological operations is also employed after every part in order to eliminate small components and complete the final stained objects.

%Median filter is often used for reducing impulse noise. Several studies have used it to enhance microscopic images of peripheral blood smear towards characterization of malaria followed by adaptive or local histogram equalization. Local low pass filter and local adaptive histogram equalization techniques have also been applied to enhance the pathological image quality. Das et al. \cite{Das2015} showed that geometric mean filter provides better performance towards enhancing peripheral blood smear images. Another preprocessing method like space-time manifolds-based image mosaicing has been used to obtain large field of view \cite{Zou2010} towards identifying malaria parasites, while Di Ruberto et al. \cite{Put16c} proposed an approach to overcome the problem of uneven illumination conditions in image acquisition. For this purpose they designed an illumination pattern that simulates the classic visual defects introduced by the digital microscope lenses, that
%is the vignetting effect. Starting from the smallest radius, they applied different illumination patterns, created modifying the radius of the Gaussian curve, to the original images. Then, a similarity value measures the difference in terms of pixels between the original images and the corrupted ones.

\subsection{Segmentation}
Segmentation is a key step in image analysis because it permits the identification and separation of the regions that compose an image, according to certain criteria of homogeneity and separation. Its main target is to divide the image into parts that have a strong correlation with objects or areas of the real world contained in the image.
The commonly used segmentation methods essentially operate considering characteristics such as the brightness value, colour and reflection of the individual pixels, identifying groups of pixels that correspond to spatially connected regions. As for many problems of image processing, there is no standard solution valid in general, so different segmentation techniques can be applied, according to the characteristics of the images to process and of the objects to segment.
Medical images segmentation is typically performed using two main strategies: the first level aims to separate whole cells or tissues from the background and the second one aims to separate the tissue structure in different regions or the cell in their components as the nucleus from the cytoplasm or intracellular parasites. The latter case is commonly used in applications in which the cell class depends on the morphological characteristics of its components.
%The segmentation techniques can generally be divided into three main categories: Pixel Based, Edge Based and Region Based.
%
%%\subsection{Erythrocytes segmentation}
%%\subsubsection{Thresholding}
%%THRESHOLDING METHODS

Several other authors attempted to use thresholding as segmentation method in their computer-aided systems.
In \cite{Ross2006} the authors propose a method which provides a positive or negative diagnosis of malaria and differentiates parasites by species. The segmentation step relies on a thresholding strategy which aims to identify and segment potential parasites and erythrocytes from the image background after a six steps threshold selection. Mathematical morphology has been used for parasite size estimation, erythrocytes reconstruction and cells bigger than erythrocytes removal.
%
%Makkapati and Rao proposed two different methodologies. The first one \cite{Makkapati2009} is based on HSV colour space with the purpose of performing a red blood cells (RBCs) segmentation and an identification of RBCs affected from MP with a chromatin dot identification by thresholding the Saturation channel. The RBCs segmentation is achieved by dividing the entire hue range of $360^{\circ}$ into six segments and finding the dominant colour type to be the representative of the background. On the other hand, the second work \cite{Makkapati2011} is devoted to the detection of the parasite stage and identification of the species. It is based on ontological description and uses the morphology of the two basic components of a parasite, namely, chromatin dot(s) and cytoplasm. The authors state that the green component of an image offers maximum separation between the background and RBC pixels. The green component is first segmented to detect the RBCs by choosing an optimal threshold [10]. The pixels belonging to chromatin dots have higher saturation values and are segmented by thresholding the saturation component of the image. However, several artefacts may also be segmented as chromatin
%dots, so a further evaluation is performed to establish if a chromatin dot is valid by identifying whether it is inside a RBC or not.
%
The authors in \cite{Kareem2011} used the Annular Ring Ratio (ARR) transform method. Before applying it, a pre processing phase for removing platelets, parasites and other artefacts in the image has been performed. In the proposed method, the image after being converted to grayscale undergoes a morphological opening similar to closing. Unlike conventional closing (dilation followed by erosion) which uses the same structuring element, two different structuring elements are used, a concentric ring for dilation and a disk for erosion. The inner and outer diameter of the dilation ring is set to 35\% and 70\% of RBCs size, respectively and the erosion disk has the same diameter. Therefore, considering that fixed manually defined parameters are used for this strategy, the results may substantially differ depending on the image resolution. This approach results in locating only the stained components in the image instead of all the cells and hence will not only speed up the operation but reduces the complexity.
%
%The morphology of P.falciparum at gametocyte stage is the key parameter in the segmentation approach proposed in \cite{Kareem2012}. It is based on a modification of the ARR transform. The method can be used as an efficient tool for gametocyte detection for post-treatment malaria diagnosis. It also identifies the presence of any White Blood Cells (WBCs) in the image, and discards other artefacts and non infected cells. It utilizes the information based on structure, colour and geometry of the cells and does not require any segmentation or non-illumination correction techniques that are commonly used for cell detection.
%
%Fang et al. \cite{Fang2011} used a threshold to distinguish the unhealthy cells from the healthy cells after a malaria detection phase performed with a phase spectrum methodology. This method uses a the Quaternion Fourier Transform (QFT) to obtain the amplitude and phase spectrum for blood smear images. Then, the reconstructed image is obtained using the Inverse Quaternion Fourier transform (IQFT) on a constant amplitude spectrum and the original phase spectrum. The authors concluded that B and G channels from the RGB colour space and I channel from the HSI colour space are the most sensitive channels for MP detection.
%
Anggraini et al. \cite{Anggraini2011} proposed a method for separating blood cells, parasites and other components from background in a microscopic field of a thin blood smear. They applied several global thresholding methods and visually compared the results to qualitatively determine which technique yields the best result. The binary image was then subjected to hole filling morphological operator and applied as marker to label blood cells. From each identified cell (RBC and WBC), constituents of the parasite (nucleus and cytoplasm) were extracted using multiple threshold.
%
Savkare et al. \cite{Savkare2011} worked on thin blood films with Giemsa staining and used global threshold and Otsu threshold \cite{Otsu1975} on grayscale enhanced image (green channel) for separating foreground from background. Hole filling has been performed on identified cells and morphological operators have been used to identify overlapping cells. Then, watershed transform has been applied for separating overlapped cells.
%
In \cite{Ghosh2011}, authors used divergence based threshold selection in order to segment P.vivax parasites from Leishman-stained thin blood films. This method is based on Cauchy membership function \cite{Put14a} and is applied to the C channel of CMYK colour space. Morphological operators of opening and closing have been used for artefacts removal.
%
Tek et al. \cite{Tek2010} addressed the problem of parasite detection and identification on thin blood film. They performed a pre-processing phase, in which morphology has been used with a granulometry-based cell size estimation, considering the peak index of the granulometry distribution as the average cell area.
Then, two different segmentation steps followed: the foreground-background segmentation has been realized using morphological area top-hats (using the average cell area value) and morphological double thresholding; the stained pixels segmentation has been performed by modelling the stained and unstained pixel distribution with RGB space histograms and used the probability density function to determine if a pixel is stained or not. This work also addresses non-uniform illumination problem, removing it by using a pre-recorded illumination image or using a morphological closing operation.
%
%Malihi et al. \cite{Malihi2013} proposed a method for malaria parasite detection in Giemsa-stained blood samples. They used both binary image masks obtained with Otsu method and the edge detection images from Canny method for RBC segmentation, followed by a hole filling process.
%
Authors in \cite{Arco2015} worked on thick blood films and proposed a method that uses an adaptive thresholding based scheme, which also allows an effective classification of pixels. This means that the election of whether a pixel belongs to the background or to the signal (parasites and white blood cells) is only established by the pixels around it, that is its neighbourhood. Then, morphological methods are applied to evaluate the area of connected components, labelling those belonging to parasites and counting their number. The following strategy has been used:
\begin{enumerate}
	\item an algorithm based on run-length encoding has been used for analysing the connected components of the binary masks obtained in segmentation phase;
	\item preliminary tagging, keeping the label equivalences or connected regions with different labels in a table of local matches. Resolution of the equivalences is, then, performed;
	\item measurements on the remaining connected components: WBCs and MP region areas are checked and WBCs connected components are removed with a closing operation.
\end{enumerate}
%
Also in the method proposed in \cite{Somasekar2017} the segmentation of the infected parasites is based on thresholding. The segmentation is achieved in two stages by maximizing between-class variance of an original image and consequently by an iterative threshold selection from a stage-one threshold image with suitable stopping criteria. The segmented results are post processed to improve the accuracy of the detection of malaria
parasites by morphological operators (erosion and closing).

Mushabe et al. \cite{Mushabe2013} used morphological and statistical classification to detect malaria in blood smears by identifying and counting red blood cells and Plasmodium parasites. Morphological operations and histogram-based thresholding are used to extract RBCs and boundary curvature calculations and Delaunay triangulation are used for splitting clumped RBCs. They worked on Giemsa-stained thin blood smears.
%
The proposed automated method in \cite{Elter2011} for parasite detection and identification worked on thin blood film acquired with Giemsa stain. The authors found that the G and B channels of the RGB colour are very good features to identify objects containing chromatin in Giemsa stained blood films, being not only considered highly discriminative but also almost independent of differences in illumination and staining intensity. They transformed the colour input image into a monochrome image \textit{I(x,y)}, that highlights objects containing chromatin: $ I(x,y) = arctan \frac{I_{green}(x,y)}{I_{blue}(x,y)} $. In this work, mathematical morphology has been used with a black top-hat operator to separate MP from both leukocytes and platelets, with a non-flat paraboloid structuring element of radius of 9 and a slope of 1 pixel. It should be taken into account that these fixed parameters might not be suitable for images with different pixel resolutions. The black top-hat operator is followed by a thresholding operation with a fixed threshold, which according to the authors is reliable given the independence of the G and B channels with regard to illumination and staining intensity. However, the authors do not define the value of this fixed threshold on the publication.
%
%% Morphology and/or granulometry in Segmentation stage:
Di Ruberto et al. \cite{DiRuberto2002} aimed to detect the parasites by means of an automatic thresholding based on a morphological approach applied to cell image segmentation, that is more accurate than the classical watershed-based algorithm. They have applied grey scale granulometries based on opening with disk-shaped elements, flat and hemispherical. They have used a hemispherical disk-shaped structuring element to enhance the roundness and the compactness of the red cells improving the accuracy of the classical watershed algorithm, while they have used a disk-shaped flat structuring element to separate overlapping cells. These methods make use of the red blood cell structure knowledge, that is not used in existing watershed-based algorithms.
%
Soni et al. \cite{Soni2011} performed segmentation of erythrocytes by using granulometry as well. The size and eccentricity of the erythrocytes are also required for the calculation of some feature values (as these can be indicative of infection). The shape of the objects (circular erythrocytes) is known a priori, but the image must be analysed to determine the size distribution of objects in the image and to find the average eccentricity of erythrocytes present.
Gray-scale granulometries based on opening with disk shape elements are then used. Non flat disk shaped structural element are applied to enhance the roundness and compactness of the red blood cells and flat disk shaped structural element applied to segment overlapping cells. The object to be segmented differs greatly in contrast from the background image. Changes in contrast can be detected by operators that calculate the gradient of an image. The gradient image can be computed and a threshold can be applied to create a binary mask containing the segmented cell. The binary gradient mask is dilated using the vertical structuring element followed by the horizontal structuring element. The cell of interest has been successfully segmented, but it is not the only object that has been found. Any objects that are connected to the border of the image can be removed.
%
Also Airwhar et al. \cite{Ahirwar2012} based their approach on thresholding and granulometry. The histogram of the complemented, green component has been used and it is said to be a bimodal distribution in all the considered images. Then, both local and global thresholds are used, and the union of the two binary images is chosen as the parasite marker image. A morphological opening filter, using a disk-shaped SE with radius equal to the mean erythrocyte radius less the standard deviation, is applied to the grayscale morphologically filtered green component of the image to remove any objects smaller than an erythrocyte.
The morphological gradient is then calculated using a diamond-shaped SE with unity length. The segmentation method is applied to each object in the reconstructed binary image of erythrocytes individually. Those objects that do not exceed the area of a circle with radius equal to the mean erythrocyte radius plus the standard deviation are regarded as being single cells, and are unmodified.
On the other hand, the clumped cells are segmented as follows. First, the intersection of the morphological gradient image and the dilated cell cluster is taken. This image is then transformed to a binary image by thresholding any value greater than zero. A series of morphological operations, namely a closing operation, thinning, and spur removal are then applied to generate a contour of the segmented erythrocytes. The contours are filled, and the segmented mask is again reconstructed with the valid parasite marker image to result in a segmented mask of infected cells.
%
Several authors used marker controlled watershed with morphological approach.
Khan et al. \cite{Khan2011}, among other experimentations, used it in order to try to separate overlapping cells because, according to their statements, watershed transform can separate touching cells but it is not sufficient for overlapping cells.
In \cite{Springl2009} red blood cell segmentation has been performed by using marker-controlled watershed transformation based on the image gradient. Markers are computed as a combination of the binary mask of the red blood cells and centres of the cells which are computed using a similar algorithm that was utilized for the evaluation of the average cell radius. The binary mask is obtained by thresholding the grey-scale image with an automatically estimated threshold using Otsu method\cite{Otsu1975}.
Das et al. \cite{Das2011}, \cite{Das2013}, \cite{Das2012}, \cite{Das2015} segmented erythrocytes as aforesaid and then morphological operators are used to eliminate unwanted cells like leukocytes and platelets. To conclude, overlapping erythrocytes are segmented by using marker controlled watershed segmentation technique.

In the paper \cite{Devi2017} the authors propose a computer assisted system for quantification of erythrocytes in microscopic images of thin blood smears. The proposed method consists of preprocessing, segmentation, morphological filtering, cell separation and clump cell segmentation. The performance of the system in classifying the isolated and clump erythrocytes by geometric features is evaluated for the different classifiers. Moreover, the clump erythrocytes are segmented using marker controlled watershed with h-minima as internal marker. Based on the experimental results, it may be concluded that the proposed model provides satisfactory results with an accuracy of 98.02\% in comparison to the state of art.

%
%%BOUNDARY BASED METHODS \subsubsection{Boundary based methods}
%Zou et al. \cite{Zou2010} based their method on the consideration that RBCs have a circular shape. They propose an improved version of the Circle Hough Transform (CHT) for RBCs detection. The authors state that overlapping and oval cells can be detected as well as non-clumped RBCs and affirm that the green channel of the RGB colour space permits to better distinguish RBCs. However, some RBCs may not have circular shape and it could affect the detection results.
%
In \cite{Yunda2012}, Yunda et al. proposed a method for P.vivax parasites detection. The segmentation phase is a combination of border and region detection that allows rejection of the image background and permits identifying each of the objects. Initially, the morphological gradient method is used to enhance the borders of previously found objects. This is followed by a threshold detection stage using the K-Median method.
Furthermore, Laplacian operator was used to discriminate the pixels that are interior or exterior in relation to the regions of the images and then erosion operation followed by two dilations were applied to delete the pixels which did not make part of any object. In the end, Absence of Gradients and Nernstian Equilibrium Stripping (AGNES) and K-Median techniques were applied to assign the remaining number of pixels to each region, using the image regions previously identified as objects and background as the starting point.
%
%%CLUSTERING METHODS \subsection{Clustering methods}: no method uses mathematical morphology applied to clustering.
%
%%CLASSIFIER-BASED METHODS \subsection{Classifier-based methods}
Tek et al. \cite{Tek2006} proposed a classifier-based method, for the segmentation stage, which relies on a Bayesian pixel classifier to distinguish among stained and non-stained pixels. In particular, they used a non-parametric method based on histograms in order to produce the probability density functions of stained and non-stained classes. Stained pixels can belong to other components such as WBCs, platelets or artefacts, in addition to the parasites and so the detection procedure requires a further classification to distinguish among parasite and non-parasite pixels. However, the stained pixels have to be represented as connected sets, representing stained objects, to extract features for the classifier. Furthermore, top-hat extraction and infinite reconstruction were applied to find the regions that include the objects.
%
\subsection{Feature extraction}
Feature extraction has the target of reducing the computational complexity of the subsequent process and facilitating a reliable and accurate recognition for unknown novel data, considering that the input data to an algorithm could be too large to be processed and it could be redundant (e.g. repetitiveness of pixels patterns in an image). Moreover, the in-depth understanding of the domain-specific knowledge gained by human experts on the problem being addressed can be of extreme importance for the design of a reliable and effective feature extraction engine \cite{Jiang2009}.
It starts from determining a subset of the initial features and this procedure is called feature selection. The selected features are expected to contain the relevant information from the input data, so that the desired task can be performed by using this reduced representation instead of the complete initial data.
Malaria parasite infection causes micro structural changes in erythrocytes. The microscopic features of the RBCs are usually specific to morphology, intensity and texture. They may also represent the differences that occur among healthy and unhealthy cells. Most of the studies have reported both textural and geometric features for describing malaria infection stages \cite{Das2015}.
Generally speaking, features may be distinguished according to the following characteristics: morphological features and textural and intensity features.\\
It is a well known mathematical morphology approach to compute a size distribution of grains in binary images, using a series of morphological opening operations. It is the basis for the characterization of the concept of size. Some authors used area granulometry for pre-processing purposes in malaria characterization \cite{Tek2010} even though it is certainly effective for extracting cells size features information \cite{Tek2006},\cite{Malihi2013},\cite{Springl2009}.
%
%
%%\input{Test.tex}
%
%\section{Discussions}
%A comprehensive synthesis of the analysed works is presented in table \ref{tab:summary_table}. It only considers the methods which employed mathematical morphology in at least one step of the pipelines and it has been structured by considering the following information: staining technique, segmentation methods, extracted features, classification strategy, specificity and sensitivity percentages relative to the classification steps.
%Most of the studies are based on P. vivax and P. falciparum characterization.
%With regards to the showed approaches and related results, it is clear that malaria parasites detection and segmentation techniques in microscopic images needs further experiments and improvements. In general, the analysed works have been tested with a limited number of images and the datasets are not publicly available; therefore, a comparison between different approaches is very difficult. Despite promising results reported during the past years, the great majority of the computer-aided methods found on the literature for malaria diagnosis are based on images acquired under well controlled conditions and with proper microscopic equipment. However, one should take into account that 80\% of malaria cases occur in Africa, where this type of equipment is scarce or even nonexistent in common healthcare facilities \cite{Rosado2016}.
%Moreover, this review showed that P. falciparum is the most analysed if we refer to segmentation and detection, considering that it is the most widespread among malaria parasite types.
%The majority of the works used thin blood smear. It is typically used for identification of malaria infected
%stages, types of parasitic infection and percentage of parasitemia, while thick blood smear is used for identification of and quantification of malaria parasite count against leukocyte count per microliter blood.
%Pre-processing phase is typically taken on with filters and the most used in the analysed works is certainly the median filter which permits to preserve sharp edges. Apart from the classic histogram equalization and contrast stretching techniques, other filters have been employed, e.g., geometric mean filter to remove Gaussian noise preserving edges, Laplacian filter, in order to find edges, and so on.
%Median filter has been found to be effective for reducing impulse noises from the microscopic images, even though recent studies have shown that geometric mean filter provides better performance than the median filter \cite{Das2013}, \cite{Das2015}.
%
%Malaria parasites may be discriminated according to two different strategies: by segmenting the whole erythrocyte from the blood smear image on the basis of which malaria infection is detected, otherwise by segmenting chromatin dot or parasite infection region for characterizing parasite infection stages based on some extracted target features.
%In general, thresholding-based approach is still widely used for segmentation purposes. In particular, a lot of authors affirm that Otsu thresholding suffers from limitations when textural variation is high, while histogram thresholding can not deal sufficiently good in identifying valley regions in case of unimodal histograms. Another greatly used segmentation approach is clearly the watershed transform. The classic watershed approach is reported to produce over segmentation results \cite{Savkare2011}, whereas the marker controlled approach does not suffer from this issue and it is reported to be very effective for overlapping cells segmentation even though some authors affirm that it may fail to segment highly overlapped cells \cite{Khan2011}, \cite{Springl2009}. Other authors \cite{Ahirwar2012}, \cite{DiRuberto2002}, \cite{Soni2011}, employed granulometry and stated that it is very effective to segment cells with regular size.
%The analysed works performed classification phase for different purposes. The majority of them aimed to distinguish among two classes \cite{Anggraini2011}, \cite{Annaldas2014},  \cite{Das2015}, \cite{Fang2011}, \cite{Ghosh2011}, \cite{Kareem2012}, \cite{Mushabe2013}: malaria infected and noninfected RBCs, whereas more complex classification strategies aimed to classify from three \cite{Das2012}, \cite{Maity2012} to up to seven different classes \cite{Das2013}, \cite{Das2015}, \cite{Makkapati2011}, \cite{Malihi2013}.
%
%\section{Conclusions}
%This work reviewed several computational microscopic imaging techniques oriented to mathematical morphology approach, proposed in literature for malaria parasites detection and segmentation in blood smear microscopic images.
%
%The computer vision methodologies reported in the literature are based on light microscopic images of human peripheral blood smears for computer-aided detection of malaria parasite and its various stages. Image pre-processing, segmentation of erythrocytes and parasites, malaria parasite feature extraction, malaria detection techniques have been discussed here.
%
%It is worth to remember that mathematical morphology techniques have been widely used for image processing purposes. Among the application fields, it has been applied for fingerprint feature extraction, recognition of handwritten digits, license plate detection, border extraction, denoising using morphological filters, text extraction and detection of imperfection in printed circuit boards \cite{Kaur2013}. Apart from this kind of fields, mathematical morphology has been employed in biomedical image analysis, especially in pre-processing and segmentation techniques.
%
%Morphological cell analysis is used to face off abnormality identification and classification, early cancer detection. It has been integrated in new methods for biomedical applications,
%such as automatic segmentation and analysis of histological tumour sections, boundary detection of cervical cell nuclei considering overlapping and clustering, the granules segmentation and spatial distribution analysis, morphological characteristics analysis of specific biomedical cells, understanding the chemotactic response and drug influences, or identifying cell morphogenesis in different cell cycle progression.
%Morphological feature quantification for grading cancerous or precancerous cells is especially widely researched in the literature, such as nuclei segmentation based on marker-controlled watershed transform and snake model for hepatocellular carcinoma feature extraction and classification, which is important for prognosis and treatment planning, nuclei feature quantification for cancer cell cycle analysis, and using feature extraction including image morphological analysis, wavelet analysis, and texture analysis for automated classification of renal cell \cite{Chen2012}.
%
%Moreover, non-linear filtering has become increasingly important in many image processing applications. Initially, the attraction to non-linear filters was mostly limited to the impulse-removing and edge-preserving qualities of the median filter. However, as the number and sophistication of non-linear filters have increased, so has the variety of applications for these filters. The shape-based methods of mathematical morphology, in particular, are now used in a wide variety of medical applications, including electrocardiography, ultrasound imaging, radiology, and histologic image analysis \cite{Schulze1994}.
%
%Furthermore, microscopic image analysis and, in particular, malaria detection and classification can greatly benefit from the use of mathematical morphology. The interest in this approach to image processing ad analysis is proved by the increasing number of works proposing methods for malaria image analysis based on mathematical morphology techniques.
%
%It is worth to mention that the development of new mobility-aware microscopic devices (and ideally low cost) is an area that can greatly improve the chances of the successful deployment of computer vision CAD solutions for malaria diagnosis in the field.
%The mobile phone is currently Africa's most important digital technology, and is boosting African health as it emerges as a platform for diagnosis and treatment.
%Considering the recent significant improvements of the new generation of mobile devices in terms of image acquisition and processing power, if a reliable automatic diagnostic performance is ensured through the usage of those devices, one would dramatically reduce the effort in the exhaustive and time consuming activity of microscopic examination.
%Moreover, the lack of highly trained microscopists on malaria diagnosis in rural areas could then be complemented by a significantly less specialized technician that knows how to operate the system and prepare blood smears. The usage of mobile devices in the system architecture can also bring significant improvements in terms of portability and data transmission. Finally, malaria diagnosis might be just one
%element of a suite of diagnostic software tests running on this type of system. Several other tests could simultaneously be carried out using the same images, for instance cell counting or detection of other hemoparasites like microfilaria or trypanosoma \cite{Rosado2016}.

\inputencoding{latin2}
\bibliographystyle{plain}
\bibliography{main}
\inputencoding{latin1}

\end{document}

### 2017-11-09 Paper.tex
Mod 1 CDR

@@
 	On the other hand, noise filtering aims to remove the noise introduced by mishandling the slides and/or the camera settings.
+	Morphological operators have been extensively used as preprocessing for image enhancement in major studies.
+	Erosion and dilation operations on raw smear images allow discarding undesired patterns and help in the selection of required cells or regions of interest.
+Morphological operators are useful for removal of unwanted
+	objects, holes filling, splitting, thinning and thinking.\texttt{Preprocessing techniques used by different researchers during automated diagnosis of malaria are listed below.}
 
@@
 After the stained objects extraction process, a series of morphological operations is also employed after every part in order to eliminate small components and complete the final stained objects.
+	In \cite{Gonzalez2016} a system to determine markers for watershed segmentation based on the Radon transform and mathematical operators is proposed.
+In the first step of the process small irrelevant structures and part of the noise are eliminated by a morphological filter, in order to ensure the preservation of the cells edges.
+Image smoothing is performed by a morphological erosion-reconstruction and dilation-reconstruction (ERDR) filter with a disc structuring element of radius equal to 20 pixels, which is $0.274$ times smaller than the average radius of the RBCs.
+In this way the influences of the size and the shape of the structures can be separated in the smoothing process.
+At the same time the objects which are not eliminated remain unchanged.
+Also, a morphological closing is performed with a disc structuring element having radius smaller than half the average of the RBCs' radii, in order to connect the possible (more than one) markers that can appear on a single cell.
 
@@
 
-\subsection{Segmentation}
+	\subsection{Segmentation of RBCs and parasites}
 	Segmentation is a key step in image analysis because it permits the identification and separation of the regions that compose an image, according to certain criteria of homogeneity and separation.
@@
 In the end, Absence of Gradients and Nernstian Equilibrium Stripping (AGNES) and K-Median techniques were applied to assign the remaining number of pixels to each region, using the image regions previously identified as objects and background as the starting point.
-%
-%%CLUSTERING METHODS \subsection{Clustering methods}: no method uses mathematical morphology applied to clustering.
-%
-%%CLASSIFIER-BASED METHODS \subsection{Classifier-based methods}
+
+	In \cite{Rosado2017} Rosado et al.
+proposed a system using supervised classification to assess the presence of malaria parasites and determine the species and life cycle stage in Giemsa-stained thin blood smears.
+For the RBCs segmentation, they used an adaptive thresholding approach followed by a closing morphological operation with an elliptical structuring element.
+
+	In the algorithm proposed by Romero-Rondon et al.
+in \cite{Romero2016} the detection of overlapped RBCs is based on marker-controlled watershed transform.
+To define the suitable markers in watershed transform they used three different approaches, based on a morphological erosion operation, on Hough transform and on clustering method of k-means.
+
 	Tek et al.
@@
 Furthermore, top-hat extraction and infinite reconstruction were applied to find the regions that include the objects.
+
+	Sovkare in \cite{Savkare2015} segmented cells using k-mean clustering and global threshold.
+Overlapping cells are separated using Sobel edge detector and watershed transform.
+Watershed transform is applied on each cluster separately.
+Over-segmentation is minimized by series of morphological operations, like erosion and dilation utilizing disk shaped structuring element.
+
+	In \cite{Dey2015} the authors proposed an automatic system for segmenting platelets, useful for identifying disease as malaria, using a color based segmentation and mathematical morphology (opening operations with a disk element of radius 2).
+
+	In \cite{Sulist2015} morphological operations (erosion, dilation, opening and closing) and blob analysis have been combined to segment and identify malaria parasites with a high degree of accuracy.
+
+	In \cite{Savkare2011_1} an approach  to detect red blood cells with consecutive classification into parasite infected and normal cells for further estimation of parasitemia is  proposed.
+For separation of overlapping cells watershed transform is applied on distance transform of binary mask of cells having larger area.
+
+	In \cite{Tek2010} the localisation of the parasites is achieved after a foreground and background segmentation step.
+Firstly, a rough foreground image using morphological area top-hats (using the average cell area value) is extracted.
+Then, from these rough foreground and background regions two different threshold values are determined and used in morphological double thresholding of the input grey level image to produce a refined binary foreground mask.
+From the foreground image the stained pixels are detected using a thresholding approach and finally used as markers to extract the stained objects by morphological area top-hats based on the estimated average area value.
+
+	In the study presented in \cite{Diaz2009} for quantification and classification of erythrocytes in stained thin blood films infected with Plasmodium falciparum, the authors used connected morphological operators in the segmentation step.
+The RBCs are detected as follows: firstly, a pixel classification allowed to label each image pixel as either background or foreground, based on its color features.
+Afterward, an inclusion-tree structure represented the hierarchical object relations between background and foreground so that a filtering process
+	allows to remove irrelevant structures such as artifacts generated at the staining or digitization processes.
+
+	Many systems for computer aided diagnosis of malaria disease made use of mathematical operatations in order to smoothen the boundary of the regions obtained from the segmentation process, like in \cite{Dave2017}.
+
 	%
@@
 Some authors used area granulometry for pre-processing purposes in malaria characterization \cite{Tek2010} even though it is certainly effective for extracting cells size features information \cite{Tek2006},\cite{Malihi2013},\cite{Springl2009}.
+In \cite{Tek2010} local area granulometry combined with colour histogram are used as features.
+The area granulometry feature is calculated locally on the binary mask of the stained objects, for the RGB channels and then concatenated.
+Morphological features are also used in \cite{Ross2006} (erosion dilation) and in \cite{Das2011} (opening, closing).
+
 	%
@@
 	%
-%\section{Discussions}
-%A comprehensive synthesis of the analysed works is presented in table \ref{tab:summary_table}.
+	\section{Discussions}
+	A comprehensive synthesis of the analysed works is presented in table \ref{tab:summary_table}.
 It only considers the methods which employed mathematical morphology in at least one step of the pipelines and it has been structured by considering the following information: staining technique, segmentation methods, extracted features, classification strategy, specificity and sensitivity percentages relative to the classification steps.
-%Most of the studies are based on P. vivax and P. falciparum characterization.
-%With regards to the showed approaches and related results, it is clear that malaria parasites detection and segmentation techniques in microscopic images needs further experiments and improvements.
+	Most of the studies are based on P. vivax and P. falciparum characterization.
+	With regards to the showed approaches and related results, it is clear that malaria parasites detection and segmentation techniques in microscopic images needs further experiments and improvements.
 In general, the analysed works have been tested with a limited number of images and the datasets are not publicly available; therefore, a comparison between different approaches is very difficult.
@@
 However, one should take into account that 80\% of malaria cases occur in Africa, where this type of equipment is scarce or even nonexistent in common healthcare facilities \cite{Rosado2016}.
-%Moreover, this review showed that P. falciparum is the most analysed if we refer to segmentation and detection, considering that it is the most widespread among malaria parasite types.
-%The majority of the works used thin blood smear. It is typically used for identification of malaria infected
-%stages, types of parasitic infection and percentage of parasitemia, while thick blood smear is used for identification of and quantification of malaria parasite count against leukocyte count per microliter blood.
-%Pre-processing phase is typically taken on with filters and the most used in the analysed works is certainly the median filter which permits to preserve sharp edges.
+	Moreover, this review showed that P. falciparum is the most analysed if we refer to segmentation and detection, considering that it is the most widespread among malaria parasite types.
+	The majority of the works used thin blood smear. It is typically used for identification of malaria infected
+	stages, types of parasitic infection and percentage of parasitemia, while thick blood smear is used for identification of and quantification of malaria parasite count against leukocyte count per microliter blood.
+	Pre-processing phase is typically taken on with filters and the most used in the analysed works is certainly the median filter which permits to preserve sharp edges.
 Apart from the classic histogram equalization and contrast stretching techniques, other filters have been employed, e.g., geometric mean filter to remove Gaussian noise preserving edges, Laplacian filter, in order to find edges, and so on.
-%Median filter has been found to be effective for reducing impulse noises from the microscopic images, even though recent studies have shown that geometric mean filter provides better performance than the median filter \cite{Das2013}, \cite{Das2015}.
-%
-%Malaria parasites may be discriminated according to two different strategies: by segmenting the whole erythrocyte from the blood smear image on the basis of which malaria infection is detected, otherwise by segmenting chromatin dot or parasite infection region for characterizing parasite infection stages based on some extracted target features.
-%In general, thresholding-based approach is still widely used for segmentation purposes.
+	Median filter has been found to be effective for reducing impulse noises from the microscopic images, even though recent studies have shown that geometric mean filter provides better performance than the median filter \cite{Das2013}, \cite{Das2015}.
+
+	Malaria parasites may be discriminated according to two different strategies: by segmenting the whole erythrocyte from the blood smear image on the basis of which malaria infection is detected, otherwise by segmenting chromatin dot or parasite infection region for characterizing parasite infection stages based on some extracted target features.
+	In general, thresholding-based approach is still widely used for segmentation purposes.
 In particular, a lot of authors affirm that Otsu thresholding suffers from limitations when textural variation is high, while histogram thresholding can not deal sufficiently good in identifying valley regions in case of unimodal histograms.
@@
 Other authors \cite{Ahirwar2012}, \cite{DiRuberto2002}, \cite{Soni2011}, employed granulometry and stated that it is very effective to segment cells with regular size.
-%The analysed works performed classification phase for different purposes.
+	The analysed works performed classification phase for different purposes.
 The majority of them aimed to distinguish among two classes \cite{Anggraini2011}, \cite{Annaldas2014},  \cite{Das2015}, \cite{Fang2011}, \cite{Ghosh2011}, \cite{Kareem2012}, \cite{Mushabe2013}: malaria infected and noninfected RBCs, whereas more complex classification strategies aimed to classify from three \cite{Das2012}, \cite{Maity2012} to up to seven different classes \cite{Das2013}, \cite{Das2015}, \cite{Makkapati2011}, \cite{Malihi2013}.
-%
-%\section{Conclusions}
-%This work reviewed several computational microscopic imaging techniques oriented to mathematical morphology approach, proposed in literature for malaria parasites detection and segmentation in blood smear microscopic images.
-%
-%The computer vision methodologies reported in the literature are based on light microscopic images of human peripheral blood smears for computer-aided detection of malaria parasite and its various stages.
+
+	\section{Conclusions}
+	This work reviewed several computational microscopic imaging techniques oriented to mathematical morphology approach, proposed in literature for malaria parasites detection and segmentation in blood smear microscopic images.
+
+	The computer vision methodologies reported in the literature are based on light microscopic images of human peripheral blood smears for computer-aided detection of malaria parasite and its various stages.
 Image pre-processing, segmentation of erythrocytes and parasites, malaria parasite feature extraction, malaria detection techniques have been discussed here.
-%
-%It is worth to remember that mathematical morphology techniques have been widely used for image processing purposes.
+
+	It is worth to remember that mathematical morphology techniques have been widely used for image processing purposes.
 Among the application fields, it has been applied for fingerprint feature extraction, recognition of handwritten digits, license plate detection, border extraction, denoising using morphological filters, text extraction and detection of imperfection in printed circuit boards \cite{Kaur2013}.
 Apart from this kind of fields, mathematical morphology has been employed in biomedical image analysis, especially in pre-processing and segmentation techniques.
-%
-%Morphological cell analysis is used to face off abnormality identification and classification, early cancer detection. It has been integrated in new methods for biomedical applications,
-%such as automatic segmentation and analysis of histological tumour sections, boundary detection of cervical cell nuclei considering overlapping and clustering, the granules segmentation and spatial distribution analysis, morphological characteristics analysis of specific biomedical cells, understanding the chemotactic response and drug influences, or identifying cell morphogenesis in different cell cycle progression.
-%Morphological feature quantification for grading cancerous or precancerous cells is especially widely researched in the literature, such as nuclei segmentation based on marker-controlled watershed transform and snake model for hepatocellular carcinoma feature extraction and classification, which is important for prognosis and treatment planning, nuclei feature quantification for cancer cell cycle analysis, and using feature extraction including image morphological analysis, wavelet analysis, and texture analysis for automated classification of renal cell \cite{Chen2012}.
-%
-%Moreover, non-linear filtering has become increasingly important in many image processing applications.
+
+	Morphological cell analysis is used to face off abnormality identification and classification, early cancer detection. It has been integrated in new methods for biomedical applications,
+	such as automatic segmentation and analysis of histological tumour sections, boundary detection of cervical cell nuclei considering overlapping and clustering, the granules segmentation and spatial distribution analysis, morphological characteristics analysis of specific biomedical cells, understanding the chemotactic response and drug influences, or identifying cell morphogenesis in different cell cycle progression.
+	Morphological feature quantification for grading cancerous or precancerous cells is especially widely researched in the literature, such as nuclei segmentation based on marker-controlled watershed transform and snake model for hepatocellular carcinoma feature extraction and classification, which is important for prognosis and treatment planning, nuclei feature quantification for cancer cell cycle analysis, and using feature extraction including image morphological analysis, wavelet analysis, and texture analysis for automated classification of renal cell \cite{Chen2012}.
+
+	Moreover, non-linear filtering has become increasingly important in many image processing applications.
 Initially, the attraction to non-linear filters was mostly limited to the impulse-removing and edge-preserving qualities of the median filter.
@@
 The shape-based methods of mathematical morphology, in particular, are now used in a wide variety of medical applications, including electrocardiography, ultrasound imaging, radiology, and histologic image analysis \cite{Schulze1994}.
-%
-%Furthermore, microscopic image analysis and, in particular, malaria detection and classification can greatly benefit from the use of mathematical morphology.
+
+	Furthermore, microscopic image analysis and, in particular, malaria detection and classification can greatly benefit from the use of mathematical morphology.
 The interest in this approach to image processing ad analysis is proved by the increasing number of works proposing methods for malaria image analysis based on mathematical morphology techniques.
-%
-%It is worth to mention that the development of new mobility-aware microscopic devices (and ideally low cost) is an area that can greatly improve the chances of the successful deployment of computer vision CAD solutions for malaria diagnosis in the field.
-%The mobile phone is currently Africa's most important digital technology, and is boosting African health as it emerges as a platform for diagnosis and treatment.
-%Considering the recent significant improvements of the new generation of mobile devices in terms of image acquisition and processing power, if a reliable automatic diagnostic performance is ensured through the usage of those devices, one would dramatically reduce the effort in the exhaustive and time consuming activity of microscopic examination.
-%Moreover, the lack of highly trained microscopists on malaria diagnosis in rural areas could then be complemented by a significantly less specialized technician that knows how to operate the system and prepare blood smears.
+
+	It is worth to mention that the development of new mobility-aware microscopic devices (and ideally low cost) is an area that can greatly improve the chances of the successful deployment of computer vision CAD solutions for malaria diagnosis in the field.
+	The mobile phone is currently Africa's most important digital technology, and is boosting African health as it emerges as a platform for diagnosis and treatment.
+	Considering the recent significant improvements of the new generation of mobile devices in terms of image acquisition and processing power, if a reliable automatic diagnostic performance is ensured through the usage of those devices, one would dramatically reduce the effort in the exhaustive and time consuming activity of microscopic examination.
+	Moreover, the lack of highly trained microscopists on malaria diagnosis in rural areas could then be complemented by a significantly less specialized technician that knows how to operate the system and prepare blood smears.
 The usage of mobile devices in the system architecture can also bring significant improvements in terms of portability and data transmission.
 Finally, malaria diagnosis might be just one
-%element of a suite of diagnostic software tests running on this type of system.
+	element of a suite of diagnostic software tests running on this type of system.
 Several other tests could simultaneously be carried out using the same images, for instance cell counting or detection of other hemoparasites like microfilaria or trypanosoma \cite{Rosado2016}.

### 2017-11-16 Paper.tex
Review Andre Discussions e Conclusions

@@
 
+	\section{Scope of this review}
 	In this paper we present a review of computer-aided methods oriented to malaria parasites detection and segmentation by mathematical morphology based techniques.
+	Most of the studies were followed Di Ruberto's work \cite{DiRuberto2002}, which first proposed a system to evaluate parasitaemia in the blood.
+The system was able to detect the parasites by using an automatic thresholding and morphological operators.
+A morphological approach to cell segmentation which is more efficient than watershed algorithm was proposed.
+Finally, the parasites classification was still based on morphological operators.
+Since then many systems for computer aided diagnosis of malaria have been proposed.
+Most of them make use of mathematical morphology to process and analyse malaria-infected peripheral blood cells images.
+The scope of this paper is to review and analyze the recent works of different researchers in the area of malaria parasite recognition using computer vision which benefit from mathematical morphology.
+
 	The rest of the paper is organised as follows.
-Section 2 introduces the adopted materials and methods for malaria-infected peripheral blood cells images analysis.
+Section 2 presents a review of the considered works, according to a typical pipeline of a computer-aided image analysis process: preprocessing, segmentation, feature extraction.
+	%introduces the adopted materials and methods for malaria-infected peripheral blood cells images analysis.
 It has been structured in preprocessing, segmentation, feature extraction, feature selection and classification subsections, consistently with a typical pipeline of a computer-aided image analysis process.
@@
 
-	\section{Materials and methods}
+	\section{Computer aided diagnosis of malaria by using mathematical morphology}
 	This section presents a review of some of the main recent studies existing in literature regarding the analysis of malaria infected blood smears using mathematical morphology.
@@
 \ref{fig:thick_thin_film}.
-	Giemsa stained blood smear is considered in most of the analysed literatures whereas Leishman stain is considered in few studies, as showed in table \ref{tab:summary_table}.
+	Giemsa stained blood smear is considered in most of the analysed literatures whereas Leishman stain is considered in few studies. %, as showed in table \ref{tab:summary_table}.
 
-	It is reported that Leishman stain has bigger sensitivity for parasite detection than Giemsa \cite{Khan2014} and is superior for visualization of red and white blood cell morphology.
-\cite{Sathpathi2014}.
+	It is reported that Leishman stain has bigger sensitivity for parasite detection than Giemsa \cite{Khan2014} and is superior for visualization of red and white blood cell morphology \cite{Sathpathi2014}.
 On the contrary, Giemsa stain highlights both malaria parasites and white blood cells and, therefore, it is an additional issue to deal with.
@@
 Morphological operators are useful for removal of unwanted
-	objects, holes filling, splitting, thinning and thinking.\texttt{Preprocessing techniques used by different researchers during automated diagnosis of malaria are listed below.}
+	objects, holes filling, splitting, thinning and thinking.
+Different researchers during automated diagnosis of malaria used morphological operations in preprocessing phase and the most recent are listed below.
 
@@
 The algorithm uses the integral image and haar-like features concepts, and weak classifiers with adaptive boosting learning.
-The search scope of the learning algorithm is reduced in the preprocessing step by removing the background around blood cells.
+The search scope of the learning algorithm is reduced in the preprocessing step by removing the background around blood cells by means of morphological erosions both for training and for testing.
 	%Morphological operations have been applied to images to enhance the geometric structures on a set of pixels.
@@
 During this step, some noise, the WBC cytoplasm and platelets still remain on the image.
-Therefore, the small objects are removed using a morphological opening and then the image is dilated with a disc-shaped structuring element.
+Therefore, the small objects are removed using a morphological opening and then the image is dilated with a disk-shaped structuring element.
 
@@
 A fuzzy C-means clustering is applied to extract infected erythrocytes, which is further processed for the final segmentation.
-The experimental results showed that the proposed method can gain 98\%, 93.3\%, 98.65\% and 90.33\% of sensitivity, specificity, prediction value positive and prediction value negative, respectively.
 A morphological erosion is used to erase some small noises and spots before the segmentation and holes inside the infected erythrocytes are filled using a morphological hole filling operation for the final segmentation.
+
 	The work \cite{Reni2015} proposes a new algorithm for morphological filtering of the blood images as a pre-processing tool for segmentation.
@@
 On the opposite the proposed method preserves the necessary information of foreground components while removing noise and artefacts.
+
 	The paper in \cite{Kareem2012_1} illustrates a morphological approach for blood cell identification and uses the image features such as intensity, histogram, relative size and geometry for further analysis.
@@
 The radius of the structuring element depends on the radius of the RBC, so that all the components smaller than the RBCs are removed.
+
 	In \cite{Tek2010} a complete framework to detect and identify malaria parasites in images of Giemsa stained thin blood film specimens is proposed.
@@
 The estimation is based on a morphological closing operation using a sufficiently large structuring element.
-The "sufficiently" large size for an input image is determined automatically with respect to its average cell size computed from the area granulometry distribution.
-	In the method proposed in \cite{Sheik2013} the first phase is the stain object extraction which extracts candidates objects that can be infected by malaria parasites by using intensity and colour.
-Before extracting the stained objects the method firstly extracts the foreground and enhanced images.
+The sufficiently large size for an input image is determined automatically with respect to its average cell size computed from the area granulometry distribution.
+
+	In the method proposed in \cite{Sheik2013} the first phase is the stained object extraction which detects candidates objects that can be infected by malaria parasites using intensity and colour.
+Before detecting the stained objects the method firstly extracts the foreground.
 Foreground image is a binary image which is produced after applying morphological hole filling on such pixels which have lower intensity value than average intensity value of green layer.
 After the stained objects extraction process, a series of morphological operations is also employed after every part in order to eliminate small components and complete the final stained objects.
+
 	In \cite{Gonzalez2016} a system to determine markers for watershed segmentation based on the Radon transform and mathematical operators is proposed.
 In the first step of the process small irrelevant structures and part of the noise are eliminated by a morphological filter, in order to ensure the preservation of the cells edges.
-Image smoothing is performed by a morphological erosion-reconstruction and dilation-reconstruction (ERDR) filter with a disc structuring element of radius equal to 20 pixels, which is $0.274$ times smaller than the average radius of the RBCs.
+Image smoothing is performed by a morphological erosion-reconstruction and dilation-reconstruction filter with a disk structuring element of radius equal to 20 pixels, which is $0.274$ times smaller than the average radius of the RBCs.
 In this way the influences of the size and the shape of the structures can be separated in the smoothing process.
@@
 
-	Several other authors attempted to use thresholding as segmentation method in their computer-aided systems.
+	Several other authors attempted to use thresholding combined with morphological operation as segmentation method in their computer-aided systems.
+
 	In \cite{Ross2006} the authors propose a method which provides a positive or negative diagnosis of malaria and differentiates parasites by species.
@@
 	%dots, so a further evaluation is performed to establish if a chromatin dot is valid by identifying whether it is inside a RBC or not.
-	%
+
 	The authors in \cite{Kareem2011} used the Annular Ring Ratio (ARR) transform method.
@@
 This approach results in locating only the stained components in the image instead of all the cells and hence will not only speed up the operation but reduces the complexity.
-	%
+
 	%The morphology of P.falciparum at gametocyte stage is the key parameter in the segmentation approach proposed in \cite{Kareem2012}.
@@
 The authors concluded that B and G channels from the RGB colour space and I channel from the HSI colour space are the most sensitive channels for MP detection.
-	%
+
 	Anggraini et al.
@@
 From each identified cell (RBC and WBC), constituents of the parasite (nucleus and cytoplasm) were extracted using multiple threshold.
-	%
+
 	Savkare et al.
@@
 Then, watershed transform has been applied for separating overlapped cells.
-	%
+
 	In \cite{Ghosh2011}, authors used divergence based threshold selection in order to segment P.vivax parasites from Leishman-stained thin blood films.
-This method is based on Cauchy membership function \cite{Put14a} and is applied to the C channel of CMYK colour space.
+This method is based on Cauchy membership function \cite{DiRuberto2014} and is applied to the C channel of CMYK colour space.
 Morphological operators of opening and closing have been used for artefacts removal.
-	%
+
 	Tek et al.
@@
 They used both binary image masks obtained with Otsu method and the edge detection images from Canny method for RBC segmentation, followed by a hole filling process.
-	%
-	Authors in \cite{Arco2015} worked on thick blood films and proposed a method that uses an adaptive thresholding based scheme, which also allows an effective classification of pixels.
+
+	The authors in \cite{Arco2015} worked on thick blood films and proposed a method that uses an adaptive thresholding based scheme, which also allows an effective classification of pixels.
 This means that the election of whether a pixel belongs to the background or to the signal (parasites and white blood cells) is only established by the pixels around it, that is its neighbourhood.
@@
 They worked on Giemsa-stained thin blood smears.
-	%
+
 	The proposed automated method in \cite{Elter2011} for parasite detection and identification worked on thin blood film acquired with Giemsa stain.
@@
 However, the authors do not define the value of this fixed threshold on the publication.
-	%
+
 	%% Morphology and/or granulometry in Segmentation stage:
@@
 These methods make use of the red blood cell structure knowledge, that is not used in existing watershed-based algorithms.
-	%
+
 	Soni et al.
@@
 Any objects that are connected to the border of the image can be removed.
-	%
+
 	Also Airwhar et al.
@@
 The contours are filled, and the segmented mask is again reconstructed with the valid parasite marker image to result in a segmented mask of infected cells.
-	%
-	Several authors used marker controlled watershed with morphological approach.
-	Khan et al.
-\cite{Khan2011}, among other experimentations, used it in order to try to separate overlapping cells because, according to their statements, watershed transform can separate touching cells but it is not sufficient for overlapping cells.
-	In \cite{Springl2009} red blood cell segmentation has been performed by using marker-controlled watershed transformation based on the image gradient.
-Markers are computed as a combination of the binary mask of the red blood cells and centres of the cells which are computed using a similar algorithm that was utilized for the evaluation of the average cell radius.
-The binary mask is obtained by thresholding the grey-scale image with an automatically estimated threshold using Otsu method\cite{Otsu1975}.
-	Das et al.
-\cite{Das2011}, \cite{Das2013}, \cite{Das2012}, \cite{Das2015} segmented erythrocytes as aforesaid and then morphological operators are used to eliminate unwanted cells like leukocytes and platelets.
-To conclude, overlapping erythrocytes are segmented by using marker controlled watershed segmentation technique.
 
-	In the paper \cite{Devi2017} the authors propose a computer assisted system for quantification of erythrocytes in microscopic images of thin blood smears.
-The proposed method consists of preprocessing, segmentation, morphological filtering, cell separation and clump cell segmentation.
-The performance of the system in classifying the isolated and clump erythrocytes by geometric features is evaluated for the different classifiers.
-Moreover, the clump erythrocytes are segmented using marker controlled watershed with h-minima as internal marker.
-Based on the experimental results, it may be concluded that the proposed model provides satisfactory results with an accuracy of 98.02\% in comparison to the state of art.
-
-	%
-	%%BOUNDARY BASED METHODS \subsubsection{Boundary based methods}
-	%Zou et al.
-\cite{Zou2010} based their method on the consideration that RBCs have a circular shape.
-They propose an improved version of the Circle Hough Transform (CHT) for RBCs detection.
-The authors state that overlapping and oval cells can be detected as well as non-clumped RBCs and affirm that the green channel of the RGB colour space permits to better distinguish RBCs.
-However, some RBCs may not have circular shape and it could affect the detection results.
-	%
 	In \cite{Yunda2012}, Yunda et al.
@@
 
-	In the algorithm proposed by Romero-Rondon et al.
+	In \cite{Tek2010} the localisation of the parasites is achieved after a foreground and background segmentation step.
+Firstly, a rough foreground image using morphological area top-hats (using the average cell area value) is extracted.
+Then, from these rough foreground and background regions two different threshold values are determined and used in morphological double thresholding of the input grey level image to produce a refined binary foreground mask.
+From the foreground image the stained pixels are detected using again a thresholding approach and finally used as markers to extract the stained objects by morphological area top-hats based on the estimated average area value.
+
+	Several authors used marker controlled watershed with morphological approach.
+	Khan et al.
+\cite{Khan2011}, among other experimentations, used it in order to try to separate overlapping cells because, according to their statements, watershed transform can separate touching cells but it is not sufficient for overlapping cells.
+
+	In \cite{Springl2009} red blood cell segmentation has been performed by using marker-controlled watershed transformation based on the image gradient.
+Markers are computed as a combination of the binary mask of the red blood cells and centres of the cells which are computed using a similar algorithm that was utilized for the evaluation of the average cell radius.
+The binary mask is obtained by thresholding the grey-scale image with an automatically estimated threshold using Otsu method\cite{Otsu1975}.
+
+	Das et al.
+\cite{Das2011}, \cite{Das2013}, \cite{Das2014}, \cite{Das2015} segmented erythrocytes as aforesaid and then morphological operators are used to eliminate unwanted cells like leukocytes and platelets.
+To conclude, overlapping erythrocytes are segmented by using marker controlled watershed segmentation technique.
+
+	In the paper \cite{Devi2017} the authors propose a computer assisted system for quantification of erythrocytes in microscopic images of thin blood smears.
+	%The proposed method consists of preprocessing, segmentation, morphological filtering, cell separation and clump cell segmentation.
+	The performance of the system in classifying the isolated and clump erythrocytes by geometric features is evaluated for the different classifiers.
+The clump erythrocytes are segmented using marker controlled watershed with h-minima as internal marker.
+	%Based on the experimental results, it may be concluded that the proposed model provides satisfactory results with an accuracy of 98.02\% in comparison to the state of art.
+
+	%
+	%%BOUNDARY BASED METHODS \subsubsection{Boundary based methods}
+	%Zou et al.
+\cite{Zou2010} based their method on the consideration that RBCs have a circular shape.
+They propose an improved version of the Circle Hough Transform (CHT) for RBCs detection.
+The authors state that overlapping and oval cells can be detected as well as non-clumped RBCs and affirm that the green channel of the RGB colour space permits to better distinguish RBCs.
+However, some RBCs may not have circular shape and it could affect the detection results.
+	%
+
+	In the algorithm described by Romero-Rondon et al.
 in \cite{Romero2016} the detection of overlapped RBCs is based on marker-controlled watershed transform.
 To define the suitable markers in watershed transform they used three different approaches, based on a morphological erosion operation, on Hough transform and on clustering method of k-means.
+
+	Sovkare in \cite{Savkare2015} segmented cells using k-mean clustering and global threshold.
+Overlapping cells are separated using Sobel edge detector and watershed transform.
+Watershed transform is applied on each cluster separately.
+Over-segmentation is minimized by series of morphological operations, like erosion and dilation utilizing disk shaped structuring element.
+
+	In \cite{Savkare2011_1} an approach  to detect red blood cells with consecutive classification into parasite infected and normal cells for further estimation of parasitemia is  proposed.
+For separation of overlapping cells watershed transform is applied on distance transform of binary mask of cells having larger area.
 
@@
 
-	Sovkare in \cite{Savkare2015} segmented cells using k-mean clustering and global threshold.
-Overlapping cells are separated using Sobel edge detector and watershed transform.
-Watershed transform is applied on each cluster separately.
-Over-segmentation is minimized by series of morphological operations, like erosion and dilation utilizing disk shaped structuring element.
-
 	In \cite{Dey2015} the authors proposed an automatic system for segmenting platelets, useful for identifying disease as malaria, using a color based segmentation and mathematical morphology (opening operations with a disk element of radius 2).
@@
 	In \cite{Sulist2015} morphological operations (erosion, dilation, opening and closing) and blob analysis have been combined to segment and identify malaria parasites with a high degree of accuracy.
-
-	In \cite{Savkare2011_1} an approach  to detect red blood cells with consecutive classification into parasite infected and normal cells for further estimation of parasitemia is  proposed.
-For separation of overlapping cells watershed transform is applied on distance transform of binary mask of cells having larger area.
-
-	In \cite{Tek2010} the localisation of the parasites is achieved after a foreground and background segmentation step.
-Firstly, a rough foreground image using morphological area top-hats (using the average cell area value) is extracted.
-Then, from these rough foreground and background regions two different threshold values are determined and used in morphological double thresholding of the input grey level image to produce a refined binary foreground mask.
-From the foreground image the stained pixels are detected using a thresholding approach and finally used as markers to extract the stained objects by morphological area top-hats based on the estimated average area value.
 
@@
 
-	Many systems for computer aided diagnosis of malaria disease made use of mathematical operatations in order to smoothen the boundary of the regions obtained from the segmentation process, like in \cite{Dave2017}.
+	To conlude many systems for computer aided diagnosis of malaria disease made use of mathematical operatations in order to smoothen the boundary of the regions obtained from the segmentation process, like in \cite{Dave2017}.
 
@@
 The area granulometry feature is calculated locally on the binary mask of the stained objects, for the RGB channels and then concatenated.
-Morphological features are also used in \cite{Ross2006} (erosion dilation) and in \cite{Das2011} (opening, closing).
-
+Morphological features are also used in \cite{Ross2006} (erosion dilation), in \cite{Das2011} (opening, closing) and in \cite{DiRuberto2002} (skeleton) to classify parasites.
 	%
@@
 	\section{Discussions}
-	A comprehensive synthesis of the analysed works is presented in table \ref{tab:summary_table}.
-It only considers the methods which employed mathematical morphology in at least one step of the pipelines and it has been structured by considering the following information: staining technique, segmentation methods, extracted features, classification strategy, specificity and sensitivity percentages relative to the classification steps.
+	%A comprehensive synthesis of the analysed works is presented in table \ref{tab:summary_table}.
+	In the review we have only considered the methods which employed mathematical morphology in at least one step of the pipelines and it has been structured by considering the following information: preprocessing, segmentation, features extraction.
+	%, classification strategy, specificity and sensitivity percentages relative to the classification steps.
 	Most of the studies are based on P. vivax and P. falciparum characterization.
@@
 Despite promising results reported during the past years, the great majority of the computer-aided methods found on the literature for malaria diagnosis are based on images acquired under well controlled conditions and with proper microscopic equipment.
-However, one should take into account that 80\% of malaria cases occur in Africa, where this type of equipment is scarce or even nonexistent in common healthcare facilities \cite{Rosado2016}.
+However, one should take into account that 80\% of malaria cases occur in Africa, where this type of equipment is scarce or even nonexistent in common healthcare facilities \cite{Rosado2017}.
 	Moreover, this review showed that P. falciparum is the most analysed if we refer to segmentation and detection, considering that it is the most widespread among malaria parasite types.
-	The majority of the works used thin blood smear. It is typically used for identification of malaria infected
-	stages, types of parasitic infection and percentage of parasitemia, while thick blood smear is used for identification of and quantification of malaria parasite count against leukocyte count per microliter blood.
+	The majority of the works used thin blood smear.
+It is typically used for identification of malaria infected stages, types of parasitic infection and percentage of parasitemia, while thick blood smear is used for identification and quantification of malaria parasite count against leukocyte count per microliter blood.
+
 	Pre-processing phase is typically taken on with filters and the most used in the analysed works is certainly the median filter which permits to preserve sharp edges.
@@
 	Median filter has been found to be effective for reducing impulse noises from the microscopic images, even though recent studies have shown that geometric mean filter provides better performance than the median filter \cite{Das2013}, \cite{Das2015}.
+However, morphological operators have been greatly used with successful performances, imposing themselves as powerful alternatives to more common and used techniques for image enhancement and noise filtering \cite{Oliveira2017}, \cite{Reni2015}, \cite{Romero2016},  \cite{Sheik2013}, \cite{Somasekar2015}, \cite{Tek2010}.
 
@@
 In particular, a lot of authors affirm that Otsu thresholding suffers from limitations when textural variation is high, while histogram thresholding can not deal sufficiently good in identifying valley regions in case of unimodal histograms.
+However, such a simple and fast approach can greatly benefit from mathematical morphology as recent studies demonstrate (\cite{Ahirwar2012}, \cite{Anggraini2011}, \cite{Arco2014}, \cite{DiRuberto2002}, \cite{Elter2011}, \cite{Ghosh2011},  \cite{Kareem2011}, \cite{Mushabe2013},  \cite{Rosado2017}, \cite{Ross2006}, \cite{Savkare2011}, \cite{Savkare2015}, \cite{Somasekar2017}, \cite{Tek2010}).
+
 	Another greatly used segmentation approach is clearly the watershed transform.
-The classic watershed approach is reported to produce over segmentation results \cite{Savkare2011}, whereas the marker controlled approach does not suffer from this issue and it is reported to be very effective for overlapping cells segmentation even though some authors affirm that it may fail to segment highly overlapped cells \cite{Khan2011}, \cite{Springl2009}.
-Other authors \cite{Ahirwar2012}, \cite{DiRuberto2002}, \cite{Soni2011}, employed granulometry and stated that it is very effective to segment cells with regular size.
+The classic watershed approach is reported to produce over segmentation results \cite{Savkare2011}, whereas the marker controlled approach does not suffer from this issue and it is reported to be very effective for overlapping cells segmentation even though some authors affirm that it may fail to segment highly overlapped cells  (\cite{Das2011}, \cite{Das2013}, \cite{Das2014}, \cite{Das2015}, \cite{Devi2017}, \cite{Khan2011}, \cite{Romero2016}, \cite{Savkare2015}, \cite{Savkare2011_1}, \cite{Springl2009}).
+
+	Other authors (\cite{Ahirwar2012}, \cite{DiRuberto2002}, \cite{Soni2011}, \cite{Tek2010}) employed granulometry and stated that it is very effective to segment cells with regular size.
+
 	The analysed works performed classification phase for different purposes.
-The majority of them aimed to distinguish among two classes \cite{Anggraini2011}, \cite{Annaldas2014},  \cite{Das2015}, \cite{Fang2011}, \cite{Ghosh2011}, \cite{Kareem2012}, \cite{Mushabe2013}: malaria infected and noninfected RBCs, whereas more complex classification strategies aimed to classify from three \cite{Das2012}, \cite{Maity2012} to up to seven different classes \cite{Das2013}, \cite{Das2015}, \cite{Makkapati2011}, \cite{Malihi2013}.
+The majority of them aimed to distinguish among two classes only, malaria infected and noninfected RBCs, or to detect and count parasites in a malaria blood image (\cite{Anggraini2011}, \cite{Arco2014}, \cite{Das2011}, \cite{Dave2017}, \cite{Das2015}, \cite{DiRuberto2002}, \cite{Elter2011}, \cite{Ghosh2011}, \cite{Kareem2012}, \cite{Khan2014}, \cite{Malihi2013}, \cite{Mushabe2013}, \cite{Oliveira2017}, \cite{Savkare2011}, \cite{Savkare2011_1}, \cite{Somasekar2011}, \cite{Somasekar2015}, \cite{Somasekar2017}, \cite{Soni2011}, \cite{Sulist2015}, \cite{Tek2006}).
+
+	More complex classification strategies aimed to classify parasites into different classes, i.e.
+different human parasites species (\cite{Ahirwar2012}, \cite{Das2013}, \cite{Das2014}, \cite{Khan2011}, \cite{Tek2010}),
+	and/or different parasites life stages (\cite{Anggraini2011}, \cite{Das2013}, \cite{Das2014}, \cite{DiRuberto2002}, \cite{Diaz2009}, \cite{Tek2010}).
 
@@
 	The computer vision methodologies reported in the literature are based on light microscopic images of human peripheral blood smears for computer-aided detection of malaria parasite and its various stages.
-Image pre-processing, segmentation of erythrocytes and parasites, malaria parasite feature extraction, malaria detection techniques have been discussed here.
+Image preprocessing, segmentation of erythrocytes and parasites, malaria parasite feature extraction, malaria detection techniques have been discussed here.
 
@@
 Among the application fields, it has been applied for fingerprint feature extraction, recognition of handwritten digits, license plate detection, border extraction, denoising using morphological filters, text extraction and detection of imperfection in printed circuit boards \cite{Kaur2013}.
-Apart from this kind of fields, mathematical morphology has been employed in biomedical image analysis, especially in pre-processing and segmentation techniques.
+Apart from this kind of fields, mathematical morphology has been employed in biomedical image analysis, especially in preprocessing and segmentation techniques.
 

### 2017-11-16 v1/Paper.tex
Review Andre Discussions e Conclusions

--- 2017-11-09 Paper.tex
@@
 \begin{document}
-	\title{Recent advances of Malaria parasites detection systems based on mathematical morphology techniques}
+\title{Malaria parasites detection using mathematical morphology techniques: a review}
 
@@
 	\centering
-		\includegraphics[width=2.75cm]{img/f3_Pfalciparum}
-		\includegraphics[width=2.75cm]{img/f3_Pvivax}
-		\includegraphics[width=2.75cm]{img/f3_Povale}
-		\includegraphics[width=2.75cm]{img/f3_Pmalariae}
+	\includegraphics[width=2.75cm]{f3_Pfalciparum}
+	\includegraphics[width=2.75cm]{f3_Pvivax}
+	\includegraphics[width=2.75cm]{f3_Povale}
+	\includegraphics[width=2.75cm]{f3_Pmalariae}
 	\caption{\label{fig:malaria_types}Types of human malaria: from left to right, P. Falciparum, P. Vivax, P. Ovale, P. Malariae \cite{Khan2011}.}
@@
 	\centering
-		\includegraphics[width=2.75cm, height=2.75cm]{img/1_Pfalciparum_ring}
-		\includegraphics[width=2.75cm, height=2.75cm]{img/2_Pfalciparum_trophozoite}
-		\includegraphics[width=2.75cm, height=2.75cm]{img/3_Pfalciparum_gametocyte}
-		\includegraphics[width=2.75cm, height=2.75cm]{img/4_Pfalciparum_schizont}
-		\includegraphics[width=2.75cm, height=2.75cm]{img/5_Pmalariae_ring}
-		\includegraphics[width=2.75cm, height=2.75cm]{img/6_Pmalariae_schizont}
-		\includegraphics[width=2.75cm, height=2.75cm]{img/7_Povale_trophozoites}
-		\includegraphics[width=2.75cm, height=2.75cm]{img/8_Pvivax_trophozoites}
-		\includegraphics[width=2.75cm, height=2.75cm]{img/9_Pvivax_ring}
-		\includegraphics[width=2.75cm, height=2.75cm]{img/10_Pvivax_gametocyte}
-		\includegraphics[width=2.75cm, height=2.75cm]{img/11_Pvivax_ring}
+	\includegraphics[width=2.75cm, height=2.75cm]{1_Pfalciparum_ring}
+	\includegraphics[width=2.75cm, height=2.75cm]{2_Pfalciparum_trophozoite}
+	\includegraphics[width=2.75cm, height=2.75cm]{3_Pfalciparum_gametocyte}
+	\includegraphics[width=2.75cm, height=2.75cm]{4_Pfalciparum_schizont}
+	\includegraphics[width=2.75cm, height=2.75cm]{5_Pmalariae_ring}
+	\includegraphics[width=2.75cm, height=2.75cm]{6_Pmalariae_schizont}
+	\includegraphics[width=2.75cm, height=2.75cm]{7_Povale_trophozoites}
+	\includegraphics[width=2.75cm, height=2.75cm]{8_Pvivax_trophozoites}
+	\includegraphics[width=2.75cm, height=2.75cm]{9_Pvivax_ring}
+	\includegraphics[width=2.75cm, height=2.75cm]{10_Pvivax_gametocyte}
+	\includegraphics[width=2.75cm, height=2.75cm]{11_Pvivax_ring}
 	\caption{\label{fig:malaria_stages}Examples of malaria parasite stages.
@@
 \section{Materials and methods}
-	This section presents a review of some of the main recent studies existing in literature regarding the analysis of malaria infected blood smears using mathematical morphology.
-	A typical approach usually comprises five different image processing and analysis tasks, as follows:
+This section critically reviews some of the main studies existing in literature regarding the analysis of malaria infected blood smears using mathematical morphology.
+Since typical approaches usually comprise five different image processing and analysis tasks, the reviewed works have been divided into the following sub-sections:
 \begin{enumerate}
@@
 \end{enumerate}
-	Since morphological techniques have been used in the first three phases, the reviewed works have been divided into the following sub-sections: pre-processing, segmentation and feature extraction.
 Each sub-section contains description about methods that cope with malaria parasites (MP) stained components analysis, both on thin and thick blood smears, without distinction.
@@
 %\subsection{Search of related literatures}
-	Extensive search of articles has been made in PubMed and Google Scholar search engines based on the keywords: "malaria, mathematical morphology, automated malaria diagnosis" up to October  2017.
-The search includes papers published in English and titles and abstracts of potentially relevant studies were selected and presented from the most recent ones.
+Extensive search of articles has been made in PubMed and Google Scholar search engines based on the keywords: "malaria, mathematical morphology, automated malaria diagnosis" up to January 2017.
+The search includes papers published in English and titles and abstracts of potentially relevant studies were selected.
 Thereafter, the full texts of these studies were evaluated as per the exclusion criteria.
@@
 	\centering
-		\includegraphics[height=0.15\textheight]{img/f2_ThickGiemsa}
-		\includegraphics[height=0.15\textheight]{img/f2_ThickLeishman}
-		\includegraphics[height=0.15\textheight]{img/f2_ThinGiemsa}
-		\includegraphics[height=0.15\textheight]{img/f2_ThinLeishman}
+	\includegraphics[height=0.15\textheight]{f2_ThickGiemsa}
+	\includegraphics[height=0.15\textheight]{f2_ThickLeishman}
+	\includegraphics[height=0.15\textheight]{f2_ThinGiemsa}
+	\includegraphics[height=0.15\textheight]{f2_ThinLeishman}
 	\caption{\label{fig:thick_thin_film}Malaria infected blood smear images.
@@
 In image analysis field, especially when we refer to complex computer-aided pipelines, pre-processing methods are particularly used in order to improve the image data by suppressing unwanted noise or enhancing some image features for further processing.
-	%Typically, pre-processing methods are classified according to the following schema:
-	%\begin{enumerate}
-	%	\item pixel brightness transformations;
-	%	\item geometric transformations;
-	%	\item pre-processing methods that use a local neighbourhood of the processed pixel;
-	%	\item image restoration that requires knowledge about the entire image.
-	%\end{enumerate}
+Typically, pre-processing methods are classified according to the following schema:
+\begin{enumerate}
+	\item pixel brightness transformations;
+	\item geometric transformations;
+	\item pre-processing methods that use a local neighbourhood of the processed pixel;
+	\item image restoration that requires knowledge about the entire image.
+\end{enumerate}
 It is worth to mention pre-processing methods because they are an important step regarding image analysis field but, for what concerns the malaria-affected blood image analysis, in our review we particularly found methods which operate for illumination correction and noise filtering purposes.
 Generally speaking, digital microscopy images can be acquired in different lighting conditions, with several types of acquisition devices or from blood smears stained with various staining protocol and, consequently, the features of similar images could differ a lot.
-	Different techniques for illumination correction have been suggested to reduce such variation, e.g., a lot of authors work with grayscale-converted images as an illumination correction method %\cite{Tek2010},\cite{Das2015},\cite{Tek2006}.
+Different techniques for illumination correction have been suggested to reduce such variation, e.g., a lot of authors work with grayscale-converted images as an illumination correction method \cite{Tek2010},\cite{Das2015},\cite{Tek2006}.
 On the other hand, noise filtering aims to remove the noise introduced by mishandling the slides and/or the camera settings.
-	Morphological operators have been extensively used as preprocessing for image enhancement in major studies.
-	Erosion and dilation operations on raw smear images allow discarding undesired patterns and help in the selection of required cells or regions of interest.
-Morphological operators are useful for removal of unwanted
-	objects, holes filling, splitting, thinning and thinking.\texttt{Preprocessing techniques used by different researchers during automated diagnosis of malaria are listed below.}
-
-	The system proposed in \cite{Oliveira2017} is based on image processing, artificial intelligence techniques and an adapted face detection algorithm to identify Plasmodium parasites.
-The algorithm uses the integral image and haar-like features concepts, and weak classifiers with adaptive boosting learning.
-The search scope of the learning algorithm is reduced in the preprocessing step by removing the background around blood cells.
-	%Morphological operations have been applied to images to enhance the geometric structures on a set of pixels.
-In particular, an erosion operation has been applied after a Gaussian filtering in order to locate the RBCs and restrict the search space of the algorithm.
-	%The accuracy of the system was, on average, 91, meaning that for every 100 parasite-infected samples, 91 were identified correctly.
-The system has been developed for mobile phones and tablets and can represent a useful tool
-	%in health centers in remote communities without the need for specific expertise.
-
-	In \cite{Romero2016} an algorithm that uses morphological operations, the watershed method, the Hough transform and the clustering method of k-means to detect overlapped RBCs is presented.
-In the pre-processing stage white blood cells and platelets are removed before the segmentation task.
-During this step, some noise, the WBC cytoplasm and platelets still remain on the image.
-Therefore, the small objects are removed using a morphological opening and then the image is dilated with a disc-shaped structuring element.
-
-	An edge-based segmentation of erythrocytes infected with malaria parasites using microscopic images is proposed in \cite{Somasekar2015}.
-A fuzzy C-means clustering is applied to extract infected erythrocytes, which is further processed for the final segmentation.
-The experimental results showed that the proposed method can gain 98\%, 93.3\%, 98.65\% and 90.33\% of sensitivity, specificity, prediction value positive and prediction value negative, respectively.
-A morphological erosion is used to erase some small noises and spots before the segmentation and holes inside the infected erythrocytes are filled using a morphological hole filling operation for the final segmentation.
-	The work \cite{Reni2015} proposes a new algorithm for morphological filtering of the blood images as a pre-processing tool for segmentation.
-Conventional morphological closing on blood images removes the unwanted components but also useful information.
-On the opposite the proposed method preserves the necessary information of foreground components while removing noise and artefacts.
-	The paper in \cite{Kareem2012_1} illustrates a morphological approach for blood cell identification and uses the image features such as intensity, histogram, relative size and geometry for further analysis.
-Before the identification of blood cells, the authors  propose a novel morphological filtering based on the size of RBC for platelets and/or artifacts elimination.
-A dilation is performed by a concentric ring structuring element and erosion by disk shaped structuring element.
-The radius of the structuring element depends on the radius of the RBC, so that all the components smaller than the RBCs are removed.
-	In \cite{Tek2010} a complete framework to detect and identify malaria parasites in images of Giemsa stained thin blood film specimens is proposed.
-Also, the system is able to identify the infecting species and life-cycle stages.
-	The preprocessing step of the proposed method is applied to reduce the variations in the observed size, intensity, and colour of the cells and stained objects before the detection and classification steps.
-The aim is to correct the non-uniform illumination in the images.
-The estimation is based on a morphological closing operation using a sufficiently large structuring element.
-The "sufficiently" large size for an input image is determined automatically with respect to its average cell size computed from the area granulometry distribution.
-	In the method proposed in \cite{Sheik2013} the first phase is the stain object extraction which extracts candidates objects that can be infected by malaria parasites by using intensity and colour.
-Before extracting the stained objects the method firstly extracts the foreground and enhanced images.
-Foreground image is a binary image which is produced after applying morphological hole filling on such pixels which have lower intensity value than average intensity value of green layer.
-After the stained objects extraction process, a series of morphological operations is also employed after every part in order to eliminate small components and complete the final stained objects.
-	In \cite{Gonzalez2016} a system to determine markers for watershed segmentation based on the Radon transform and mathematical operators is proposed.
-In the first step of the process small irrelevant structures and part of the noise are eliminated by a morphological filter, in order to ensure the preservation of the cells edges.
-Image smoothing is performed by a morphological erosion-reconstruction and dilation-reconstruction (ERDR) filter with a disc structuring element of radius equal to 20 pixels, which is $0.274$ times smaller than the average radius of the RBCs.
-In this way the influences of the size and the shape of the structures can be separated in the smoothing process.
-At the same time the objects which are not eliminated remain unchanged.
-Also, a morphological closing is performed with a disc structuring element having radius smaller than half the average of the RBCs' radii, in order to connect the possible (more than one) markers that can appear on a single cell.
-
-	%Median filter is often used for reducing impulse noise.
+Median filter is often used for reducing impulse noise.
 Several studies have used it to enhance microscopic images of peripheral blood smear towards characterization of malaria followed by adaptive or local histogram equalization.
@@
 For this purpose they designed an illumination pattern that simulates the classic visual defects introduced by the digital microscope lenses, that
-	%is the vignetting effect.
+is the vignetting effect.
 Starting from the smallest radius, they applied different illumination patterns, created modifying the radius of the Gaussian curve, to the original images.
@@
 
-	\subsection{Segmentation of RBCs and parasites}
+\subsection{Segmentation}
 Segmentation is a key step in image analysis because it permits the identification and separation of the regions that compose an image, according to certain criteria of homogeneity and separation.
@@
 The latter case is commonly used in applications in which the cell class depends on the morphological characteristics of its components.
-	%The segmentation techniques can generally be divided into three main categories: Pixel Based, Edge Based and Region Based.
-	%
-	%%\subsection{Erythrocytes segmentation}
-	%%\subsubsection{Thresholding}
-	%%THRESHOLDING METHODS
+The segmentation techniques can generally be divided into three main categories: Pixel Based, Edge Based and Region Based.
 
+%\subsection{Erythrocytes segmentation}
+%\subsubsection{Thresholding}
+%THRESHOLDING METHODS
 Several other authors attempted to use thresholding as segmentation method in their computer-aided systems.
-	In \cite{Ross2006} the authors propose a method which provides a positive or negative diagnosis of malaria and differentiates parasites by species.
+In \cite{Ross2006}, the authors propose a method which provides a positive or negative diagnosis of malaria and differentiates parasites by species.
 The segmentation step relies on a thresholding strategy which aims to identify and segment potential parasites and erythrocytes from the image background after a six steps threshold selection.
 Mathematical morphology has been used for parasite size estimation, erythrocytes reconstruction and cells bigger than erythrocytes removal.
-	%
-	%Makkapati and Rao proposed two different methodologies.
+
+Makkapati and Rao proposed two different methodologies.
 The first one \cite{Makkapati2009} is based on HSV colour space with the purpose of performing a red blood cells (RBCs) segmentation and an identification of RBCs affected from MP with a chromatin dot identification by thresholding the Saturation channel.
@@
 However, several artefacts may also be segmented as chromatin
-	%dots, so a further evaluation is performed to establish if a chromatin dot is valid by identifying whether it is inside a RBC or not.
-	%
+dots, so a further evaluation is performed to establish if a chromatin dot is valid by identifying whether it is inside a RBC or not.
+
 The authors in \cite{Kareem2011} used the Annular Ring Ratio (ARR) transform method.
@@
 This approach results in locating only the stained components in the image instead of all the cells and hence will not only speed up the operation but reduces the complexity.
-	%
-	%The morphology of P.falciparum at gametocyte stage is the key parameter in the segmentation approach proposed in \cite{Kareem2012}.
+
+The morphology of P.falciparum at gametocyte stage is the key parameter in the segmentation approach proposed in \cite{Kareem2012}.
 It is based on a modification of the ARR transform.
@@
 It utilizes the information based on structure, colour and geometry of the cells and does not require any segmentation or non-illumination correction techniques that are commonly used for cell detection.
-	%
-	%Fang et al.
+
+Fang et al.
 \cite{Fang2011} used a threshold to distinguish the unhealthy cells from the healthy cells after a malaria detection phase performed with a phase spectrum methodology.
@@
 The authors concluded that B and G channels from the RGB colour space and I channel from the HSI colour space are the most sensitive channels for MP detection.
-	%
+
 Anggraini et al.
@@
 From each identified cell (RBC and WBC), constituents of the parasite (nucleus and cytoplasm) were extracted using multiple threshold.
-	%
+
 Savkare et al.
@@
 Then, watershed transform has been applied for separating overlapped cells.
-	%
+
 In \cite{Ghosh2011}, authors used divergence based threshold selection in order to segment P.vivax parasites from Leishman-stained thin blood films.
@@
 Morphological operators of opening and closing have been used for artefacts removal.
-	%
+
 Tek et al.
@@
 This work also addresses non-uniform illumination problem, removing it by using a pre-recorded illumination image or using a morphological closing operation.
-	%
-	%Malihi et al.
+
+Malihi et al.
 \cite{Malihi2013} proposed a method for malaria parasite detection in Giemsa-stained blood samples.
 They used both binary image masks obtained with Otsu method and the edge detection images from Canny method for RBC segmentation, followed by a hole filling process.
-	%
+
 Authors in \cite{Arco2015} worked on thick blood films and proposed a method that uses an adaptive thresholding based scheme, which also allows an effective classification of pixels.
@@
 \end{enumerate}
-	%
-	Also in the method proposed in \cite{Somasekar2017} the segmentation of the infected parasites is based on thresholding.
-The segmentation is achieved in two stages by maximizing between-class variance of an original image and consequently by an iterative threshold selection from a stage-one threshold image with suitable stopping criteria.
-The segmented results are post processed to improve the accuracy of the detection of malaria
-	parasites by morphological operators (erosion and closing).
 
@@
 They worked on Giemsa-stained thin blood smears.
-	%
+
 The proposed automated method in \cite{Elter2011} for parasite detection and identification worked on thin blood film acquired with Giemsa stain.
@@
 However, the authors do not define the value of this fixed threshold on the publication.
-	%
-	%% Morphology and/or granulometry in Segmentation stage:
+
+% Morphology and/or granulometry in Segmentation stage:
 Di Ruberto et al.
@@
 These methods make use of the red blood cell structure knowledge, that is not used in existing watershed-based algorithms.
-	%
+
 Soni et al.
@@
 Any objects that are connected to the border of the image can be removed.
-	%
+
 Also Airwhar et al.
@@
 The contours are filled, and the segmented mask is again reconstructed with the valid parasite marker image to result in a segmented mask of infected cells.
-	%
+
 Several authors used marker controlled watershed with morphological approach.
@@
 Das et al.
-\cite{Das2011}, \cite{Das2013}, \cite{Das2012}, \cite{Das2015} segmented erythrocytes as aforesaid and then morphological operators are used to eliminate unwanted cells like leukocytes and platelets.
+\cite{Das2013}, \cite{Das2012}, \cite{Das2011}, \cite{Das2015} segmented erythrocytes as aforesaid and then morphological operators are used to eliminate unwanted cells like leukocytes and platelets.
 To conclude, overlapping erythrocytes are segmented by using marker controlled watershed segmentation technique.
 
-	In the paper \cite{Devi2017} the authors propose a computer assisted system for quantification of erythrocytes in microscopic images of thin blood smears.
-The proposed method consists of preprocessing, segmentation, morphological filtering, cell separation and clump cell segmentation.
-The performance of the system in classifying the isolated and clump erythrocytes by geometric features is evaluated for the different classifiers.
-Moreover, the clump erythrocytes are segmented using marker controlled watershed with h-minima as internal marker.
-Based on the experimental results, it may be concluded that the proposed model provides satisfactory results with an accuracy of 98.02\% in comparison to the state of art.
-
-	%
-	%%BOUNDARY BASED METHODS \subsubsection{Boundary based methods}
-	%Zou et al.
+%BOUNDARY BASED METHODS \subsubsection{Boundary based methods}
+Zou et al.
 \cite{Zou2010} based their method on the consideration that RBCs have a circular shape.
@@
 However, some RBCs may not have circular shape and it could affect the detection results.
-	%
+
 In \cite{Yunda2012}, Yunda et al.
@@
 
-	In \cite{Rosado2017} Rosado et al.
-proposed a system using supervised classification to assess the presence of malaria parasites and determine the species and life cycle stage in Giemsa-stained thin blood smears.
-For the RBCs segmentation, they used an adaptive thresholding approach followed by a closing morphological operation with an elliptical structuring element.
+%CLUSTERING METHODS \subsection{Clustering methods}: no method uses mathematical morphology applied to clustering.
 
-	In the algorithm proposed by Romero-Rondon et al.
-in \cite{Romero2016} the detection of overlapped RBCs is based on marker-controlled watershed transform.
-To define the suitable markers in watershed transform they used three different approaches, based on a morphological erosion operation, on Hough transform and on clustering method of k-means.
-
+%CLASSIFIER-BASED METHODS \subsection{Classifier-based methods}
 Tek et al.
@@
 
-	Sovkare in \cite{Savkare2015} segmented cells using k-mean clustering and global threshold.
-Overlapping cells are separated using Sobel edge detector and watershed transform.
-Watershed transform is applied on each cluster separately.
-Over-segmentation is minimized by series of morphological operations, like erosion and dilation utilizing disk shaped structuring element.
-
-	In \cite{Dey2015} the authors proposed an automatic system for segmenting platelets, useful for identifying disease as malaria, using a color based segmentation and mathematical morphology (opening operations with a disk element of radius 2).
-
-	In \cite{Sulist2015} morphological operations (erosion, dilation, opening and closing) and blob analysis have been combined to segment and identify malaria parasites with a high degree of accuracy.
-
-	In \cite{Savkare2011_1} an approach  to detect red blood cells with consecutive classification into parasite infected and normal cells for further estimation of parasitemia is  proposed.
-For separation of overlapping cells watershed transform is applied on distance transform of binary mask of cells having larger area.
-
-	In \cite{Tek2010} the localisation of the parasites is achieved after a foreground and background segmentation step.
-Firstly, a rough foreground image using morphological area top-hats (using the average cell area value) is extracted.
-Then, from these rough foreground and background regions two different threshold values are determined and used in morphological double thresholding of the input grey level image to produce a refined binary foreground mask.
-From the foreground image the stained pixels are detected using a thresholding approach and finally used as markers to extract the stained objects by morphological area top-hats based on the estimated average area value.
-
-	In the study presented in \cite{Diaz2009} for quantification and classification of erythrocytes in stained thin blood films infected with Plasmodium falciparum, the authors used connected morphological operators in the segmentation step.
-The RBCs are detected as follows: firstly, a pixel classification allowed to label each image pixel as either background or foreground, based on its color features.
-Afterward, an inclusion-tree structure represented the hierarchical object relations between background and foreground so that a filtering process
-	allows to remove irrelevant structures such as artifacts generated at the staining or digitization processes.
-
-	Many systems for computer aided diagnosis of malaria disease made use of mathematical operatations in order to smoothen the boundary of the regions obtained from the segmentation process, like in \cite{Dave2017}.
-
-	%
 \subsection{Feature extraction}
@@
 Most of the studies have reported both textural and geometric features for describing malaria infection stages \cite{Das2015}.
+
 Generally speaking, features may be distinguished according to the following characteristics: morphological features and textural and intensity features.\\
+Suwanarusk et al.
+\cite{Suwanarusk2004} affirms that P.vivax infected RBCs become larger while P.falciparum infected RBCs become rigid even though they do not increase in size.
+These statements are very relevant because they permit to let morphological features play an important role in order to extract malaria-related information.
+Several morphological features can be mentioned, as follows.
+\begin{enumerate}
+	\item \textit{Area granulometry}.
 It is a well known mathematical morphology approach to compute a size distribution of grains in binary images, using a series of morphological opening operations.
@@
 Some authors used area granulometry for pre-processing purposes in malaria characterization \cite{Tek2010} even though it is certainly effective for extracting cells size features information \cite{Tek2006},\cite{Malihi2013},\cite{Springl2009}.
-In \cite{Tek2010} local area granulometry combined with colour histogram are used as features.
-The area granulometry feature is calculated locally on the binary mask of the stained objects, for the RGB channels and then concatenated.
-Morphological features are also used in \cite{Ross2006} (erosion dilation) and in \cite{Das2011} (opening, closing).
+	\item \textit{Shape features}.
+Several shape features may be used for malaria infected red blood cells characterization \cite{Das2015},\cite{Tek2006},\cite{Das2013}.
+They are strongly related with the geometric characteristic of the objects.
+Some classic examples are following listed: area, perimeter, circularity, eccentricity, orientation, major axis, minor axis.
+More complex shape features have been used as well: roundness ratio \cite{Ross2006},\cite{Theodoridis2009}, compactness and moments of inertia \cite{Tek2010}, relative ratios between chromatin dots and RBCs terms of area and centroids distances \cite{Makkapati2009}.
+	\item \textit{Relative shape measurement}.
+This kind of parameters are independent from object's position and orientation.
+Some examples are relative area, relative perimeter, major and minor axis of best-fit ellipse and largest inscribable circle \cite{Springl2009},\cite{Tek2006}.
+	\item \textit{Chain code}.
+It is generally used to represent the object boundary by covering it starting from a point and operating through the straight line of a connected component.
+Therefore, it may be used to describe the RBCs' morphology \cite{Ross2006}.
+	\item \textit{Bending energy}. It is the measure of the curvature of a boundary \cite{Theodoridis2009}.
+	\item \textit{Hu's moments}. The Hu's moments are known to be invariant under translation, rotation and scaling \cite{HuMK} and are particularly useful for RBCs' feature extraction due to this fact.
+\end{enumerate}
 
-	%
-	%
-	%%\input{Test.tex}
-	%
+Concerning textural and intensity features, texture is defined as a connected set of pixels satisfying a given grey level property which occurs repeatedly in an image region.
+It represents the variation of the intensity of a surface, quantifying properties such as smoothness, coarseness, and regularity.
+The most used methods for texture analysis are following listed:
+
+\begin{enumerate}
+	\item \textit{Colour channels intensity}.
+Several colour intensity information has been considered to infected RBCs characterization, like HSI's saturation channel, RGB's red or L and B of LAB colour space \cite{Kareem2012}, \cite{Yunda2012}.
+	\item \textit{Colour histogram}.
+Several colour channels have been involved in order to extract significant information from their histogram, e.g.
+RGB (Red, Green, Blue), HSV (Hue, Saturation, Value), HSI (Hue, Saturation, Intensity), LAB (Luminance, A-red green axis and B-blue yellow axis) \cite{Tek2010},\cite{Malihi2013},\cite{DiRuberto2002}.
+	\item \textit{Histogram}.
+Some features can be extracted from channels histogram and they are commonly known as first order statistical features, e.g.
+energy, entropy, average intensity, skewness, kurtosis \cite{Purnama2013}, \cite{Das2013}, \cite{Das2012}, \cite{Das2011}, \cite{Das2015}, \cite{Yunda2012}.
+	\item \textit{Haralick texture feature}.
+Haralick texture features are extracted from grey-level co-occurrence matrices \cite{Haralick1987}.
+A co-occurrence matrix is defined over an image to be the distribution of co-occurring pixel values at a given offset \cite{Haralick1973}.
+Gray-Level Co-Occurrence Matrix (GLCM) has been used to calculate several texture measures, like energy, entropy, correlation \cite{Elter2011}, \cite{Ross2006}, \cite{Springl2009}, \cite{Yunda2012}.
+	\item \textit{Local binary pattern}.
+Local Binary Pattern (LBP) is a texture operator which labels the pixels of an image by thresholding the neighbourhood of each pixel and considers the result as a binary number.
+Several authors recently used it to discriminate among infected and not infected cells \cite{Das2013}, \cite{Das2012}, \cite{Das2011}, \cite{Das2015}, \cite{Linder2014}, \cite{Maity2012}.
+	\item \textit{Grey-level run length matrices}.
+Grey-level run length matrices (GLRLM) are typically used to describe the texture of images, particularly the property of coarseness \cite{Galloway1975}.
+Run-length encoding is used to represent strings of symbols in an image matrix.
+For a given image a gray level run is defined as a set of consecutive, collinear pixels having the same gray level and the final length of the run is given by the number of pixels in the run.
+The property of coarseness has been used in characterization of malaria infected RBCs \cite{Das2012}, \cite{Das2011}, \cite{Das2015}, \cite{Maity2012}, \cite{Ross2006}, \cite{Springl2009}.
+	\item \textit{Entropy}.
+It measures the uncertainty associated with randomness of an image intensity \cite{Pharwaha2009}.
+Five different entropy measures have been considered for malaria infection characterization \cite{Das2013}, \cite{Das2015}.
+	\item \textit{Fractal}.
+It measures the roughness of an image over spatial distribution by using differential box counting methods \cite{Sarkar1994} and it has been involved by \cite{Das2013}, \cite{Das2015} in malaria characterization.
+	\item \textit{Flat, Gradient and Laplacian textures}.
+Flat texture is based on the difference between target images with median filter.
+Both gradient and Laplacian textures measure the velocity changes of grey intensity value of a target erythrocyte and edge information is calculated from gradient images.
+Moreover, Laplacian texture value for an infected portion of erythrocytes is higher due to its higher gradient changes \cite{Springl2009}.
+	\item \textit{Phase of an image}.
+It has been used to discriminate among parasite infected and non infected RBCs.
+It is related to the phase information of a Fourier transform image \cite{Annaldas2014}.
+	\item \textit{Wavelet feature}. Energy, mean and standard deviation are calculated following the application of discrete wavelet transform with two decomposition levels \cite{Yunda2012}.
+	\item \textit{Colour auto correlogram}.
+It defines the spatial correlation that occurs between the colour values of two pixels after applying a distance measure \cite{Huang1997}.
+Correlation of spatial colour distribution can be described globally by using colour auto correlogram \cite{Cesario2012}.
+\end{enumerate}
+
+\subsection{Feature selection}
+Feature selection is the process of selecting a subset of relevant features in order to let them be used in model construction.
+It is typically performed either to lighten the load of features extracted in the previous step and to lighten the subsequent classification step.
+In general, feature selection permits the simplification of models, the reduction of training time, to avoid the dimensionality problem and to reduce over-fitting.
+Feature selection techniques may be organized into three categories: filter, wrapper and embedded methods.
+
+Filter methods use some univariate searching function in order to rank each feature and select the highest-ranking features, where the scoring is given according to the discriminative power of each feature.
+Bayesian Network, Information Gain, Signal-to-Ratio, Euclidean Distance are among the most common univariate filter methods.
+They select variables regardless of the model and tend to suppress the least interesting variables.
+They are typically very fast to compute even though they tend to select redundant variables because they do not consider the relationships between them.
+Therefore, they are mainly used as a pre-processing method.
+
+Wrapper and Embedded Methods are specific to a given machine-learning algorithm, in spite of filter techniques that consider the ranking of each feature independently. 
+Wrapper methods evaluate subsets of variables which allows, unlike filter approaches, to detect the possible interactions between variables.
+They conduct a search by using a specific classifier in order to find the subset of features with which the classification algorithm performs the best.
+
+Embedded methods try to combine the advantages of both previous methods: learning algorithm takes advantage of its own variable selection process and performs feature selection and classification simultaneously.
+
+In malaria detection, a lot of different techniques have been used in order to find the most significant features.
+In the work \cite{Elter2011}, Elter et al.
+employed a 2-step feature selection methodology, after a normalization process to have zero mean and a standard deviation of one: an univariate ranking is first applied in order to keep only the 60 features that have the highest univariate discriminative power out of the 174 previously extracted features for MP detection, and then a genetic algorithm (GA) is used for automatic selection of an even smaller feature subset.
+Yunda et al. \cite{Yunda2012} used the Principal Component Analysis (PCA) to identify features that are redundant.
+Other examples of feature selection techniques certainly is one-way ANOVA \cite{Das2013}.
+F-statistics and information gain measure based raking has been made and their performances have been characterized with their ranking \cite{Das2015}.
+
+\subsection{Classification}
+Classification is, in general, the problem of assigning a new observation to a particular set of categories, on the basis of a training set of data in which category membership is known.
+In the terminology of machine learning, classification is considered an instance of supervised learning, i.e.
+learning where a training set of correctly identified observations is available \cite{Jayavani2014}.
+Several classification methods are well known in the literature, e.g., Naïve Bayes classifier, K-Nearest Neighbours (k-NN), Support Vector Machine (SVM), Neural Networks (NN), Genetic programming, Extremely Randomized Trees and so on.
+
+\begin{enumerate}
+	\item \textit{Naïve Bayes classifier}.
+The Naïve Bayes classifiers are a family of simple probabilistic classifiers based on the application of Bayes' theorem with strong (naive) independence assumptions between the features.
+	It assumes that the presence (or absence) of a particular feature of a class is unrelated to the presence (or absence) of any other feature, meaning that it assumes the independency of variables given the class.
+In spite of its naive design and apparently over-simplified assumptions, Naïve Bayes can often outperform more sophisticated classification methods applied in many complex real world situations.
+An advantage of the Naïve Bayes classifier is that it requires a small amount of training data to estimate the parameters of the model \cite{John1995_Bayes}.
+
+	Anggraini et al.
+\cite{Anggraini2011}, realized a Bayes Decision Rule classification model for the identification of infected RBCs and it achieved a sensitivity (SE) of 92.59\% and a specificity (SP) of 99.65\%.
+However, only 60 images were used for training and 20 for testing.
+
+	\item \textit{K-Nearest Neighbour}.
+k-NN is a non-parametric and instance-based learning method.
+It is often used when there is little or no prior knowledge about the data distribution.
+It is also based on the principle that the instances within a dataset will generally exist in close proximity to other instances with similar properties so that the classification of unknown samples is performed by comparing with already stored training data \cite{Aha1991_kNN}.
+
+	Tek et al.
+\cite{Tek2006} proposed a distance weighted k-NN classifier trained with the previously defined features in order to perform malaria parasite detection.
+They affirm that the proposed method achieved a SE of 74\% and a SP of 98\%.
+	In a different work the same aforementioned authors \cite{Tek2010} proposed an automated method for MP detection and identification.
+They tested k-NN performances along with two others classifiers, that is the Fisher linear discriminant (FLD) and the Back Propagation Neural Network (BPNN).
+The authors state that, for what regards the MP detection and identification, the k-NN method achieved 72.4\% of SE and 97.6\%, while FLD achieved 71.5 \% of SE and 93.9\% of SP and BPNN achieved $70.5\% \pm 1.7\%$ of SE and $96.6\% \pm 0.4\%$ of SP.
+
+	In the work of Malihi et al.
+\cite{Malihi2013} a method to detect malaria parasite in blood samples has been proposed and k-NN has been tested along with other five different classifiers, that is SVM, nearest mean (NM), 1-NN and Fisher linear discriminant.
+K-NN shows the best results with 91\% of accuracy (ACC), 80\% of SE and 95.5\% of SP.
+
+	\item \textit{Support Vector Machine}.
+SVM is a supervised and adaptive classification algorithm and it is one of the most widely used in medical image analysis field.
+This method constructs a hyperplane or set of hyperplanes in a high-or infinite-dimensional space, which can be used for classification.
+Intuitively, a good separation is achieved by the hyperplane that has the largest distance to the nearest training data point of any class (so-called functional margin), since in general the larger the margin the lower is the generalization error of the classifier \cite{Boser1992_SVM}, \cite{Cortes1995_SVM}.
+
+	Tek et al.
+\cite{Tek2006} used a SVM approach with Radial Basis Kernel (RBF) as function kernel.
+Considering the non linearity of the approach, additional parameters have been set: gamma G = 0.125, that is the inverse of the radius of influence of samples selected by the model as support vectors, and a cost factor of C = 1.0, that is the penalty associated to the instances which are either misclassified or violates the maximal margin.
+The authors state a SE of 97\% combined with 0.8 false-positive detections per image.
+	In \cite{Ghosh2011} the authors aim to classify malaria infected and not infected using SVM strategy.
+They state an accuracy of 98\%.
+Das et al.
+\cite{Das2013} tried to distinguish six different stages of malaria parasites: ring trophozoite, gametocyte for P.vivax and P.falciparum, schizont for P.vivax and not infected RBCs.
+They reach an accuracy of 84\%.
+On the other hand, Linder et al.
+\cite{Linder2014} classify P.falciparum infected and not infected RBCs, with 92.5\% of SE and 100\% of SP.
+	Savkare et al. \cite{Savkare2011} realized a parasite count method with a SVM as classification method, tuned with an RBF kernel. They state 93.12\% of SE and 93.17\% of SP as results.
+
+	\item \textit{Neural networks}.
+Neural networks (NNs) take their name from biological neural networks found in animal brains.
+They are especially well suited to problems that people are good at solving, but the algorithm solution is too complex to be defined.
+The most common NN model is the multilayer perceptron, which is a feed-forward artificial NN that maps sets of input data onto a set of appropriate output, using three or more layers of neurons with non-linear activation functions \cite{Umbaugh1997}.
+	Yunda et al.
+\cite{Yunda2012} used a NN for classification of gametocytes of P.falciparum, and schizonts and gametocytes of the P.
+vivax type.
+They state a 76,45\%, 76,96\% and 67,62\% of SP, respectively.
+	In \cite{Diaz}, the authors used a feed forward back propagation neural network (BFF), that is the most common NN model.
+It has three layers (input, hidden and output layers) and, typically, each one is formed of from 2 to 10 neurons.
+	Also Ross \cite{Ross2006} employed a classification based on BFF, in particular the target was to establish if an erythrocyte was infected with malaria or not, and if so, to identify the species of the parasite.
+	The first classifier is an 8-5-1 three layer network (two hidden layers and one output layer), with tan-sigmoid transfer functions.
+The output layer is determined by the number of output classes (two), while the hidden layers were designed by trial and error.
+The second classifier is a 10-10-4 three layer network.
+There are four output classes, and the hidden layers were again designed by trial and error.
+The authors state 92.07\% of SE for initial parasite detection and 85.13\% after first classification.
+
+	Some others authors used a multilayer perceptron network.
+It basically works like Feed Forward Back Propagation Neural Network (FFBPNN), even though it has multiple layers.
+Das et al.
+\cite{Das2012} classify six classes: ring trophozoite, gametocyte for P.vivax and P.falciparum, schizont for P.vivax and non-infected RBCs).
+They state a 99.72\% of SE, 84.39\% of SP, 98.64\% of Positive predictive value (PPV) and 96.3\% of ACC.
+	The same authors tried to perform two different classification in another work \cite{Das2015}: the first one aimed to distinguish among P.vivax infected RBCs and non-infected RBCs, while the second one aimed to classify six different classes of malaria (ring trophozoite, gametocyte for P.vivax and P.falciparum, schizont for P.vivax and non-infected RBCs).
+
+	Some other classification methods have been used for malaria detection or parasite count.
+In this review, the following methods are stated for knowledge, because they are not focused on mathematical morphology techniques.
+
+	\item \textit{Genetic programming}.
+In artificial intelligence, genetic programming (GP) is a technique whereby computer programs are encoded as a set of genes that are then modified using an evolutionary algorithm (often a genetic algorithm, "GA").
+Essentially genetic programming is a set of instructions and a fitness function to measure how well a computer has performed a task \cite{wiki_geneticProgramming}.
+	The authors in \cite{Purnama2013} realised an genetic programming based algorithm to identify parasites and also to detect their type and stage.
+Two classification models have been realized: the firs one aimed to distinguish among parasite and non-parasite affected cells.
+They state an ACC of 95.49\% and 95.58\%, respectively.
+The second model, instead, perform a six classes classification among the several types and stage of malaria parasites (non-parasites, P.vivax thropozoites, P.vivax schizonts, P.vivax gametocytes, P.falciparum thropozoites and P.falciparum gametocytes) with an average accuracy of 90.25\%, 82.25\%, 75.83\%, 81.75\%, 90.75\%, 86.75\%, respectively.
+
+	\item \textit{Extremely randomized trees}.
+Extremely randomized trees is a tree-based ensemble method for supervised classification and regression problems.
+It essentially consists of randomizing strongly both attribute	and cut-point choice while splitting a tree node.
+In the extreme case, it builds totally randomized trees whose structures are independent of the output values of the learning sample.
+The	strength of the randomization can be tuned to problem specifics by the appropriate choice of a parameter \cite{Geurts2006}.
+	The authors in \cite{Quinn2014} used a Extremely Randomized Trees classifier with an ensemble of 250 trees and a maximum tree depth of 5. 
+	They stated that this classifier has the advantages of being fast and memory-efficient, condition that can be very useful whether classification has to be performed for instance on a mobile device with limited computational resources.
+The authors stated an area under the curve (AUC) of 0.97 for the receiver operating characteristic (ROC) curve.
+
+	\item \textit{C4.5 classifier}.
+It is a decision tree model, realized to overcome the limitations of the previously defined ID3 classification algorithm, that is used to generate a decision tree from a dataset.
+ID3 is the precursor to the C4.5 algorithm, and is typically used in the machine learning and natural language processing domains \cite{Quinlan1993}.
+C4.5 can handle both continuous and discrete data.
+Like in ID3 the data is sorted at every node of the tree in order to determine the best splitting attribute.
+At each node of the tree, C4.5 chooses one attribute of the data that most effectively splits, by using gain ratio - a normalized version of information gain, its set of samples into subsets enriched in one class or the other.
+	Maity et al.
+\cite{Maity2012} implemented a C4.5 classifier in order to classify three classes (ring trophozoite, gametocyte for P.vivax and P.falciparum, schizont for P.vivax and non-infected RBC).
+They state 97.2\% of SE and 99.5\% of SP.
+	\item \textit{Ada-Boost}.
+The Ada-Boost algorithm of Freund and Schapire \cite{Schapire2013} is based on the concept of boosting, that is an approach to machine learning based on the idea of creating a highly accurate prediction rule by combining many relatively weak and inaccurate rules.
+This methodology has been utilized towards classifying malaria infection using 10 relevant features \cite{Vink2013}.
+They state 75\% of SE and 99.9\% of SP.
+	\item \textit{Classification and regression tree (CART)}.
+CART are machine-learning methods for constructing prediction models from data, with a top down recursive divide-and-conquer approach.
+The models are obtained by recursively partitioning the data space and fitting a simple prediction model within each partition.
+The authors in \cite{Das2012} used this method for RBCs infection stage classification.
+They used different sizes for feature set, from 10 to 80, and they state, as best results, 70.11\% of SP (with feature set of size 0) and and 97.94\% of SE (with feature set of size 70).
+	\item \textit{Logistic regression}.
+It is a supervised probabilistic classification approach and it is based on regression parameters.
+Multivariate logistic regression model is a good predictive	model for predicting whether erythrocytes are P.vivax infected or not.
+Das et al.
+\cite{Das2011} used this method in order to predict statistics of normal and P.vivax infection.
+They report an ACC of 88.77\%.
+\end{enumerate}
+
+%\input{Test.tex}
+
 \section{Discussions}

### 2017-11-17 Paper.tex
Added table comparisons

@@
 
-	The paper in \cite{Kareem2012_1} illustrates a morphological approach for blood cell identification and uses the image features such as intensity, histogram, relative size and geometry for further analysis.
+	The paper in \cite{Kareem2012} illustrates a morphological approach for blood cell identification and uses the image features such as intensity, histogram, relative size and geometry for further analysis.
 Before the identification of blood cells, the authors  propose a novel morphological filtering based on the size of RBC for platelets and/or artifacts elimination.
@@
 	Several other authors attempted to use thresholding combined with morphological operation as segmentation method in their computer-aided systems.
+
+	Dave et al.
+in \cite{Dave2017} image segmentation is performed using histogram based adaptive thresholding followed by mathematical morphological operations (erosion and dilatioon).
+The detection of infected RBC is based on a unsupervised learning technique.
 
@@
 
-	The authors in \cite{Arco2015} worked on thick blood films and proposed a method that uses an adaptive thresholding based scheme, which also allows an effective classification of pixels.
+	The authors in \cite{Arco2014} worked on thick blood films and proposed a method that uses an adaptive thresholding based scheme, which also allows an effective classification of pixels.
 This means that the election of whether a pixel belongs to the background or to the signal (parasites and white blood cells) is only established by the pixels around it, that is its neighbourhood.
@@
 
-	To conlude many systems for computer aided diagnosis of malaria disease made use of mathematical operatations in order to smoothen the boundary of the regions obtained from the segmentation process, like in \cite{Dave2017}.
+	To conlude many systems for computer aided diagnosis of malaria disease made use of mathematical operatations in order to smoothen the boundary of the regions obtained from the segmentation process.
 
@@
 
+	\input{Table_sum.tex}
+
 	\inputencoding{latin2}

### 2017-11-17 Table_sum.tex
Added table comparisons

	\setlength\LTleft{-0.75in}
%	\setlength\LTright{-1in}
%\small
\footnotesize
    \begin{longtable}{C{2.1cm} C{2.1cm} C{2.3cm} C{1.5cm} C{2.7cm} C{2.2cm} }
    %\begin{longtable}{|C{2cm}|c|C{2.5cm}|C{3cm}|C{2.4cm}|c|}

		\hline
    	\textbf{Authors} & \textbf{Preprocessing} & \textbf{Segmentation} & \textbf{Features} & \textbf{Classification} & \textbf{Performance}  \\[1pt] \hline

    Ahirwar \emph{et al.}, 2012  &
    	- &
    	thresholding + granulometry, opening, morphological gradient, dilation, closing, thinning, spur removal &
    	- &
    	five (\emph{P.falciparum}, \emph{P.vivax}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) &
    	- \vspace{0.6cm} \\
    	%\hline

    Anggraini \emph{et al.}, 2011 &
		- &
		thresholding + hole filling &	
		- &
		two (\emph{P.falciparum} infected and noninfected) + two life-cycle-stages &
		-
        \vspace{0.6cm}
		\\
    Arco \emph{et al.}, 2014 &
		- &
		adaptive thresholding + hole filling, closing, regional minima &	
		- &
		two (infected and noninfected) &
		-
		\vspace{0.6cm}
        \\
    Das \emph{et al.}, 2011 &
		- &
		marker controlled watershed &	
		opening, closing  &
		two (infected and noninfected) &
		-
		\vspace{0.6cm}
        \\

    Das \emph{et al.}, 2013 &
		- &
		marker controlled watershed &	
		- &
		three (\emph{P.falciparum}, \emph{P.vivax} infected and noninfected) + three life-cycle-stages per species &
		-
		\vspace{0.6cm}
        \\

    Das \emph{et al.}, 2014 &
		- &
		marker controlled watershed &	
		- &
		three (\emph{P.falciparum}, \emph{P.vivax} infected and noninfected) + three life-cycle-stages per species &
		-
		\vspace{0.6cm}
        \\

    Dave \emph{et al.}, 2017 &
		- &
		adaptive thresholding + erosion, dilation &	
		- &
		two (infected and noninfected) &
		-
		\vspace{0.6cm}
        \\

    Devi \emph{et al.}, 2017 &
		- &
		marker controlled watershed &	
		- &
		- &
		-
		\vspace{0.6cm}
        \\

        Dey \emph{et al.}, 2015 &
		- &
		opening &	
		- &
		- &
		-
		\vspace{0.6cm}
        \\

    Diaz \emph{et al.}, 2009 &
		- &
		inclusion tree &	
		- &
		two (\emph{P.falciparum} infected and noninfected) + three life-cycle-stages &
		-
\vspace{0.6cm}
		\\

    Di Ruberto \emph{et al.}, 2002 &
		- &
		thresholding + granulometry, watershed  transform &	
		skeleton &
		two (\emph{P.falciparum} infected and noninfected) + three life-cycle-stages &
		-
\vspace{0.6cm}
		\\

    Elter \emph{et al.}, 2011 &
		- &
		thresholding + black top-hat &	
		- &
		two (infected and noninfected) &
		-
		\vspace{0.6cm} \\

    Gonzalez-Betancourt \emph{et al.}, 2016  &
    	morphological filter, erosion-reconstruction, dilation-reconstruction, closing &
    	-  &
    	- &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

    Ghosh \emph{et al.}, 2011  &
    	- &
    	thresholding + opening, closing  &
    	- &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

    Kareem \emph{et al.}, 2012  &
    	dilation, erosion &
    	-  &
    	- &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

    Khan \emph{et al.}, 2011  &
    	- &
    	marker controlled watershed  &
    	- &
    	five (\emph{P.falciparum}, \emph{P.vivax}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) &
    	- \vspace{0.6cm} \\

    Khan \emph{et al.}, 2014  &
    	- &
    	- &
    	- &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

    Malihi \emph{et al.}, 2013  &
    	- &
    	-  &
    	area granulometry &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

    Mushabe \emph{et al.}, 2013  &
    	- &
    	thresholding + quali op.  &
    	- &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

    Oliveira \emph{et al.}, 2017  &
    	erosion &
    	-  &
    	- &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

    Reni \emph{et al.}, 2015  &
    	new morphological filtering &
    	-  &
    	- &
    	- &
    	- \vspace{0.6cm} \\

    Romero-Rondon \emph{et al.}, 2016  &
    	opening &
    	marker controlled watershed, erosion  &
    	- &
    	- &
    	- \vspace{0.6cm} \\

    Rosado \emph{et al.}, 2017  &
    	- &
    	adaptive thresholding + closing  &
    	- &
    	- &
    	- \vspace{0.6cm} \\

    Ross \emph{et al.}, 2006  &
    	area closing &
    	thresholding + granulometry, opening, reconstruction, morphological gradient, closing, thinning  &
    	erosion, dilation &
    	- &
    	- \vspace{0.6cm} \\

    Sheikhhosseini \emph{et al.}, 2013  &
    	hole filling, opening &
    	-  &
    	- &
    	- &
    	- \vspace{0.6cm} \\

    Savkare \emph{et al.}, 2011a  &
    	- &
    	watershed transform  &
    	- &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

    Savkare \emph{et al.}, 2011b  &
    	- &
    	thresholding + hole filling, watershed transform  &
    	- &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

    Savkare \emph{et al.}, 2015  &
    	- &
    	thresholding + watershed transform, erosion, dilation  &
    	- &
    	- &
    	- \vspace{0.6cm} \\

    {\v{S}}pringl, 2009  &
    	- &
    	thresholding + marker controlled watershed transform  &
    	area granulometry &
    	- &
    	- \vspace{0.6cm} \\

Somasekar \emph{et al.}, 2011  &
    	- &
    	- &
    	- &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

    Somasekar \emph{et al.}, 2015  &
    	erosion, hole filling &
    	-  &
    	- &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

    Somasekar \emph{et al.}, 2017  &
    	- &
    	thresholding + erosion, closing  &
    	- &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

    Soni \emph{et al.}, 2011  &
    	- &
    	thresholding + granulometry, morphological gradient, dilation  &
    	- &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

    Sulistyawati \emph{et al.}, 2015  &
    	- &
    	blob analysis + erosion, dilation, opening, closing  &
    	- &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

 Tek \emph{et al.}, 2006  &
    	- &
    	top-hat, infinite reconstruction  &
    	area granulometry &
    	two (infected and noninfected) &
    	- \vspace{0.6cm} \\

    Tek \emph{et al.}, 2010  &
    	closing, granulometry &
    	thresholding + granulometry, area top-hat, closing  &
    	area granulometry &
    	five (\emph{P.falciparum}, \emph{P.vivax}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) + 
    four life-cycle-stages for species &
    	- \vspace{0.6cm} \\

    Yunda \emph{et al.}, 2012  &
    	- &
    	thresholding + morphological gradient, erosion, dilation  &
    	- &
    	- &
    	- \\

		 \hline

	\caption{Summary of analysed methods: morphological operations used in the main phases of analysis, kind of classification and performance measures.} % needs to go inside longtable environment
	\label{tab:summary_table}
	\end{longtable}

### 2017-11-20 Paper.tex
Images to insert

@@
 	Savkare et al.
-\cite{Savkare2011} worked on thin blood films with Giemsa staining and used global threshold and Otsu threshold \cite{Otsu1975} on grayscale enhanced image (green channel) for separating foreground from background.
+\cite{Savkare2011b} worked on thin blood films with Giemsa staining and used global threshold and Otsu threshold \cite{Otsu1975} on grayscale enhanced image (green channel) for separating foreground from background.
 Hole filling has been performed on identified cells and morphological operators have been used to identify overlapping cells.
@@
 
-	In \cite{Savkare2011_1} an approach  to detect red blood cells with consecutive classification into parasite infected and normal cells for further estimation of parasitemia is  proposed.
+	In \cite{Savkare2011a} an approach  to detect red blood cells with consecutive classification into parasite infected and normal cells for further estimation of parasitemia is  proposed.
 For separation of overlapping cells watershed transform is applied on distance transform of binary mask of cells having larger area.
@@
 The area granulometry feature is calculated locally on the binary mask of the stained objects, for the RGB channels and then concatenated.
-Morphological features are also used in \cite{Ross2006} (erosion dilation), in \cite{Das2011} (opening, closing) and in \cite{DiRuberto2002} (skeleton) to classify parasites.
+Morphological features are also used
+	%in \cite{Ross2006} (erosion dilation),
+	in \cite{Das2011} (opening, closing) and in \cite{DiRuberto2002} (skeleton) to classify parasites.
 	%
@@
 In particular, a lot of authors affirm that Otsu thresholding suffers from limitations when textural variation is high, while histogram thresholding can not deal sufficiently good in identifying valley regions in case of unimodal histograms.
-However, such a simple and fast approach can greatly benefit from mathematical morphology as recent studies demonstrate (\cite{Ahirwar2012}, \cite{Anggraini2011}, \cite{Arco2014}, \cite{DiRuberto2002}, \cite{Elter2011}, \cite{Ghosh2011},  \cite{Kareem2011}, \cite{Mushabe2013},  \cite{Rosado2017}, \cite{Ross2006}, \cite{Savkare2011}, \cite{Savkare2015}, \cite{Somasekar2017}, \cite{Tek2010}).
+However, such a simple and fast approach can greatly benefit from mathematical morphology as recent studies demonstrate (\cite{Ahirwar2012}, \cite{Anggraini2011}, \cite{Arco2014}, \cite{DiRuberto2002}, \cite{Elter2011}, \cite{Ghosh2011},  \cite{Kareem2011}, \cite{Mushabe2013},  \cite{Rosado2017}, \cite{Ross2006}, \cite{Savkare2011b}, \cite{Savkare2015}, \cite{Somasekar2017}, \cite{Tek2010}).
 
 	Another greatly used segmentation approach is clearly the watershed transform.
-The classic watershed approach is reported to produce over segmentation results \cite{Savkare2011}, whereas the marker controlled approach does not suffer from this issue and it is reported to be very effective for overlapping cells segmentation even though some authors affirm that it may fail to segment highly overlapped cells  (\cite{Das2011}, \cite{Das2013}, \cite{Das2014}, \cite{Das2015}, \cite{Devi2017}, \cite{Khan2011}, \cite{Romero2016}, \cite{Savkare2015}, \cite{Savkare2011_1}, \cite{Springl2009}).
+The classic watershed approach is reported to produce over segmentation results \cite{Savkare2011}, whereas the marker controlled approach does not suffer from this issue and it is reported to be very effective for overlapping cells segmentation even though some authors affirm that it may fail to segment highly overlapped cells  (\cite{Das2011}, \cite{Das2013}, \cite{Das2014}, \cite{Das2015}, \cite{Devi2017}, \cite{Khan2011}, \cite{Romero2016}, \cite{Savkare2015}, \cite{Savkare2011a}, \cite{Springl2009}).
 
@@
 	The analysed works performed classification phase for different purposes.
-The majority of them aimed to distinguish among two classes only, malaria infected and noninfected RBCs, or to detect and count parasites in a malaria blood image (\cite{Anggraini2011}, \cite{Arco2014}, \cite{Das2011}, \cite{Dave2017}, \cite{Das2015}, \cite{DiRuberto2002}, \cite{Elter2011}, \cite{Ghosh2011}, \cite{Kareem2012}, \cite{Khan2014}, \cite{Malihi2013}, \cite{Mushabe2013}, \cite{Oliveira2017}, \cite{Savkare2011}, \cite{Savkare2011_1}, \cite{Somasekar2011}, \cite{Somasekar2015}, \cite{Somasekar2017}, \cite{Soni2011}, \cite{Sulist2015}, \cite{Tek2006}).
+The majority of them aimed to distinguish among two classes only, malaria infected and noninfected RBCs, or to detect and count parasites in a malaria blood image (\cite{Anggraini2011}, \cite{Arco2014}, \cite{Das2011}, \cite{Dave2017}, \cite{Das2015}, \cite{DiRuberto2002}, \cite{Elter2011}, \cite{Ghosh2011}, \cite{Kareem2012}, \cite{Khan2014}, \cite{Malihi2013}, \cite{Mushabe2013}, \cite{Oliveira2017}, \cite{Savkare2011b}, \cite{Savkare2011a}, \cite{Somasekar2011}, \cite{Somasekar2015}, \cite{Somasekar2017}, \cite{Soni2011}, \cite{Sulist2015}, \cite{Tek2006}).
 

### 2017-11-22 Paper.tex
Added new images and renewed fig.1. References ok. Conclusion reviewed.

@@
 Although this procedure requires highly trained experts, it is error-prone and could be affected by inter-observer variations.
-Moreover, blood cells images taken from microscope could vary in their illumination and colouration conditions.
+Moreover, blood cells images taken from microscope could vary in their illumination and colouration conditions, as shown in fig.
+\ref{fig:images_types}.
 Typical blood cells images contain three main components of interest: the platelets (or thrombocytes), the red blood cells (or erythrocytes) and the white blood cells (or leukocytes).
@@
 	\centering	
-		\includegraphics[width=2.75cm]{img/f3_Pfalciparum}
-		\includegraphics[width=2.75cm]{img/f3_Pvivax}
-		\includegraphics[width=2.75cm]{img/f3_Povale}
-		\includegraphics[width=2.75cm]{img/f3_Pmalariae}
-		\caption{\label{fig:malaria_types}Types of human malaria: from left to right, P. Falciparum, P. Vivax, P. Ovale, P. Malariae \cite{Khan2011}.}
+	\includegraphics[width=5cm]{img/f1a.eps}
+	\includegraphics[width=5cm]{img/f1b.eps}
+	\includegraphics[width=5cm]{img/f1c.eps}
+	\includegraphics[width=5cm]{img/f1d.eps}
+	\caption{\label{fig:images_types}Different illumination conditions generate different images, because of the absence of a standardized acquisition procedure.
+From left to right: acquisition of the same smear with four microscope's brightness levels.}
 \end{figure}
 
+\iftrue
 \begin{figure}[!t]
 	\centering
-		\includegraphics[width=2.75cm, height=2.75cm]{img/1_Pfalciparum_ring}
-		\includegraphics[width=2.75cm, height=2.75cm]{img/2_Pfalciparum_trophozoite}
-		\includegraphics[width=2.75cm, height=2.75cm]{img/3_Pfalciparum_gametocyte}
-		\includegraphics[width=2.75cm, height=2.75cm]{img/4_Pfalciparum_schizont}
-		\includegraphics[width=2.75cm, height=2.75cm]{img/5_Pmalariae_ring}
-		\includegraphics[width=2.75cm, height=2.75cm]{img/6_Pmalariae_schizont}
-		\includegraphics[width=2.75cm, height=2.75cm]{img/7_Povale_trophozoites}
-		\includegraphics[width=2.75cm, height=2.75cm]{img/8_Pvivax_trophozoites}
-		\includegraphics[width=2.75cm, height=2.75cm]{img/9_Pvivax_ring}
-		\includegraphics[width=2.75cm, height=2.75cm]{img/10_Pvivax_gametocyte}
-		\includegraphics[width=2.75cm, height=2.75cm]{img/11_Pvivax_ring}
-		\caption{\label{fig:malaria_stages}Examples of malaria parasite stages.
-From top left: P.falciparum ring, trophozoite, gametocyte, schizont; P.malariae ring and schizont; P.ovale and P.vivax trophozoites; P.vivax ring and gametocyte; P.vivax ring \cite{Tek2010}}
+	\includegraphics[width=3.5cm]{img/f3_Pfalciparum}
+	\includegraphics[width=3.5cm]{img/f3_Pvivax}
+	\includegraphics[width=3.5cm]{img/f3_Povale}
+%	\includegraphics[width=2.75cm]{img/f3_Pmalariae}
+	\caption{\label{fig:malaria_types}Types of human malaria: from left to right, P.
+Falciparum in its schizont stage, P.
+Vivax in two gametocytes specimens and one ring stage, P.
+Ovale in its ring stage.}
 \end{figure}
+\fi
+
+\iftrue
+\begin{figure}[!t]
+	\centering
+	\includegraphics[width=2.75cm, height=2.75cm]{img/falciparum_1_ring}
+	\includegraphics[width=2.75cm, height=2.75cm]{img/falciparum_2_trophozoiteAge}
+	\includegraphics[width=2.75cm, height=2.75cm]{img/falciparum_3_schizont}
+	\includegraphics[width=2.75cm, height=2.75cm]{img/falciparum_4_gametocyte}
+	\includegraphics[width=2.75cm, height=2.75cm]{img/ovale_1_ring}
+	\includegraphics[width=2.75cm, height=2.75cm]{img/ovale_2_trophozoite}
+	\includegraphics[width=2.75cm, height=2.75cm]{img/ovale_3_schizont}
+	\includegraphics[width=2.75cm, height=2.75cm]{img/ovale_4_gametocyte}
+	\includegraphics[width=2.75cm, height=2.75cm]{img/vivax_1_ring}
+	\includegraphics[width=2.75cm, height=2.75cm]{img/vivax_2c_trophozoiteDeveloped}
+	\includegraphics[width=2.75cm, height=2.75cm]{img/vivax_4_gametocyte}
+	\caption{\label{fig:malaria_stages}Examples of malaria parasite stages. From top left: P.falciparum ring, trophozoite, schizont, gametocyte;
+		 P.ovale ring, trophozoite, schizont, gametocyte; P.vivax ring, developed trophozoite, gametocyte. \cite{Tek2010}}
+\end{figure}
+\fi
 
@@
 	\centering
-		\includegraphics[height=0.15\textheight]{img/f2_ThickGiemsa}
-		\includegraphics[height=0.15\textheight]{img/f2_ThickLeishman}
-		\includegraphics[height=0.15\textheight]{img/f2_ThinGiemsa}
-		\includegraphics[height=0.15\textheight]{img/f2_ThinLeishman}
+	\includegraphics[height=0.15\textheight]{img/f3_ThickGiemsa}
+	\includegraphics[height=0.15\textheight]{img/f3_ThickLeishman}
+	\includegraphics[height=0.15\textheight]{img/f3_ThinGiemsa}
+	\includegraphics[height=0.15\textheight]{img/f3_ThinLeishman}
 	\caption{\label{fig:thick_thin_film}Malaria infected blood smear images.
@@
 Another greatly used segmentation approach is clearly the watershed transform.
-The classic watershed approach is reported to produce over segmentation results \cite{Savkare2011}, whereas the marker controlled approach does not suffer from this issue and it is reported to be very effective for overlapping cells segmentation even though some authors affirm that it may fail to segment highly overlapped cells  (\cite{Das2011}, \cite{Das2013}, \cite{Das2014}, \cite{Das2015}, \cite{Devi2017}, \cite{Khan2011}, \cite{Romero2016}, \cite{Savkare2015}, \cite{Savkare2011a}, \cite{Springl2009}).
+The classic watershed approach is reported to produce over segmentation results \cite{Savkare2011b}, whereas the marker controlled approach does not suffer from this issue and it is reported to be very effective for overlapping cells segmentation even though some authors affirm that it may fail to segment highly overlapped cells  (\cite{Das2011}, \cite{Das2013}, \cite{Das2014}, \cite{Das2015}, \cite{Devi2017}, \cite{Khan2011}, \cite{Romero2016}, \cite{Savkare2015}, \cite{Savkare2011a}, \cite{Springl2009}).
 
@@
 
-	The computer vision methodologies reported in the literature are based on light microscopic images of human peripheral blood smears for computer-aided detection of malaria parasite and its various stages.
+The computer vision methodologies reported in the literature are based on light microscopic images of human peripheral blood smears for computer-aided detection of malaria parasites and their several stages.
 Image preprocessing, segmentation of erythrocytes and parasites, malaria parasite feature extraction, malaria detection techniques have been discussed here.
@@
 However, as the number and sophistication of non-linear filters have increased, so has the variety of applications for these filters.
-The shape-based methods of mathematical morphology, in particular, are now used in a wide variety of medical applications, including electrocardiography, ultrasound imaging, radiology, and histologic image analysis \cite{Schulze1994}.
+The shape-based methods of mathematical morphology, in particular, are now used in a wide variety of medical applications, including electrocardiography, ultrasound imaging, radiology, and histological image analysis \cite{Schulze1994}.
 

### 2017-11-22 Table_sum.tex
Images to insert
Added new images and renewed fig.1. References ok. Conclusion reviewed.

@@
 		two (\emph{P.falciparum} infected and noninfected) + two life-cycle-stages &
-		-
+		SE=93\% SP=99\%
         \vspace{0.6cm}
@@
 		two (infected and noninfected) &
-		-
+		Acc=96.46\%
 		\vspace{0.6cm}
@@
 		two (infected and noninfected) &
-		-
+		Acc=88.77\%
 		\vspace{0.6cm}
@@
 	three (\emph{P.falciparum}, \emph{P.vivax} infected and noninfected) + three life-cycle-stages per species &
-		-
+		Acc=84\%
 		\vspace{0.6cm}
@@
 	three (\emph{P.falciparum}, \emph{P.vivax} infected and noninfected) + three life-cycle-stages per species &
-		-
+		SE=99.72\% SP=84.39\%
 		\vspace{0.6cm}
@@
 		two (infected and noninfected) &
-		-
+		Acc=97.83\% thin films, Acc=89.88\% thick films
 		\vspace{0.6cm}
@@
 		- &
-		- &
-		-
-		\vspace{0.6cm}
-        \\
-
-        Dey \emph{et al.}, 2015 &
-		- &
-		opening &	
-		- &
-		- &
-		-
+		two (infected and noninfected)  &
+		Acc=98.02\%
 		\vspace{0.6cm}
@@
 		two (\emph{P.falciparum} infected and noninfected) + three life-cycle-stages &
-		-
+		SE=94\% SP=99.7\% for detection, SE=78.8\% SP=91.2\% for life-stages
 \vspace{0.6cm}
@@
     Di Ruberto \emph{et al.}, 2002 &
-		- &
+		area closing &
 		thresholding + granulometry, watershed  transform &	
@@
 		- &
-		thresholding + black top-hat &	
+		thresholding + black top-hat, dilation &	
 		- &
 		two (infected and noninfected) &
-		-
+		SE=97\%
 		\vspace{0.6cm} \\
@@
     	- &
-    	two (infected and noninfected) &
+    	- &
     	- \vspace{0.6cm} \\
@@
     	- &
-    	two (infected and noninfected) &
+    	two (\emph{P.vivax} infected and noninfected) &
     	- \vspace{0.6cm} \\
@@
     	two (infected and noninfected) &
-    	- \vspace{0.6cm} \\
+    	Acc=88\% SE=90\% SP=86\% \vspace{0.6cm} \\
 
     Khan \emph{et al.}, 2011  &
-    	- &
-    	marker controlled watershed  &
+    	area closing &
+    	thresholding + granulometry, opening, morphological reconstruction, gradient, dilation &
     	- &
     	five (\emph{P.falciparum}, \emph{P.vivax}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) &
-    	- \vspace{0.6cm} \\
+    	Acc=81\% SE=85.5\% \vspace{0.6cm} \\
 
@@
     	two (infected and noninfected) &
-    	- \vspace{0.6cm} \\
+    	SE=93\% SP=95\% \vspace{0.6cm} \\
 
     Malihi \emph{et al.}, 2013  &
-    	- &
+    	closing &
     	-  &
@@
     	two (infected and noninfected) &
-    	- \vspace{0.6cm} \\
+    	Acc=91\% SE=80\% SP=95.5\% \vspace{0.6cm} \\
 
     Mushabe \emph{et al.}, 2013  &
-    	- &
-    	thresholding + quali op.  &
+    	closing &
+    	thresholding + granulometry, dilation, erosion  &
     	- &
     	two (infected and noninfected) &
-    	- \vspace{0.6cm} \\
+    	SE=98.5 SP=97.2\% \vspace{0.6cm} \\
 
@@
     	two (infected and noninfected) &
-    	- \vspace{0.6cm} \\
+    	Acc=91\% \vspace{0.6cm} \\
 
@@
     	- &
-    	- &
-    	- \vspace{0.6cm} \\
+    	four (\emph{P.falciparum}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) +
+    three life-cycle-stages for species &
+    	SE=73.9-96.2\% SP=92.6-99.3\% \vspace{0.6cm} \\
 
@@
     	thresholding + granulometry, opening, reconstruction, morphological gradient, closing, thinning  &
-    	erosion, dilation &
         - &
-    	- \vspace{0.6cm} \\
-
-    Sheikhhosseini \emph{et al.}, 2013  &
-    	hole filling, opening &
-    	-  &
-    	- &
-    	- &
-    	- \vspace{0.6cm} \\
+    	five (\emph{P.falciparum}, \emph{P.vivax}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) &
+    	SE=85\% for detection, Acc=73\% for classification
+    	 \vspace{0.6cm} \\
 
@@
     	- &
-    	watershed transform  &
+    	thresholding + hole filling, watershed transform  &
     	- &
@@
     	two (infected and noninfected) &
-    	- \vspace{0.6cm} \\
+    	SE=93.12\% SP=93.17\% \vspace{0.6cm} \\
 
@@
     	- &
+    	two (infected and noninfected) &
+    	Acc=95.5\% \vspace{0.6cm} \\
+
+    Sheikhhosseini \emph{et al.}, 2013  &
     	- &
-    	- \vspace{0.6cm} \\
-
-    {\v{S}}pringl, 2009  &
-    	- &
-    	thresholding + marker controlled watershed transform  &
-    	area granulometry &
-    	- &
-    	- \vspace{0.6cm} \\
-
-Somasekar \emph{et al.}, 2011  &
-    	- &
-    	- &
+    	thresholding + hole filling, opening  &
     	- &
     	two (infected and noninfected) &
-    	- \vspace{0.6cm} \\
+    	Acc=97.25\% SE=82.21\% SP=98.02\% \vspace{0.6cm} \\
 
     Somasekar \emph{et al.}, 2015  &
-    	erosion, hole filling &
     	- &
+    	fuzzy C-means clustering + erosion, hole filling  &
     	- &
     	two (infected and noninfected) &
-    	- \vspace{0.6cm} \\
+    	SE=98\% SP=93.3\% \vspace{0.6cm} \\
 
@@
     	- &
-    	thresholding + erosion, closing  &
+    	thresholding + erosion, closing, hole filling  &
     	- &
     	two (infected and noninfected) &
-    	- \vspace{0.6cm} \\
+    	average DSC=0.8 \vspace{0.6cm} \\
 
@@
     	- &
+    	five (\emph{P.falciparum}, \emph{P.vivax}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) &
+    	SE=98\% for detection  \vspace{0.6cm} \\
+
+{\v{S}}pringl, 2009  &
+    	closing &
+    	thresholding + marker controlled watershed transform, hole filling, dilation, opening, erosion  &
+    	- &
     	two (infected and noninfected) &
-    	- \vspace{0.6cm} \\
+    	AUC=0.98 \vspace{0.6cm} \\
 
@@
     	- &
-    	blob analysis + erosion, dilation, opening, closing  &
+    	blob analysis + erosion, dilation, opening, closing, hole filling  &
     	- &
     	two (infected and noninfected) &
-    	- \vspace{0.6cm} \\
+    	Acc=99.39\% \vspace{0.6cm} \\
 
@@
     	two (infected and noninfected) &
-    	- \vspace{0.6cm} \\
+    	SE=74\% SP=98\% \vspace{0.6cm} \\
 
@@
     four life-cycle-stages for species &
-    	- \vspace{0.6cm} \\
+    	SE=72\% SP=98\% \vspace{0.6cm} \\
 
@@
     	- &
-    	- &
-    	- \\
+    	three (\emph{P.falciparum}, \emph{P.vivax} infected, and noninfected) + two life-cycle-stages for \emph{P.falciparum}&
+    	SE=77.19\% \\
 
@@
 
-	\caption{Summary of analysed methods: morphological operations used in the main phases of analysis, kind of classification and performance measures.} % needs to go inside longtable environment
+	\caption{Summary of analysed methods: morphological operations used in the main phases of analysis, kind of classification and performance measures (Sensitivity, Specificity, Accuracy, if reported).} % needs to go inside longtable environment
 	\label{tab:summary_table}

### 2017-11-24 Paper.tex
Updated conclusion and named-ordered authors in pre-proc and segmentation sections

@@
 \begin{document}
-\title{Recent advances of Malaria parasites detection systems based on mathematical morphology techniques}
+\title{Recent advances of malaria parasites detection systems based on mathematical morphology}
 
@@
 The differential white blood cell count includes measurements of the different types of white blood cells that constitute the total white blood cell count: the band neutrophils, segmented neutrophils, lymphocytes, monocytes, eosinophils, and basophils.
-A specific infection can be suspected on the basis of the type of leukocyte that has an abnormal value.
+A specific infection can be suspected on the basis of the type of leukocyte that has an abnormal value \cite{DiRuberto2016}.
 
@@
 Numerous methods of automatic malaria diagnosis have been proposed so far, in order to overcome the issues before mentioned.
-The aim of this paper is to review and analyse the works of different researchers who in particular have used mathematical morphology as a tool for computer aided malaria detection and classification.
+The aim of this paper is to review and analyse the works of different researchers who in particular have used mathematical morphology as a powerful tool for computer aided malaria detection and classification.
 
@@
 They are defined in terms of more elementary set operations, but are employed as the basic elements of many algorithms.
-Both dilation and erosion are produced by the interaction of a set called structuring element with a set of pixels of interest in the image.
+Both dilation and erosion are produced by the interaction of a set called structuring element (SE) with a set of pixels of interest in the image.
 The structuring element has both a shape and an origin.
+From these two basic operators, others have been derived (opening, closing, hit-or-miss).
+They can be applied to extract image components useful in the representation and descriptions of region shapes, such as boundaries, skeleton, or convex hull.
+Also morphological operators can be used for image preprocessing and postprocessing, such as morphological filtering, thinning, and especially for segmentation.
 
@@
 This section presents a review of some of the main recent studies existing in literature regarding the analysis of malaria infected blood smears using mathematical morphology.
-A typical approach usually comprises five different image processing and analysis tasks, as follows:
+A typical approach usually comprises four different image processing and analysis tasks, as follows:
 \begin{enumerate}
@@
 	\item Feature extraction.
-	\item Feature selection.
+	%\item Feature selection.
 	\item Classification.
@@
 It is worth to mention pre-processing methods because they are an important step regarding image analysis field but, for what concerns the malaria-affected blood image analysis, in our review we particularly found methods which operate for illumination correction and noise filtering purposes.
-Generally speaking, digital microscopy images can be acquired in different lighting conditions, with several types of acquisition devices or from blood smears stained with various staining protocol and, consequently, the features of similar images could differ a lot.
-Different techniques for illumination correction have been suggested to reduce such variation, e.g., a lot of authors work with grayscale-converted images as an illumination correction method %\cite{Tek2010},\cite{Das2015},\cite{Tek2006}.
+Generally speaking, digital microscopy images can be acquired in different lighting conditions, with several types of acquisition devices or from blood smears stained with various staining protocols and, consequently, the features of similar images could differ a lot.
+Different techniques for illumination correction have been suggested to reduce such variation, e.g., a lot of authors work with grayscale-converted images as an illumination correction method.
+%\cite{Tek2010},\cite{Das2015},\cite{Tek2006}.
 On the other hand, noise filtering aims to remove the noise introduced by mishandling the slides and/or the camera settings.
@@
 Morphological operators are useful for removal of unwanted
-objects, holes filling, splitting, thinning and thinking.
+objects, holes filling, splitting, thinning and thickening.
 Different researchers during automated diagnosis of malaria used morphological operations in preprocessing phase and the most recent are listed below.
 
-The system proposed in \cite{Oliveira2017} is based on image processing, artificial intelligence techniques and an adapted face detection algorithm to identify Plasmodium parasites.
-The algorithm uses the integral image and haar-like features concepts, and weak classifiers with adaptive boosting learning.
+In \cite{Gonzalez2016} Gonzalez-Betancourt \emph{et al.} proposed a system to determine markers for watershed segmentation based on the Radon transform and mathematical operators.
+In the first step of the process small irrelevant structures and part of the noise are eliminated by a morphological filter, in order to ensure the preservation of the cells edges.
+Image smoothing is performed by a morphological erosion-reconstruction and dilation-reconstruction filter with a disk structuring element of radius equal to 20 pixels, which is $0.274$ times smaller than the average radius of the RBCs.
+In this way the influences of the size and the shape of the structures can be separated in the smoothing process.
+At the same time the objects which are not eliminated remain unchanged.
+Also, a morphological closing is performed with a disc structuring element having radius smaller than half the average of the RBCs' radii, in order to connect the possible (more than one) markers that can appear on a single cell.
+
+In \cite{Kareem2012} Kareen \emph{et al.} illustrated a morphological approach for blood cell identification and use the image features such as intensity, histogram, relative size and geometry for further analysis.
+Before the identification of blood cells, the authors  propose a novel morphological filtering based on the size of RBC for platelets and/or artifacts elimination.
+A dilation is performed by a concentric ring structuring element and erosion by disk shaped structuring element.
+The radius of the structuring element depends on the radius of the RBC, so that all the components smaller than the RBCs can be removed.
+
+The system proposed in \cite{Oliveira2017} by Oliveira \emph{et al.} is based on image processing, artificial intelligence techniques and an adapted face detection algorithm to identify Plasmodium parasites.
+The latter uses the integral image and haar-like features concepts, and weak classifiers with adaptive boosting learning.
 The search scope of the learning algorithm is reduced in the preprocessing step by removing the background around blood cells by means of morphological erosions both for training and for testing.
-%Morphological operations have been applied to images to enhance the geometric structures on a set of pixels.
-In particular, an erosion operation has been applied after a Gaussian filtering in order to locate the RBCs and restrict the search space of the algorithm.
-%The accuracy of the system was, on average, 91, meaning that for every 100 parasite-infected samples, 91 were identified correctly.
-The system has been developed for mobile phones and tablets and can represent a useful tool
-%in health centers in remote communities without the need for specific expertise.
 
-In \cite{Romero2016} an algorithm that uses morphological operations, the watershed method, the Hough transform and the clustering method of k-means to detect overlapped RBCs is presented.
+Romero \emph{et al.} in \cite{Romero2016} presented an algorithm that uses morphological operations, the watershed method, the Hough transform and the clustering method of k-means to detect overlapped RBCs.
 In the pre-processing stage white blood cells and platelets are removed before the segmentation task.
@@
 
-An edge-based segmentation of erythrocytes infected with malaria parasites using microscopic images is proposed in \cite{Somasekar2015}.
+Reni \emph{et al.} in \cite{Reni2015} described a new algorithm for morphological filtering of the blood images as a pre-processing tool for segmentation.
+Conventional morphological closing on blood images removes the unwanted components but also useful information.
+On the opposite the proposed method preserves the necessary information of foreground components while removing noise and artefacts.
+
+In the method proposed in \cite{Sheik2013} by Sheikhhosseini \emph{et al.} the first phase is the stained object extraction which detects candidates objects that can be infected by malaria parasites using intensity and colour.
+Before detecting the stained objects the method firstly extracts the foreground.
+Foreground image is a binary image which is produced after applying morphological hole filling on such pixels which have lower intensity value than average intensity value of green layer.
+After the stained objects extraction process, a series of morphological operations is also employed in order to eliminate small components and complete the final stained objects.
+
+An edge-based segmentation of erythrocytes infected with malaria parasites using microscopic images is proposed by Somasekar \emph{et al.} in \cite{Somasekar2015}.
 A fuzzy C-means clustering is applied to extract infected erythrocytes, which is further processed for the final segmentation.
@@
 
-The work \cite{Reni2015} proposes a new algorithm for morphological filtering of the blood images as a pre-processing tool for segmentation.
-Conventional morphological closing on blood images removes the unwanted components but also useful information.
-On the opposite the proposed method preserves the necessary information of foreground components while removing noise and artefacts.
-
-The paper in \cite{Kareem2012} illustrates a morphological approach for blood cell identification and uses the image features such as intensity, histogram, relative size and geometry for further analysis.
-Before the identification of blood cells, the authors  propose a novel morphological filtering based on the size of RBC for platelets and/or artifacts elimination.
-A dilation is performed by a concentric ring structuring element and erosion by disk shaped structuring element.
-The radius of the structuring element depends on the radius of the RBC, so that all the components smaller than the RBCs are removed.
-
-In \cite{Tek2010} a complete framework to detect and identify malaria parasites in images of Giemsa stained thin blood film specimens is proposed.
+In \cite{Tek2010} Tek \emph{et al.} presented a complete framework to detect and identify malaria parasites in images of Giemsa stained thin blood film specimens.
 Also, the system is able to identify the infecting species and life-cycle stages.
@@
 The sufficiently large size for an input image is determined automatically with respect to its average cell size computed from the area granulometry distribution.
-
-In the method proposed in \cite{Sheik2013} the first phase is the stained object extraction which detects candidates objects that can be infected by malaria parasites using intensity and colour.
-Before detecting the stained objects the method firstly extracts the foreground.
-Foreground image is a binary image which is produced after applying morphological hole filling on such pixels which have lower intensity value than average intensity value of green layer.
-After the stained objects extraction process, a series of morphological operations is also employed after every part in order to eliminate small components and complete the final stained objects.
-
-In \cite{Gonzalez2016} a system to determine markers for watershed segmentation based on the Radon transform and mathematical operators is proposed.
-In the first step of the process small irrelevant structures and part of the noise are eliminated by a morphological filter, in order to ensure the preservation of the cells edges.
-Image smoothing is performed by a morphological erosion-reconstruction and dilation-reconstruction filter with a disk structuring element of radius equal to 20 pixels, which is $0.274$ times smaller than the average radius of the RBCs.
-In this way the influences of the size and the shape of the structures can be separated in the smoothing process.
-At the same time the objects which are not eliminated remain unchanged.
-Also, a morphological closing is performed with a disc structuring element having radius smaller than half the average of the RBCs' radii, in order to connect the possible (more than one) markers that can appear on a single cell.
 
@@
 
-Several other authors attempted to use thresholding combined with morphological operation as segmentation method in their computer-aided systems.
+Several other authors attempted to use thresholding combined with morphological operation as segmentation method in their computer-aided systems and they are described as follows.
 
-Dave et al.
-in \cite{Dave2017} image segmentation is performed using histogram based adaptive thresholding followed by mathematical morphological operations (erosion and dilatioon).
-The detection of infected RBC is based on a unsupervised learning technique.
+Arco \emph{et al.} in \cite{Arco2014} worked on thick blood films and proposed a method that uses an adaptive thresholding based scheme, which also allows an effective classification of pixels.
+This means that the election of whether a pixel belongs to the background or to the signal (parasites and white blood cells) is only established by the pixels around it, that is its neighbourhood.
+Then, morphological methods are applied to evaluate the area of connected components, labelling those belonging to parasites and counting their number.
 
-In \cite{Ross2006} the authors propose a method which provides a positive or negative diagnosis of malaria and differentiates parasites by species.
+Anggraini \emph{et al.} \cite{Anggraini2011} proposed a method for separating blood cells, parasites and other components from background in a microscopic field of a thin blood smear.
+They applied several global thresholding methods and visually compared the results to qualitatively determine which technique yields the best result.
+The binary image was then subjected to hole filling morphological operator and applied as marker to label blood cells.
+From each identified cell (RBC and WBC), constituents of the parasite (nucleus and cytoplasm) were extracted using multiple threshold.
+
+Dave \emph{et al.} in \cite{Dave2017} performed image segmentation using histogram based adaptive thresholding followed by mathematical morphological operations (erosion and dilation).
+The detection of infected RBCs is based on a unsupervised learning technique.
+
+The proposed automated method in \cite{Elter2011} by Elter \emph{et al.} for parasite detection and identification worked on thin blood film acquired with Giemsa stain.
+The authors found that the G and B channels of the RGB colour are very good features to identify objects containing chromatin in Giemsa stained blood films, being not only considered highly discriminative but also almost independent of differences in illumination and staining intensity.
+They transformed the colour input image into a monochrome image \textit{I(x,y)}, that highlights objects containing chromatin: $ I(x,y) = arctan \frac{I_{green}(x,y)}{I_{blue}(x,y)} $.
+In this work, mathematical morphology has been used with a black top-hat operator to separate MP from both leukocytes and platelets, with a non-flat paraboloid structuring element of radius of 9 and a slope of 1 pixel.
+It should be taken into account that these fixed parameters might not be suitable for images with different pixel resolutions.
+The black top-hat operator is followed by a thresholding operation with a fixed threshold, which according to the authors is reliable given the independence of the G and B channels with regard to illumination and staining intensity.
+However, the authors do not define the value of this fixed threshold on the publication.
+
+In \cite{Ghosh2011} Ghosh \emph{et al.} used divergence based threshold selection in order to segment P.vivax parasites from Leishman-stained thin blood films.
+This method is based on Cauchy membership function \cite{DiRuberto2014} and is applied to the C channel of CMYK colour space.
+Morphological operators of opening and closing have been used for artefacts removal.
+
+Kareem \emph{et al.} in \cite{Kareem2011} used the Annular Ring Ratio transform method.
+Before applying it, a pre processing phase for removing platelets, parasites and other artefacts in the image has been performed.
+In the proposed method, the image after being converted to grayscale undergoes a morphological opening similar to closing.
+Unlike conventional closing (dilation followed by erosion) which uses the same structuring element, two different structuring elements are used, a concentric ring for dilation and a disk for erosion.
+The inner and outer diameter of the dilation ring is set to 35\% and 70\% of RBCs size, respectively and the erosion disk has the same diameter.
+Therefore, considering that fixed manually defined parameters are used for this strategy, the results may substantially differ depending on the image resolution.
+This approach results in locating only the stained components in the image instead of all the cells and hence will not only speed up the operation but reduces the complexity.
+
+Mushabe \emph{et al.} \cite{Mushabe2013} used morphological and statistical classification to detect malaria in blood smears by identifying and counting red blood cells and Plasmodium parasites.
+Morphological operations and histogram-based thresholding are used to extract RBCs and boundary curvature calculations and Delaunay triangulation are used for splitting clumped RBCs.
+They worked on Giemsa-stained thin blood smears.
+
+In \cite{Ross2006} Ross \emph{et al.} proposed a method which provides a positive or negative diagnosis of malaria and differentiates parasites by species.
 The segmentation step relies on a thresholding strategy which aims to identify and segment potential parasites and erythrocytes from the image background after a six steps threshold selection.
 Mathematical morphology has been used for parasite size estimation, erythrocytes reconstruction and cells bigger than erythrocytes removal.
+
+Savkare \emph{et al.} \cite{Savkare2011b} worked on thin blood films with Giemsa staining and used global threshold and Otsu threshold \cite{Otsu1975} on grayscale enhanced image (green channel) for separating foreground from background.
+Hole filling has been performed on identified cells and morphological operators have been used to identify overlapping cells.
+Then, watershed transform has been applied for separating overlapped cells.
+
+Also in the method proposed in \cite{Somasekar2017} by Somasekar \emph{et al.} the segmentation of the infected parasites is based on thresholding.
+The segmentation is achieved in two stages by maximizing between-class variance of an original image and consequently by an iterative threshold selection from a stage-one threshold image with suitable stopping criteria.
+The segmented results are post processed to improve the accuracy of the detection of malaria parasites by morphological operators (erosion and closing).
+
 %
@@
 %dots, so a further evaluation is performed to establish if a chromatin dot is valid by identifying whether it is inside a RBC or not.
-
-The authors in \cite{Kareem2011} used the Annular Ring Ratio (ARR) transform method.
-Before applying it, a pre processing phase for removing platelets, parasites and other artefacts in the image has been performed.
-In the proposed method, the image after being converted to grayscale undergoes a morphological opening similar to closing.
-Unlike conventional closing (dilation followed by erosion) which uses the same structuring element, two different structuring elements are used, a concentric ring for dilation and a disk for erosion.
-The inner and outer diameter of the dilation ring is set to 35\% and 70\% of RBCs size, respectively and the erosion disk has the same diameter.
-Therefore, considering that fixed manually defined parameters are used for this strategy, the results may substantially differ depending on the image resolution.
-This approach results in locating only the stained components in the image instead of all the cells and hence will not only speed up the operation but reduces the complexity.
 
@@
 
-Anggraini et al.
-\cite{Anggraini2011} proposed a method for separating blood cells, parasites and other components from background in a microscopic field of a thin blood smear.
-They applied several global thresholding methods and visually compared the results to qualitatively determine which technique yields the best result.
-The binary image was then subjected to hole filling morphological operator and applied as marker to label blood cells.
-From each identified cell (RBC and WBC), constituents of the parasite (nucleus and cytoplasm) were extracted using multiple threshold.
-
-Savkare et al.
-\cite{Savkare2011b} worked on thin blood films with Giemsa staining and used global threshold and Otsu threshold \cite{Otsu1975} on grayscale enhanced image (green channel) for separating foreground from background.
-Hole filling has been performed on identified cells and morphological operators have been used to identify overlapping cells.
-Then, watershed transform has been applied for separating overlapped cells.
-
-In \cite{Ghosh2011}, authors used divergence based threshold selection in order to segment P.vivax parasites from Leishman-stained thin blood films.
-This method is based on Cauchy membership function \cite{DiRuberto2014} and is applied to the C channel of CMYK colour space.
-Morphological operators of opening and closing have been used for artefacts removal.
-
-Tek et al.
-\cite{Tek2010} addressed the problem of parasite detection and identification on thin blood film.
-They performed a pre-processing phase, in which morphology has been used with a granulometry-based cell size estimation, considering the peak index of the granulometry distribution as the average cell area.
-Then, two different segmentation steps followed: the foreground-background segmentation has been realized using morphological area top-hats (using the average cell area value) and morphological double thresholding; the stained pixels segmentation has been performed by modelling the stained and unstained pixel distribution with RGB space histograms and used the probability density function to determine if a pixel is stained or not.
-This work also addresses non-uniform illumination problem, removing it by using a pre-recorded illumination image or using a morphological closing operation.
 %
@@
 
-The authors in \cite{Arco2014} worked on thick blood films and proposed a method that uses an adaptive thresholding based scheme, which also allows an effective classification of pixels.
-This means that the election of whether a pixel belongs to the background or to the signal (parasites and white blood cells) is only established by the pixels around it, that is its neighbourhood.
-Then, morphological methods are applied to evaluate the area of connected components, labelling those belonging to parasites and counting their number.
-The following strategy has been used:
-\begin{enumerate}
-	\item an algorithm based on run-length encoding has been used for analysing the connected components of the binary masks obtained in segmentation phase;
-	\item preliminary tagging, keeping the label equivalences or connected regions with different labels in a table of local matches. Resolution of the equivalences is, then, performed;
-	\item measurements on the remaining connected components: WBCs and MP region areas are checked and WBCs connected components are removed with a closing operation.
-\end{enumerate}
+%The following strategy has been used:
+%\begin{enumerate}
+%	\item an algorithm based on run-length encoding has been used for analysing the connected components of the binary masks obtained in segmentation phase;
+%	\item preliminary tagging, keeping the label equivalences or connected regions with different labels in a table of local matches. Resolution of the equivalences is, then, performed;
+%	\item measurements on the remaining connected components: WBCs and MP region areas are checked and WBCs connected components are removed with a closing operation.
+%\end{enumerate}
 %
-Also in the method proposed in \cite{Somasekar2017} the segmentation of the infected parasites is based on thresholding.
-The segmentation is achieved in two stages by maximizing between-class variance of an original image and consequently by an iterative threshold selection from a stage-one threshold image with suitable stopping criteria.
-The segmented results are post processed to improve the accuracy of the detection of malaria
-parasites by morphological operators (erosion and closing).
-
-Mushabe et al.
-\cite{Mushabe2013} used morphological and statistical classification to detect malaria in blood smears by identifying and counting red blood cells and Plasmodium parasites.
-Morphological operations and histogram-based thresholding are used to extract RBCs and boundary curvature calculations and Delaunay triangulation are used for splitting clumped RBCs.
-They worked on Giemsa-stained thin blood smears.
-
-The proposed automated method in \cite{Elter2011} for parasite detection and identification worked on thin blood film acquired with Giemsa stain.
-The authors found that the G and B channels of the RGB colour are very good features to identify objects containing chromatin in Giemsa stained blood films, being not only considered highly discriminative but also almost independent of differences in illumination and staining intensity.
-They transformed the colour input image into a monochrome image \textit{I(x,y)}, that highlights objects containing chromatin: $ I(x,y) = arctan \frac{I_{green}(x,y)}{I_{blue}(x,y)} $.
-In this work, mathematical morphology has been used with a black top-hat operator to separate MP from both leukocytes and platelets, with a non-flat paraboloid structuring element of radius of 9 and a slope of 1 pixel.
-It should be taken into account that these fixed parameters might not be suitable for images with different pixel resolutions.
-The black top-hat operator is followed by a thresholding operation with a fixed threshold, which according to the authors is reliable given the independence of the G and B channels with regard to illumination and staining intensity.
-However, the authors do not define the value of this fixed threshold on the publication.
 
 %% Morphology and/or granulometry in Segmentation stage:
-Di Ruberto et al.
-\cite{DiRuberto2002} aimed to detect the parasites by means of an automatic thresholding based on a morphological approach applied to cell image segmentation, that is more accurate than the classical watershed-based algorithm.
-They have applied grey scale granulometries based on opening with disk-shaped elements, flat and hemispherical.
-They have used a hemispherical disk-shaped structuring element to enhance the roundness and the compactness of the red cells improving the accuracy of the classical watershed algorithm, while they have used a disk-shaped flat structuring element to separate overlapping cells.
-These methods make use of the red blood cell structure knowledge, that is not used in existing watershed-based algorithms.
+On the other hand, a lot of works have been realized by means of mathematical morphology and/or granulometry in the segmentation stages, even in combination with thresholding strategies.
+They are briefly analysed underneath.
 
-Soni et al.
-\cite{Soni2011} performed segmentation of erythrocytes by using granulometry as well.
-The size and eccentricity of the erythrocytes are also required for the calculation of some feature values (as these can be indicative of infection).
-The shape of the objects (circular erythrocytes) is known a priori, but the image must be analysed to determine the size distribution of objects in the image and to find the average eccentricity of erythrocytes present.
-Gray-scale granulometries based on opening with disk shape elements are then used.
-Non flat disk shaped structural element are applied to enhance the roundness and compactness of the red blood cells and flat disk shaped structural element applied to segment overlapping cells.
-The object to be segmented differs greatly in contrast from the background image.
-Changes in contrast can be detected by operators that calculate the gradient of an image.
-The gradient image can be computed and a threshold can be applied to create a binary mask containing the segmented cell.
-The binary gradient mask is dilated using the vertical structuring element followed by the horizontal structuring element.
-The cell of interest has been successfully segmented, but it is not the only object that has been found.
-Any objects that are connected to the border of the image can be removed.
-
-Also Airwhar et al.
-\cite{Ahirwar2012} based their approach on thresholding and granulometry.
+Airwhar \emph{et al.} \cite{Ahirwar2012} based their approach on thresholding and granulometry.
 The histogram of the complemented, green component has been used and it is said to be a bimodal distribution in all the considered images.
@@
 
-In \cite{Yunda2012}, Yunda et al.
-proposed a method for P.vivax parasites detection.
+Di Ruberto \emph{et al.} \cite{DiRuberto2002} aimed to detect the parasites by means of an automatic thresholding based on a morphological approach applied to cell image segmentation, that is more accurate than the classical watershed-based algorithm.
+They have applied grey scale granulometries based on opening with disk-shaped elements, flat and hemispherical.
+They have used a hemispherical disk-shaped structuring element to enhance the roundness and the compactness of the red cells improving the accuracy of the classical watershed algorithm, while they have used a disk-shaped flat structuring element to separate overlapping cells.
+These methods make use of the red blood cell structure knowledge, that is not used in existing watershed-based algorithms.
+
+Khan \emph{et al.} in \cite{Khan2011} presented a novel threshold selection technique used to identify erythrocytes and possible parasites present on microscopic slides that greatly takes benefit of morphological operations, such as granulometry and morphological reconstruction.
+
+In \cite{Rosado2017} Rosado \emph{et al.} proposed a system using supervised classification to assess the presence of malaria parasites and determine the species and life cycle stage in Giemsa-stained thin blood smears.
+For the RBCs segmentation, they used an adaptive thresholding approach followed by a closing morphological operation with an elliptical structuring element.
+
+Soni \emph{et al.} \cite{Soni2011} performed segmentation of erythrocytes by using granulometry as well.
+The size and eccentricity of the erythrocytes are also required for the calculation of some feature values (as these can be indicative of infection).
+The shape of the objects (circular erythrocytes) is known a priori, but the image must be analysed to determine the size distribution of objects in the image and to find the average eccentricity of erythrocytes present.
+Gray-scale granulometries based on opening with disk shape elements are then used.
+Non flat disk shaped structural element are applied to enhance the roundness and compactness of the red blood cells and flat disk shaped structural element applied to segment overlapping cells.
+The object to be segmented differs greatly in contrast from the background image.
+Changes in contrast can be detected by operators that calculate the gradient of an image.
+The gradient image can be computed and a threshold can be applied to create a binary mask containing the segmented cell.
+The binary gradient mask is dilated using the vertical structuring element followed by the horizontal structuring element.
+The cell of interest has been successfully segmented, but it is not the only object that has been found.
+Any objects that are connected to the border of the image can be removed.
+
+In \cite{Tek2010} Tek \emph{et al.} the localisation of the parasites is achieved after a foreground and background segmentation step.
+Firstly, a rough foreground image using morphological area top-hats (using the average cell area value) is extracted.
+Then, from these rough foreground and background regions two different threshold values are determined and used in morphological double thresholding of the input grey level image to produce a refined binary foreground mask.
+From the foreground image the stained pixels are detected using again a thresholding approach and finally used as markers to extract the stained objects by morphological area top-hats based on the estimated average area value.
+
+In \cite{Yunda2012} Yunda \emph{et al.} proposed a method for P.vivax parasites detection.
 The segmentation phase is a combination of border and region detection that allows rejection of the image background and permits identifying each of the objects.
@@
 
-In \cite{Rosado2017} Rosado et al.
-proposed a system using supervised classification to assess the presence of malaria parasites and determine the species and life cycle stage in Giemsa-stained thin blood smears.
-For the RBCs segmentation, they used an adaptive thresholding approach followed by a closing morphological operation with an elliptical structuring element.
+Several authors used marker controlled watershed with morphological approach, as following described.
 
-In \cite{Tek2010} the localisation of the parasites is achieved after a foreground and background segmentation step.
-Firstly, a rough foreground image using morphological area top-hats (using the average cell area value) is extracted.
-Then, from these rough foreground and background regions two different threshold values are determined and used in morphological double thresholding of the input grey level image to produce a refined binary foreground mask.
-From the foreground image the stained pixels are detected using again a thresholding approach and finally used as markers to extract the stained objects by morphological area top-hats based on the estimated average area value.
+Das \emph{et al.} in \cite{Das2011}, \cite{Das2013}, \cite{Das2014}, \cite{Das2015} segmented erythrocytes as aforesaid and then morphological operators are used to eliminate unwanted cells like leukocytes and platelets.
+To conclude, overlapping erythrocytes are segmented by using marker controlled watershed segmentation technique.
 
-Several authors used marker controlled watershed with morphological approach.
-Khan et al.
-\cite{Khan2011}, among other experimentations, used it in order to try to separate overlapping cells because, according to their statements, watershed transform can separate touching cells but it is not sufficient for overlapping cells.
+In the paper \cite{Devi2017} Devi \emph{et al.} proposed a computer assisted system for quantification of erythrocytes in microscopic images of thin blood smears.
+The performance of the system in classifying the isolated and clump erythrocytes by geometric features is evaluated for the different classifiers.
+The clump erythrocytes are segmented using marker controlled watershed with h-minima as internal marker.
 
-In \cite{Springl2009} red blood cell segmentation has been performed by using marker-controlled watershed transformation based on the image gradient.
+In \cite{Dey2015} Dey \emph{et al.} presented an automatic system for segmenting platelets, useful for identifying disease as malaria, using a color based segmentation and mathematical morphology (opening operations with a disk element of radius 2).
+
+In the study presented in \cite{Diaz2009} by Diaz \emph{et al.} for quantification and classification of erythrocytes in stained thin blood films infected with Plasmodium falciparum, the authors used connected morphological operators in the segmentation step.
+The RBCs are detected as follows: firstly, a pixel classification allowed to label each image pixel as either background or foreground, based on its color features.
+Afterward, an inclusion-tree structure represented the hierarchical object relations between background and foreground so that a filtering process allows to remove irrelevant structures such as artifacts generated at the staining or digitization processes.
+
+Khan \emph{et al.} \cite{Khan2011}, among other experimentations, used it in order to try to separate overlapping cells because, according to their statements, watershed transform can separate touching cells but it is not sufficient for overlapping cells.
+
+In the algorithm described by Romero-Rondon \emph{et al.} in \cite{Romero2016} the detection of overlapped RBCs is still based on marker-controlled watershed transform.
+To define the suitable markers in watershed transform they used three different approaches, based on a morphological erosion operation, on Hough transform and on clustering method of k-means.
+
+Savkare \emph{et al.} in \cite{Savkare2015} segmented cells using k-mean clustering and global threshold.
+Overlapping cells are separated using Sobel edge detector and watershed transform.
+Watershed transform is applied on each cluster separately.
+Over-segmentation is minimized by series of morphological operations, like erosion and dilation utilizing disk shaped structuring element.
+
+In \cite{Savkare2011a} an approach  to detect red blood cells with consecutive classification into parasite infected and normal cells for further estimation of parasitemia is proposed.
+For separation of overlapping cells watershed transform is applied on distance transform of binary mask of cells having larger area.
+
+In \cite{Springl2009} {\v{S}}pringl performed red blood cell segmentation by using marker-controlled watershed transformation based on the image gradient.
 Markers are computed as a combination of the binary mask of the red blood cells and centres of the cells which are computed using a similar algorithm that was utilized for the evaluation of the average cell radius.
@@
 
-Das et al.
-\cite{Das2011}, \cite{Das2013}, \cite{Das2014}, \cite{Das2015} segmented erythrocytes as aforesaid and then morphological operators are used to eliminate unwanted cells like leukocytes and platelets.
-To conclude, overlapping erythrocytes are segmented by using marker controlled watershed segmentation technique.
+In \cite{Sulist2015} Sulistyawati \emph{et al.} combined morphological operations (erosion, dilation, opening and closing) and blob analysis to segment and identify malaria parasites with a high degree of accuracy.\\
 
-In the paper \cite{Devi2017} the authors propose a computer assisted system for quantification of erythrocytes in microscopic images of thin blood smears.
-%The proposed method consists of preprocessing, segmentation, morphological filtering, cell separation and clump cell segmentation.
-The performance of the system in classifying the isolated and clump erythrocytes by geometric features is evaluated for the different classifiers.
-The clump erythrocytes are segmented using marker controlled watershed with h-minima as internal marker.
-%Based on the experimental results, it may be concluded that the proposed model provides satisfactory results with an accuracy of 98.02\% in comparison to the state of art.
+Tek \emph{et al.} in \cite{Tek2006} proposed a classifier-based method, for the segmentation stage, which relies on a Bayesian pixel classifier to distinguish among stained and non-stained pixels.
+In particular, they used a non-parametric method based on histograms in order to produce the probability density functions of stained and non-stained classes.
+Stained pixels can belong to other components such as WBCs, platelets or artefacts, in addition to the parasites and so the detection procedure requires a further classification to distinguish among parasite and non-parasite pixels.
+However, the stained pixels have to be represented as connected sets, representing stained objects, to extract features for the classifier.
+Furthermore, top-hat extraction and infinite reconstruction were applied to find the regions that include the objects.
 
-%
+To conclude many systems for computer aided diagnosis of malaria disease made use of mathematical operations in order to smoothen the boundary of the regions obtained from the segmentation process.
+
 %%BOUNDARY BASED METHODS \subsubsection{Boundary based methods}
@@
 
-In the algorithm described by Romero-Rondon et al.
-in \cite{Romero2016} the detection of overlapped RBCs is based on marker-controlled watershed transform.
-To define the suitable markers in watershed transform they used three different approaches, based on a morphological erosion operation, on Hough transform and on clustering method of k-means.
-
-Sovkare in \cite{Savkare2015} segmented cells using k-mean clustering and global threshold.
-Overlapping cells are separated using Sobel edge detector and watershed transform.
-Watershed transform is applied on each cluster separately.
-Over-segmentation is minimized by series of morphological operations, like erosion and dilation utilizing disk shaped structuring element.
-
-In \cite{Savkare2011a} an approach  to detect red blood cells with consecutive classification into parasite infected and normal cells for further estimation of parasitemia is  proposed.
-For separation of overlapping cells watershed transform is applied on distance transform of binary mask of cells having larger area.
-
-Tek et al.
-\cite{Tek2006} proposed a classifier-based method, for the segmentation stage, which relies on a Bayesian pixel classifier to distinguish among stained and non-stained pixels.
-In particular, they used a non-parametric method based on histograms in order to produce the probability density functions of stained and non-stained classes.
-Stained pixels can belong to other components such as WBCs, platelets or artefacts, in addition to the parasites and so the detection procedure requires a further classification to distinguish among parasite and non-parasite pixels.
-However, the stained pixels have to be represented as connected sets, representing stained objects, to extract features for the classifier.
-Furthermore, top-hat extraction and infinite reconstruction were applied to find the regions that include the objects.
-
-In \cite{Dey2015} the authors proposed an automatic system for segmenting platelets, useful for identifying disease as malaria, using a color based segmentation and mathematical morphology (opening operations with a disk element of radius 2).
-
-In \cite{Sulist2015} morphological operations (erosion, dilation, opening and closing) and blob analysis have been combined to segment and identify malaria parasites with a high degree of accuracy.
-
-In the study presented in \cite{Diaz2009} for quantification and classification of erythrocytes in stained thin blood films infected with Plasmodium falciparum, the authors used connected morphological operators in the segmentation step.
-The RBCs are detected as follows: firstly, a pixel classification allowed to label each image pixel as either background or foreground, based on its color features.
-Afterward, an inclusion-tree structure represented the hierarchical object relations between background and foreground so that a filtering process
-allows to remove irrelevant structures such as artifacts generated at the staining or digitization processes.
-
-To conlude many systems for computer aided diagnosis of malaria disease made use of mathematical operatations in order to smoothen the boundary of the regions obtained from the segmentation process.
-
-%
 \subsection{Feature extraction}
@@
 
-Pre-processing phase is typically taken on with filters and the most used in the analysed works is certainly the median filter which permits to preserve sharp edges.
+Preprocessing phase is typically taken on with filters and the most used in the analysed works is certainly the median filter which permits to preserve sharp edges.
 Apart from the classic histogram equalization and contrast stretching techniques, other filters have been employed, e.g., geometric mean filter to remove Gaussian noise preserving edges, Laplacian filter, in order to find edges, and so on.
 Median filter has been found to be effective for reducing impulse noises from the microscopic images, even though recent studies have shown that geometric mean filter provides better performance than the median filter \cite{Das2013}, \cite{Das2015}.
-However, morphological operators have been greatly used with successful performances, imposing themselves as powerful alternatives to more common and used techniques for image enhancement and noise filtering \cite{Oliveira2017}, \cite{Reni2015}, \cite{Romero2016},  \cite{Sheik2013}, \cite{Somasekar2015}, \cite{Tek2010}.
+However, morphological operators have been greatly used with successful performances, imposing themselves as powerful alternatives to more common and used techniques for image enhancement and noise filtering (\cite{DiRuberto2002}, \cite{Gonzalez2016}, \cite{Kareem2011}, \cite{Kareem2012}, \cite{Malihi2013}, \cite{Mushabe2013}, \cite{Oliveira2017}, \cite{Reni2015}, \cite{Romero2016}, \cite{Ross2006} \cite{Sheik2013}, \cite{Somasekar2015}, \cite{Springl2009}, \cite{Tek2010}).
 
@@
 
-Other authors (\cite{Ahirwar2012}, \cite{DiRuberto2002}, \cite{Soni2011}, \cite{Tek2010}) employed granulometry and stated that it is very effective to segment cells with regular size.
+Other authors (\cite{Ahirwar2012}, \cite{DiRuberto2002}, \cite{Khan2011}, \cite{Mushabe2013}, \cite{Ross2006}, \cite{Soni2011}, \cite{Tek2010}) employed granulometry and stated that it is very effective to segment cells with regular size.
 
@@
 
-The computer vision methodologies reported in the literature are based on light microscopic images of human peripheral blood smears for computer-aided detection of malaria parasites and their several stages.
+The computer vision methodologies reported in the literature are based on light microscopic images of human peripheral blood smears for computer-aided detection of malaria parasites and their different life stages.
 Image preprocessing, segmentation of erythrocytes and parasites, malaria parasite feature extraction, malaria detection techniques have been discussed here.
 
-It is worth to remember that mathematical morphology techniques have been widely used for image processing purposes.
+It is worth noticing that cells colours and the colour contrast between cells and background can vary so often according to the different, existing staining techniques, thickness of smear, microscope illumination and microscope's image acquisition procedure, as shown in fig.
+\ref{fig:images_types}.
+A standardization of the procedure should be really useful to avoid superfluous differences in similar images' features and to have fair comparisons among the several proposed methods.
+The main efforts towards the realization of a fully automatic blood cells segmentation and classification system cannot leave this aspect out.
+
+Mathematical morphology techniques have been widely used for image processing purposes.
 Among the application fields, it has been applied for fingerprint feature extraction, recognition of handwritten digits, license plate detection, border extraction, denoising using morphological filters, text extraction and detection of imperfection in printed circuit boards \cite{Kaur2013}.
-Apart from this kind of fields, mathematical morphology has been employed in biomedical image analysis, especially in preprocessing and segmentation techniques.
+Apart from this kind of fields, mathematical morphology has been employed successfully in biomedical image analysis, especially in preprocessing and segmentation techniques.
 
-Morphological cell analysis is used to face off abnormality identification and classification, early cancer detection. It has been integrated in new methods for biomedical applications,
-such as automatic segmentation and analysis of histological tumour sections, boundary detection of cervical cell nuclei considering overlapping and clustering, the granules segmentation and spatial distribution analysis, morphological characteristics analysis of specific biomedical cells, understanding the chemotactic response and drug influences, or identifying cell morphogenesis in different cell cycle progression.
+Morphological cell analysis is used to face off abnormality identification and classification, early cancer detection.
+It has been integrated in new methods for biomedical applications, such as automatic segmentation and analysis of histological tumour sections, boundary detection of cervical cell nuclei considering overlapping and clustering, the granules segmentation and spatial distribution analysis, morphological characteristics analysis of specific biomedical cells, understanding the chemotactic response and drug influences, or identifying cell morphogenesis in different cell cycle progression.
 Morphological feature quantification for grading cancerous or precancerous cells is especially widely researched in the literature, such as nuclei segmentation based on marker-controlled watershed transform and snake model for hepatocellular carcinoma feature extraction and classification, which is important for prognosis and treatment planning, nuclei feature quantification for cancer cell cycle analysis, and using feature extraction including image morphological analysis, wavelet analysis, and texture analysis for automated classification of renal cell \cite{Chen2012}.
@@
 
-It is worth to mention that the development of new mobility-aware microscopic devices (and ideally low cost) is an area that can greatly improve the chances of the successful deployment of computer vision CAD solutions for malaria diagnosis in the field.
+In the end, it is worth considering that the development of new mobility-aware microscopic devices (and ideally low cost) is an area that can greatly improve the chances of the successful deployment of computer vision CAD solutions for malaria diagnosis in the field.
 The mobile phone is currently Africa's most important digital technology, and is boosting African health as it emerges as a platform for diagnosis and treatment.
@@
 Moreover, the lack of highly trained microscopists on malaria diagnosis in rural areas could then be complemented by a significantly less specialized technician that knows how to operate the system and prepare blood smears.
-The usage of mobile devices in the system architecture can also bring significant improvements in terms of portability and data transmission.
-Finally, malaria diagnosis might be just one
-element of a suite of diagnostic software tests running on this type of system.
+The usage of mobile devices in the system architecture can also bring significant improvements in terms of portability and data transmission, like the systems proposed by \cite{Oliveira2017} and \cite{Rosado2017}.
+Finally, malaria diagnosis might be just one element of a suite of diagnostic software tests running on this type of system.
 Several other tests could simultaneously be carried out using the same images, for instance cell counting or detection of other hemoparasites like microfilaria or trypanosoma \cite{Rosado2016}.

### 2017-11-24 Table_sum.tex
Updated conclusion and named-ordered authors in pre-proc and segmentation sections

@@
     Di Ruberto \emph{et al.}, 2002 &
-		area closing &
+		area closing, opening &
 		thresholding + granulometry, watershed  transform &	
@@
     	morphological filter, erosion-reconstruction, dilation-reconstruction, closing &
-    	-  &
+    	watershed  transform  &
     	- &
@@
 
-    Kareem \emph{et al.}, 2012  &
+    Kareem \emph{et al.}, 2011, 2012  &
     	dilation, erosion &
@@
     	Acc=81\% SE=85.5\% \vspace{0.6cm} \\
-
-    Khan \emph{et al.}, 2014  &
-    	- &
-    	- &
-    	- &
-    	two (infected and noninfected) &
-    	SE=93\% SP=95\% \vspace{0.6cm} \\
 
@@
     Romero-Rondon \emph{et al.}, 2016  &
-    	opening &
+    	dilation, opening &
     	marker controlled watershed, erosion  &
@@
     Sheikhhosseini \emph{et al.}, 2013  &
-    	- &
+    	hole filling &
     	thresholding + hole filling, opening  &
@@
     Somasekar \emph{et al.}, 2015  &
-    	- &
+    	erosion &
     	fuzzy C-means clustering + erosion, hole filling  &

### 2017-11-24 MDPI_template/template.tex
added new template for sensors journal

%  LaTeX support: latex@mdpi.com 
%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).

% You need to save the "mdpi.cls" and "mdpi.bst" files into the same folder as this template file.

%=================================================================
\documentclass[journal,article,submit,moreauthors,pdftex,10pt,a4paper]{mdpi} 
%
%--------------------
% Class Options:
%--------------------
% journal
%----------
% Choose between the following MDPI journals:
% actuators, addictions, admsci, aerospace, agriculture, agronomy, algorithms, animals, antibiotics, antibodies, antioxidants, applsci, arts, asi, atmosphere, atoms, axioms, batteries, bdcc, behavsci, beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsci, buildings, carbon, cancers, catalysts, cells, ceramics, challenges, chemengineering, chemosensors, children, chromatography, climate, coatings, colloids, computation, computers, condensedmatter, cosmetics, cryptography, crystals, cybersecurity, data, dentistry, designs, diagnostics, diseases, diversity, econometrics, economies, education, electrochemistry, electronics, energies, entropy, environments, epigenomes, est, fermentation, fibers, fire, fishes, fluids, foods, forests, fractalfract, futureinternet, galaxies, games, gastrointestdisord, gels, genealogy, genes, geosciences, geriatrics, hazardousmatters, healthcare, heritage, highthroughput, horticulturae, humanities, hydrology, informatics, information, infrastructures, inorganics, insects, instruments, ijerph, ijfs, ijms, ijgi, ijtpp, inventions, j, jcdd, jcm, jcs, jdb, jfb, jfmk, jimaging, jof, jintelligence, jlpea, jmmp, jmse, jpm, jrfm, jsan, land, languages, laws, life, literature, logistics, lubricants, machines, magnetochemistry, make, marinedrugs, materials, mathematics, mca, mti, medsci, medicines, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, modelling, molbank, molecules, mps, nanomaterials, ncrna, neonatalscreening, nitrogen, nutrients, ohbm, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, plants, plasma, polymers, preprints, proceedings, processes, proteomes, publications, quaternary, qubs, recycling, religions, remotesensing, resources, risks, robotics, safety, scipharm, sensors, separations, sexes, sinusitis, socsci, societies, soils, sports, standards, surgeries, sustainability, symmetry, systems, technologies, toxics, toxins, tropicalmed, universe, urbansci, vaccines, vetsci, vibration, viruses, vision, water, wem
%---------
% article
%---------
% The default type of manuscript is article, but can be replaced by: 
% abstract, addendum, article, benchmark, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, correction, conferenceproceedings, conferencereport, expressionofconcern, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimages, letter, meetingreport, newbookreceived, opinion, obituary, projectreport, reply, reprint, retraction, review, perspective, protocol, shortnote, supfile, technicalnote, viewpoint
% supfile = supplementary materials
% protocol: If you are preparing a "Protocol" paper, please refer to http://www.mdpi.com/journal/mps/instructions for details on its expected structure and content.
%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g. the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.

%=================================================================
\firstpage{1} 
\makeatletter 
\setcounter{page}{\@firstpage} 
\makeatother 
\articlenumber{x}
\doinum{10.3390/------}
\pubvolume{xx}
\pubyear{2017}
\copyrightyear{2017}
\externaleditor{Academic Editor: name}
\history{Received: date; Accepted: date; Published: date}

%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, calc, indentfirst, fancyhdr, graphicx, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, amsthm, hyphenat, natbib, hyperref, footmisc, geometry, caption, url, mdframed, tabto, soul, multirow, microtype, tikz

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Title}

% Author Orchid ID: enter ID or remove command
\newcommand{\orcidauthorA}{0000-0000-000-000X} % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-000-000X} % Add \orcidB{} behind the author's name

% Authors, for the paper (add full first names)
\Author{Firstname Lastname $^{1,\dagger,\ddagger}$\orcidA{}, Firstname Lastname $^{1,\ddagger}$ and Firstname Lastname $^{2,}$*}

% Authors, for metadata in PDF
\AuthorNames{Firstname Lastname, Firstname Lastname and Firstname Lastname}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Affiliation 1; e-mail@e-mail.com\\
$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}

% Contact information of the corresponding author
\corres{Correspondence: e-mail@e-mail.com; Tel.: +x-xxx-xxx-xxxx}

% Current address and/or shared authorship
\firstnote{Current address: Affiliation 3} 
\secondnote{These authors contributed equally to this work.}
% The commands \thirdnote{} till \eighthnote{} are available for further notes

% Simple summary
%\simplesumm{}

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{A single paragraph of about 200 words maximum. For research articles, abstracts should give a pertinent overview of the work. We strongly encourage authors to use the following style of structured abstracts, but without headings: 1) Background: Place the question addressed in a broad context and highlight the purpose of the study; 2) Methods: Describe briefly the main methods or treatments applied; 3) Results: Summarize the article's main findings; and 4) Conclusion: Indicate the main conclusions or interpretations. The abstract should be an objective representation of the article, it must not contain results which are not presented and substantiated in the main text and should not exaggerate the main conclusions.}

% Keywords
\keyword{keyword 1; keyword 2; keyword 3 (list three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline.)}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences:
%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data:
%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately. If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal. In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}

%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%\setcounter{secnumdepth}{4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Only for the journal Gels: Please place the Experimental Section after the Conclusions

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setcounter{section}{-1} %% Remove this when starting to work on the template.
\section{How to Use this Template}
The template details the sections that can be used in a manuscript. Note that the order and names of article sections may differ from the requirements of the journal (e.g. the positioning of the Materials and Methods section). Please check the instructions for authors page of the journal to verify the correct order and names. For any questions, please contact the editorial office of the journal or support@mdpi.com. For LaTeX related questions please contact Janine Daum at latex-support@mdpi.com.

\section{Introduction}

The introduction should briefly place the study in a broad context and highlight why it is important. It should define the purpose of the work and its significance. The current state of the research field should be reviewed carefully and key publications cited. Please highlight controversial and diverging hypotheses when necessary. Finally, briefly mention the main aim of the work and highlight the principal conclusions. As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research. Citing a journal paper \cite{ref-journal}. And now citing a book reference \cite{ref-book}. Please use the command \citep{ref-journal} for the following MDPI journals, which use author-date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

This section may be divided by subheadings. It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Subsection}

\subsubsection{Subsubsection}

Bulleted lists look like this:
\begin{itemize}[leftmargin=*,labelsep=5.8mm]
\item	First bullet
\item	Second bullet
\item	Third bullet
\end{itemize}

Numbered lists can be added as follows:
\begin{enumerate}[leftmargin=*,labelsep=4.9mm]
\item	First item 
\item	Second item
\item	Third item
\end{enumerate}

The text continues here.

\subsection{Figures, Tables and Schemes}

All figures and tables should be cited in the main text as Figure 1, Table 1, etc.

\begin{figure}[H]
\centering
\includegraphics[width=3 cm]{logo-mdpi}
\caption{This is a figure, Schemes follow the same formatting. If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel. (\textbf{b}) Description of what is contained in the second panel. Figures should be placed in the main text near to the first time they are cited. A caption on a single line should be centered.}
\end{figure}

\begin{table}[H]
\caption{This is a table caption. Tables should be placed in the main text near to the first time they are cited.}
\centering
%% \tablesize{} %% You can specify the fontsize here, e.g.  \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{ccc}
\toprule
\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
\midrule
entry 1		& data			& data\\
entry 2		& data			& data\\
\bottomrule
\end{tabular}
\end{table}

\subsection{Formatting of Mathematical Components}

This is an example of an equation:

\begin{equation}
\mathbb{S}
\end{equation}

%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments). This ensures correct linenumbering. The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph. 
Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
%% Example of a theorem:
\begin{Theorem}
Example text of a theorem.
\end{Theorem}

The text continues here. Proofs must be formatted as follows:

%% Example of a proof:
\begin{proof}[Proof of Theorem 1]
Text of the proof. Note that the phrase `of Theorem 1' is optional if it is clear which theorem is being referred to.
\end{proof}
The text continues here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

Authors should discuss the results and how they can be interpreted in perspective of previous studies and of the working hypotheses. The findings and their implications should be discussed in the broadest context possible. Future research directions may also be highlighted.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and Methods}

Materials and Methods should be described with sufficient details to allow others to replicate and build on published results. Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers. Please disclose at the submission stage any restrictions on the availability of materials or information. New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.

Research manuscripts reporting large datasets that are deposited in a publicly available database should specify where the data have been deposited and provide the relevant accession numbers. If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review. They must be provided prior to publication.

Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Patents}
This section is not mandatory, but may be added if there are patents resulting from the work reported in this manuscript.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\supplementary{The following are available online at www.mdpi.com/link, Figure S1: title, Table S1: title, Video S1: title.}

% Only for the journal Methods and Protocols:
% If you wish to submit a video article, please do so with any other supplementary material.
% \supplementary{The following are available at www.mdpi.com/link: Figure S1: title, Table S1: title, Video S1: title. A supporting video article is available at doi: link.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{All sources of funding of the study should be disclosed. Please clearly indicate grants that you have received in support of your research work. Clearly state if you received funds for covering the costs to publish in open access.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided. The following statements should be used ``X.X. and Y.Y. conceived and designed the experiments; X.X. performed the experiments; X.X. and Y.Y. analyzed the data; W.W. contributed reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be limited to those who have contributed substantially to the work reported.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\conflictsofinterest{Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results. Any role of the funding sponsors in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript, or in the decision to publish the results must be declared in this section. If there is no role, please state ``The founding sponsors had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, and in the decision to publish the results''.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\abbreviations{The following abbreviations are used in this manuscript:\\

\noindent 
\begin{tabular}{@{}ll}
MDPI & Multidisciplinary Digital Publishing Institute\\
DOAJ & Directory of open access journals\\
TLA & Three letter acronym\\
LD & linear dichroism
\end{tabular}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
\appendixsections{multiple} %Leave argument "multiple" if there are multiple sections. Then a counter is printed ("Appendix A"). If there is only one appendix section then change the argument to "one" and no counter is printed ("Appendix").
\appendix
\section{}
\subsection{}
The appendix is an optional section that can contain details and data supplemental to the main text. For example, explanations of experimental details that would disrupt the flow of the main text, but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data is shown in the main text can be added here if brief, or as Supplementary data. Mathematical proofs of results not central to the paper can be added as an appendix.

\section{}
All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.

%=====================================
% References, variant A: internal bibliography
%=====================================
\reftitle{References}
\begin{thebibliography}{999}
% Reference 1
\bibitem[Author1(year)]{ref-journal}
Author1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142-149, DOI.
% Reference 2
\bibitem[Author2(year)]{ref-book}
Author2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32-58, ISBN.
\end{thebibliography}

% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences. For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p. 328; \citeyear{ref-journal-3b}, p.475). This produces: Wong (1999, p. 328; 2000, p. 475)

%=====================================
% References, variant B: external bibliography
%=====================================
%\externalbibliography{yes}
%\bibliography{your_external_BibTeX_file}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\sampleavailability{Samples of the compounds ...... are available from the authors.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}

### 2017-11-30 MDPI_template/paper.tex
Template changed

--- 2017-11-24 Paper.tex
@@
-\documentclass[10pt,a4paper]{article}
-\usepackage[latin1]{inputenc}
+%  LaTeX support: latex@mdpi.com 
+%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).
+
+% You need to save the "mdpi.cls" and "mdpi.bst" files into the same folder as this template file.
+
+%=================================================================
+\documentclass[sensors,review,submit,moreauthors,pdftex,10pt,a4paper]{mdpi}
+
 \usepackage{amsmath}
@@
 
-% for institute information	
-\usepackage{authblk}
-
 % definition of keywords
-\def\keywords{\vspace{.5em}
-	{\bfseries\textit{Keywords}---\,\relax%
-}}
+\def\keywords{\vspace{.5em}{\bfseries\textit{Keywords}---\,\relax}}
 \def\endkeywords{\par}
@@
 \usepackage{longtable}
-
+\usepackage{multirow}
 \usepackage{fancyhdr}
 \usepackage{booktabs}
-\usepackage{cite}
-\usepackage{graphicx}
 \usepackage{array}
-\usepackage{multirow}
-\usepackage{amssymb}
-\usepackage{fancyhdr}
-\usepackage{booktabs}
-\usepackage{amsmath}
+
 \usepackage{float}
+\usepackage{filecontents}
+
 \usepackage[tight,footnotesize]{subfigure}
@@
 
-\usepackage{filecontents}
-\usepackage{cite}
+%
+%--------------------
+% Class Options:
+%--------------------
+% journal
+%----------
+% Choose between the following MDPI journals:
+% actuators, addictions, admsci, aerospace, agriculture, agronomy, algorithms, animals, antibiotics, antibodies, antioxidants, applsci, arts, asi, atmosphere, atoms, axioms, batteries, bdcc, behavsci, beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsci, buildings, carbon, cancers, catalysts, cells, ceramics, challenges, chemengineering, chemosensors, children, chromatography, climate, coatings, colloids, computation, computers, condensedmatter, cosmetics, cryptography, crystals, cybersecurity, data, dentistry, designs, diagnostics, diseases, diversity, econometrics, economies, education, electrochemistry, electronics, energies, entropy, environments, epigenomes, est, fermentation, fibers, fire, fishes, fluids, foods, forests, fractalfract, futureinternet, galaxies, games, gastrointestdisord, gels, genealogy, genes, geosciences, geriatrics, hazardousmatters, healthcare, heritage, highthroughput, horticulturae, humanities, hydrology, informatics, information, infrastructures, inorganics, insects, instruments, ijerph, ijfs, ijms, ijgi, ijtpp, inventions, j, jcdd, jcm, jcs, jdb, jfb, jfmk, jimaging, jof, jintelligence, jlpea, jmmp, jmse, jpm, jrfm, jsan, land, languages, laws, life, literature, logistics, lubricants, machines, magnetochemistry, make, marinedrugs, materials, mathematics, mca, mti, medsci, medicines, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, modelling, molbank, molecules, mps, nanomaterials, ncrna, neonatalscreening, nitrogen, nutrients, ohbm, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, plants, plasma, polymers, preprints, proceedings, processes, proteomes, publications, quaternary, qubs, recycling, religions, remotesensing, resources, risks, robotics, safety, scipharm, sensors, separations, sexes, sinusitis, socsci, societies, soils, sports, standards, surgeries, sustainability, symmetry, systems, technologies, toxics, toxins, tropicalmed, universe, urbansci, vaccines, vetsci, vibration, viruses, vision, water, wem
+%---------
+% article
+%---------
+% The default type of manuscript is article, but can be replaced by: 
+% abstract, addendum, article, benchmark, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, correction, conferenceproceedings, conferencereport, expressionofconcern, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimages, letter, meetingreport, newbookreceived, opinion, obituary, projectreport, reply, reprint, retraction, review, perspective, protocol, shortnote, supfile, technicalnote, viewpoint
+% supfile = supplementary materials
+% protocol: If you are preparing a "Protocol" paper, please refer to http://www.mdpi.com/journal/mps/instructions for details on its expected structure and content.
+%----------
+% submit
+%----------
+% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted.
+This will only make changes to the frontpage (e.g.
+the logo of the journal will get visible), the headings, and the copyright information.
+Also, line numbering will be removed.
+Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
+%------------------
+% moreauthors
+%------------------
+% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
+%---------
+% pdftex
+%---------
+% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.
+
+%=================================================================
+\firstpage{1} 
+\makeatletter 
+\setcounter{page}{\@firstpage} 
+\makeatother 
+\articlenumber{x}
+\doinum{10.3390/------}
+\pubvolume{xx}
+\pubyear{2017}
+\copyrightyear{2017}
+\externaleditor{Academic Editor: name}
+\history{Received: date; Accepted: date; Published: date}
+
+%------------------------------------------------------------------
+% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
+%\pdfoutput=1
+
+%=================================================================
+% Add packages and commands here.
+The following packages are loaded in our class file: fontenc, calc, indentfirst, fancyhdr, graphicx, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, amsthm, hyphenat, natbib, hyperref, footmisc, geometry, caption, url, mdframed, tabto, soul, multirow, microtype, tikz
+
+%=================================================================
+%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition
+%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).
+
+%=================================================================
+% Full title of the paper (Capitalized)
+\Title{Recent advances of malaria parasites detection systems based on mathematical morphology}
+
+% Author Orchid ID: enter ID or remove command
+\newcommand{\orcidauthorA}{0000-0000-000-000X} % Add \orcidA{} behind the author's name
+%\newcommand{\orcidauthorB}{0000-0000-000-000X} % Add \orcidB{} behind the author's name
+
+% Authors, for the paper (add full first names)
+\Author{Firstname Lastname $^{1,\dagger,\ddagger}$\orcidA{}, Firstname Lastname $^{1,\ddagger}$ and Firstname Lastname $^{2,}$*}
+
+% Authors, for metadata in PDF
+\AuthorNames{Firstname Lastname, Firstname Lastname and Firstname Lastname}
+
+% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
+\address{%
+$^{1}$ \quad Affiliation 1; e-mail@e-mail.com\\
+$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}
+
+% Contact information of the corresponding author
+\corres{Correspondence: e-mail@e-mail.com; Tel.: +x-xxx-xxx-xxxx}
+
+% Current address and/or shared authorship
+\firstnote{Current address: Affiliation 3} 
+\secondnote{These authors contributed equally to this work.}
+% The commands \thirdnote{} till \eighthnote{} are available for further notes
+
+% Simple summary
+%\simplesumm{}
+
+% Abstract (Do not insert blank lines, i.e. \\) 
+\abstract{This paper investigates existing mathematical morphology based techniques applied for performing malaria parasites detection and identification in both Giemsa and Leishman stained blood smears images.
+Malaria is an epidemic health disease and a rapid, accurate diagnosis is necessary for proper intervention.
+Generally, pathologists visually examine blood stained slides for malaria diagnosis; this kind of visual inspection is subjective, error-prone and time consuming.
+In order to cope with such issues, computer-aided methods have been increasingly evolved for abnormal erythrocyte and/or parasites detection, segmentation and semi/fully automated classification.
+The aim of this paper is to present a review of mathematical morphology based methods for malaria parasite detection.}
+
+% Keywords
+\keyword{malaria, red blood cells segmentation, mathematical morphology, medical image analysis}
+
+% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
+%\PACS{J0101}
+%\MSC{}
+%\JEL{}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Applied Sciences:
+%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Data:
+%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately.
+If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal.
+In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}
+
+%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}
+
+%\setcounter{secnumdepth}{4}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \begin{document}
-\title{Recent advances of malaria parasites detection systems based on mathematical morphology}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Only for the journal Gels: Please place the Experimental Section after the Conclusions
 
-\author[1]{Andrea Loddo}
-\author[1]{Cecilia Di Ruberto}
-\author[2]{Michel Kocher}
-
-\affil[1]{Department of Mathematics and Computer Science\\	Via Ospedale 72, 09124, Cagliari, Italy}
-\affil[ ]{\textit {\{andrea.loddo, dirubert\}@unica.it}}
-\affil[2]{Biomedical Imaging Group, École Polytechnique
-	Fédérale de Lausanne (EPFL), CH-1015 Lausanne, Switzerland}
-\affil[ ]{\textit {michel.kocher@heig-vd.ch}}
-\date{}                     %% if you don't need date to appear
-\setcounter{Maxaffil}{0}
-\renewcommand\Affilfont{\itshape\small}
-
-\maketitle
-
-\begin{abstract}
-	This paper investigates existing mathematical morphology based techniques applied for performing malaria parasites detection and identification in both Giemsa and Leishman stained blood smears images.
-	Malaria is an epidemic health disease and a rapid, accurate diagnosis is necessary for proper intervention.
-Generally, pathologists visually examine blood stained slides for malaria diagnosis; this kind of visual inspection is subjective, error-prone and time consuming.
-In order to cope with such issues, computer-aided methods have been increasingly evolved for abnormal erythrocyte detection, segmentation and semi/fully automated classification.
-The aim of this paper is to present a review on the methods that are based on mathematical morphology techniques for malaria parasite detection.
-
-	\keywords{malaria, red blood cells segmentation, mathematical morphology, medical image analysis}
-\end{abstract}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% d
 
 \section{Introduction}
+
 Haematology is the branch of medicine concerned with the study, diagnosis, monitoring, treatment, and prevention of diseases related to the blood and blood-forming organs.
@@
 The use of image processing techniques can help to analyse, count the cells in human blood and, at the same time, to provide useful and precise information about cells morphology.
-
 Peripheral blood smears analysis is a common and economical diagnosis technique by which expert pathologists may obtain health information about the patients.
@@
 It is worth considering that blood cells exist with different shapes, characteristics and colourations, according to their types.
-
 Many tests are designed to determine the number of erythrocytes and leukocytes in the blood, together with the volume, sedimentation rate, and haemoglobin concentration of the red blood cells (blood count).
@@
 There are special tests based on the presence in the blood of substances characteristic of specific infections, such as the serological tests for syphilis, hepatitis, and human immunodeficiency virus (HIV, the AIDS virus) \footnote{https://www.britannica.com/topic/blood-analysis}.
-
 Among the several available blood tests, the most common are certainly the blood cells counts, e.g., a CBC is a measure of the haematologic parameters of the blood.
@@
 	\centering	
-	\includegraphics[width=5cm]{img/f1a.eps}
-	\includegraphics[width=5cm]{img/f1b.eps}
-	\includegraphics[width=5cm]{img/f1c.eps}
-	\includegraphics[width=5cm]{img/f1d.eps}
+	\includegraphics[width=3.5cm]{img/f1a.eps}
+	\includegraphics[width=3.5cm]{img/f1b.eps}
+	\includegraphics[width=3.5cm]{img/f1c.eps}
+	\includegraphics[width=3.5cm]{img/f1d.eps}
 	\caption{\label{fig:images_types}Different illumination conditions generate different images, because of the absence of a standardized acquisition procedure.
@@
 	\centering
-	\includegraphics[width=3.5cm]{img/f3_Pfalciparum}
-	\includegraphics[width=3.5cm]{img/f3_Pvivax}
-	\includegraphics[width=3.5cm]{img/f3_Povale}
+	\includegraphics[width=4.5cm]{img/f3_Pfalciparum}
+	\includegraphics[width=4.5cm]{img/f3_Pvivax}
+	\includegraphics[width=4.5cm]{img/f3_Povale}
 	%	\includegraphics[width=2.75cm]{img/f3_Pmalariae}
@@
 	\centering
-	\includegraphics[width=2.75cm, height=2.75cm]{img/falciparum_1_ring}
-	\includegraphics[width=2.75cm, height=2.75cm]{img/falciparum_2_trophozoiteAge}
-	\includegraphics[width=2.75cm, height=2.75cm]{img/falciparum_3_schizont}
-	\includegraphics[width=2.75cm, height=2.75cm]{img/falciparum_4_gametocyte}
-	\includegraphics[width=2.75cm, height=2.75cm]{img/ovale_1_ring}
-	\includegraphics[width=2.75cm, height=2.75cm]{img/ovale_2_trophozoite}
-	\includegraphics[width=2.75cm, height=2.75cm]{img/ovale_3_schizont}
-	\includegraphics[width=2.75cm, height=2.75cm]{img/ovale_4_gametocyte}
-	\includegraphics[width=2.75cm, height=2.75cm]{img/vivax_1_ring}
-	\includegraphics[width=2.75cm, height=2.75cm]{img/vivax_2c_trophozoiteDeveloped}
-	\includegraphics[width=2.75cm, height=2.75cm]{img/vivax_4_gametocyte}
+	\includegraphics[width=2.5cm, height=2.5cm]{img/falciparum_1_ring}
+	\includegraphics[width=2.5cm, height=2.5cm]{img/falciparum_2_trophozoiteAge}
+	\includegraphics[width=2.5cm, height=2.5cm]{img/falciparum_3_schizont}
+	\includegraphics[width=2.5cm, height=2.5cm]{img/falciparum_4_gametocyte}
+	\includegraphics[width=2.5cm, height=2.5cm]{img/ovale_1_ring}
+	\includegraphics[width=2.5cm, height=2.5cm]{img/ovale_2_trophozoite}
+	\includegraphics[width=2.5cm, height=2.5cm]{img/ovale_3_schizont}
+	\includegraphics[width=2.5cm, height=2.5cm]{img/ovale_4_gametocyte}
+	\includegraphics[width=2.5cm, height=2.5cm]{img/vivax_1_ring}
+	\includegraphics[width=2.5cm, height=2.5cm]{img/vivax_2c_trophozoiteDeveloped}
+	\includegraphics[width=2.5cm, height=2.5cm]{img/vivax_4_gametocyte}
 	\caption{\label{fig:malaria_stages}Examples of malaria parasite stages. From top left: P.falciparum ring, trophozoite, schizont, gametocyte;
@@
 Some of the salient points regarding the morphological approach are as follows \cite{Giardina1988}:
-\begin{enumerate}
+\begin{itemize}[leftmargin=*,labelsep=5.8mm]
 	\item Morphological operations provide for the systematic alteration of the geometric content of an image while maintaining the stability of the important geometric characteristics.
@@
 	\item There exist rigorous representations theorems by means of which one can obtain the expression of morphological filters in terms of the primitive morphological operations.
-\end{enumerate}
+\end{itemize}
 Dilation and erosion are the basic morphological processing operations.
@@
 Section 2 presents a review of the considered works, according to a typical pipeline of a computer-aided image analysis process: preprocessing, segmentation, feature extraction.
-%introduces the adopted materials and methods for malaria-infected peripheral blood cells images analysis.
-It has been structured in preprocessing, segmentation, feature extraction, feature selection and classification subsections, consistently with a typical pipeline of a computer-aided image analysis process.
+
 All the considered works make use of morphological operators in at least one of the phases of image analysis.
@@
 	\item Feature extraction.
-	%\item Feature selection.
 	\item Classification.
@@
 In image analysis field, especially when we refer to complex computer-aided pipelines, pre-processing methods are particularly used in order to improve the image data by suppressing unwanted noise or enhancing some image features for further processing.
-%Typically, pre-processing methods are classified according to the following schema:
-%\begin{enumerate}
-%	\item pixel brightness transformations;
-%	\item geometric transformations;
-%	\item pre-processing methods that use a local neighbourhood of the processed pixel;
-%	\item image restoration that requires knowledge about the entire image.
-%\end{enumerate}
 It is worth to mention pre-processing methods because they are an important step regarding image analysis field but, for what concerns the malaria-affected blood image analysis, in our review we particularly found methods which operate for illumination correction and noise filtering purposes.
@@
 Different techniques for illumination correction have been suggested to reduce such variation, e.g., a lot of authors work with grayscale-converted images as an illumination correction method. 
-%\cite{Tek2010},\cite{Das2015},\cite{Tek2006}.
 On the other hand, noise filtering aims to remove the noise introduced by mishandling the slides and/or the camera settings.
@@
 The latter case is commonly used in applications in which the cell class depends on the morphological characteristics of its components.
-%The segmentation techniques can generally be divided into three main categories: Pixel Based, Edge Based and Region Based.
-%
-%%\subsection{Erythrocytes segmentation}
-%%\subsubsection{Thresholding}
-%%THRESHOLDING METHODS
 
+%%% THRESHOLDING METHODS
 Several other authors attempted to use thresholding combined with morphological operation as segmentation method in their computer-aided systems and they are described as follows.
@@
 
-%
 %Makkapati and Rao proposed two different methodologies.
@@
 It utilizes the information based on structure, colour and geometry of the cells and does not require any segmentation or non-illumination correction techniques that are commonly used for cell detection.
-%
+
 %Fang et al.
@@
 
-%
 %Malihi et al.
@@
 
-%% Morphology and/or granulometry in Segmentation stage:
+% Morphology and/or granulometry in Segmentation stage:
 On the other hand, a lot of works have been realized by means of mathematical morphology and/or granulometry in the segmentation stages, even in combination with thresholding strategies.
@@
 in \cite{Das2011} (opening, closing) and in \cite{DiRuberto2002} (skeleton) to classify parasites.
-%
-%
-%%\input{Test.tex}
-%
+
+%\input{Test.tex}
+
 \section{Discussions}
@@
 In the review we have only considered the methods which employed mathematical morphology in at least one step of the pipelines and it has been structured by considering the following information: preprocessing, segmentation, features extraction.
-%, classification strategy, specificity and sensitivity percentages relative to the classification steps.
-Most of the studies are based on P. vivax and P. falciparum characterization.
+Most of the studies are based on P. vivax and/or P. falciparum characterization.
 With regards to the showed approaches and related results, it is clear that malaria parasites detection and segmentation techniques in microscopic images needs further experiments and improvements.
@@
 
-\inputencoding{latin2}
-\bibliographystyle{plain}
-\bibliography{main}
-\inputencoding{latin1}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vspace{6pt}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% optional
+\supplementary{The following are available online at www.mdpi.com/link, Figure S1: title, Table S1: title, Video S1: title.}
+
+% Only for the journal Methods and Protocols:
+% If you wish to submit a video article, please do so with any other supplementary material.
+% \supplementary{The following are available at www.mdpi.com/link: Figure S1: title, Table S1: title, Video S1: title. A supporting video article is available at doi: link.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\acknowledgments{All sources of funding of the study should be disclosed.
+Please clearly indicate grants that you have received in support of your research work.
+Clearly state if you received funds for covering the costs to publish in open access.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
+The following statements should be used ``X.X.
+and Y.Y.
+conceived and designed the experiments; X.X.
+performed the experiments; X.X.
+and Y.Y.
+analyzed the data; W.W.
+contributed reagents/materials/analysis tools; Y.Y.
+wrote the paper.'' Authorship must be limited to those who have contributed substantially to the work reported.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\conflictsofinterest{Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results.
+Any role of the funding sponsors in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript, or in the decision to publish the results must be declared in this section.
+If there is no role, please state ``The founding sponsors had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, and in the decision to publish the results''.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% optional
+\abbreviations{The following abbreviations are used in this manuscript:\\
+
+\noindent 
+\begin{tabular}{@{}ll}
+MDPI & Multidisciplinary Digital Publishing Institute\\
+DOAJ & Directory of open access journals\\
+TLA & Three letter acronym\\
+LD & linear dichroism
+\end{tabular}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% optional
+\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
+\appendixsections{multiple} %Leave argument "multiple" if there are multiple sections.
+Then a counter is printed ("Appendix A").
+If there is only one appendix section then change the argument to "one" and no counter is printed ("Appendix").
+\appendix
+\section{}
+\subsection{}
+The appendix is an optional section that can contain details and data supplemental to the main text.
+For example, explanations of experimental details that would disrupt the flow of the main text, but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data is shown in the main text can be added here if brief, or as Supplementary data.
+Mathematical proofs of results not central to the paper can be added as an appendix.
+
+\section{}
+All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.
+
+%=====================================
+% References, variant A: internal bibliography
+%=====================================
+%\reftitle{References}
+%\begin{thebibliography}{999}
+%% Reference 1
+%\bibitem[Author1(year)]{ref-journal}
+%Author1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142-149, DOI.
+%% Reference 2
+%\bibitem[Author2(year)]{ref-book}
+%Author2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32-58, ISBN.
+%\end{thebibliography}
+
+% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences.
+For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
+% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
+% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p.
+328; \citeyear{ref-journal-3b}, p.475).
+This produces: Wong (1999, p.
+328; 2000, p.
+475)
+
+%=====================================
+% References, variant B: external bibliography
+%=====================================
+\externalbibliography{yes}
+\bibliography{bibliography}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% optional
+\sampleavailability{Samples of the compounds ...... are available from the authors.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \end{document}

### 2017-11-30 MDPI_template/paper.tex
Template changed, table updated.

@@
 \usepackage[tight,footnotesize]{subfigure}
+
 \newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
@@
 % Author Orchid ID: enter ID or remove command
-\newcommand{\orcidauthorA}{0000-0000-000-000X} % Add \orcidA{} behind the author's name
+\newcommand{\orcidauthorA}{0000-0002-6571-3816} % Add \orcidA{} behind the author's name
 %\newcommand{\orcidauthorB}{0000-0000-000-000X} % Add \orcidB{} behind the author's name
@@
 % Authors, for the paper (add full first names)
-\Author{Firstname Lastname $^{1,\dagger,\ddagger}$\orcidA{}, Firstname Lastname $^{1,\ddagger}$ and Firstname Lastname $^{2,}$*}
+\Author{Andrea Loddo $^{1,\dagger,\ddagger}$\orcidA{}, Cecilia Di Ruberto $^{1,\ddagger}$ and Michel Kocher$^{2,\ddagger}$}
 
 % Authors, for metadata in PDF
-\AuthorNames{Firstname Lastname, Firstname Lastname and Firstname Lastname}
+\AuthorNames{Andrea Loddo, Cecilia Di Ruberto and Michel Kocher}
 
@@
 \address{%
-$^{1}$ \quad Affiliation 1; e-mail@e-mail.com\\
-$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}
+$^{1}$ \quad Department of Mathematics and Computer Science, University of Cagliari; {andrea.loddo, dirubert}@unica.it\\
+$^{2}$ \quad Biomedical Imaging Group, École Polytechnique Fédérale de Lausanne (EPFL); michel.kocher@heig-vd.ch}
 
 % Contact information of the corresponding author
-\corres{Correspondence: e-mail@e-mail.com; Tel.: +x-xxx-xxx-xxxx}
+\corres{Correspondence: andrea.loddo@unica.it; Tel.: +390706758503}
 
 % Current address and/or shared authorship
-\firstnote{Current address: Affiliation 3} 
+\firstnote{Current address: Via Ospedale 72, 09124, Cagliari, Italy} 
 \secondnote{These authors contributed equally to this work.}
@@
 % Simple summary
-%\simplesumm{}
+\simplesumm{}
 
@@
 %%BOUNDARY BASED METHODS \subsubsection{Boundary based methods}
-%Zou et al.
+Zou et al.
 \cite{Zou2010} based their method on the consideration that RBCs have a circular shape.
@@
 
-\input{Table_sum.tex}
+%\input{Table_sum.tex}
+\input{Table.tex}
 
@@
 %% optional
-\supplementary{The following are available online at www.mdpi.com/link, Figure S1: title, Table S1: title, Video S1: title.}
+%\supplementary{The following are available online at www.mdpi.com/link, Figure S1: title, Table S1: title, Video S1: title.}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\conflictsofinterest{Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results.
-Any role of the funding sponsors in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript, or in the decision to publish the results must be declared in this section.
-If there is no role, please state ``The founding sponsors had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, and in the decision to publish the results''.}
+\conflictsofinterest{The authors declare no conflict of interest.}
 
@@
 \begin{tabular}{@{}ll}
-MDPI & Multidisciplinary Digital Publishing Institute\\
-DOAJ & Directory of open access journals\\
-TLA & Three letter acronym\\
-LD & linear dichroism
+WBC &White Blood Cell \\
+RBC & Red Blood Cell \\
+MP & Malaria Parasite \\
 \end{tabular}}

### 2017-11-30 MDPI_template/Table.tex

--- 2017-11-24 Table_sum.tex
@@
-	\setlength\LTleft{-0.75in}
-%	\setlength\LTright{-1in}
+%\setlength\LTleft{-0.75in}
+%\setlength\LTright{-1in}
 %\small
 \footnotesize
-    \begin{longtable}{C{2.1cm} C{2.1cm} C{2.3cm} C{1.5cm} C{2.7cm} C{2.2cm} }
+\newcolumntype{H}{>{\setbox0=\hbox\bgroup}c<{\egroup}@{}}
+    \begin{longtable}{C{2.1cm} C{2.1cm} C{2.3cm} H C{2.7cm} C{2.2cm} }
     %\begin{longtable}{|C{2cm}|c|C{2.5cm}|C{3cm}|C{2.4cm}|c|}
@@
 		- &
-		marker controlled watershed &	
+		marker controlled watershed + opening, closing &	
 		opening, closing  &
@@
 		area closing, opening &
-		thresholding + granulometry, watershed  transform &	
+		thresholding + granulometry, watershed  transform + skeleton &	
 		skeleton &
@@
     	closing &
-    	-  &
+    	area granulometry  &
     	area granulometry &
@@
     	- &
-    	top-hat, infinite reconstruction  &
+    	top-hat, infinite reconstruction, area granulometry  &
     	area granulometry &
@@
     	closing, granulometry &
-    	thresholding + granulometry, area top-hat, closing  &
+    	thresholding + granulometry, area top-hat, closing, area granulometry &
     	area granulometry &

### 2017-12-07 MDPI_template/paper.tex
Added malariae imgs and courtesy

@@
 In order to cope with such issues, computer-aided methods have been increasingly evolved for abnormal erythrocyte and/or parasites detection, segmentation and semi/fully automated classification.
-The aim of this paper is to present a review of mathematical morphology based methods for malaria parasite detection.}
+The aim of this paper is to present a review of recent mathematical morphology based methods for malaria parasite detection.}
 
@@
 		\centering	
-	\includegraphics[width=3.5cm]{img/f1a.eps}
-	\includegraphics[width=3.5cm]{img/f1b.eps}
-	\includegraphics[width=3.5cm]{img/f1c.eps}
-	\includegraphics[width=3.5cm]{img/f1d.eps}
+		\includegraphics[width=3.5cm]{img/f1a}
+		\includegraphics[width=3.5cm]{img/f1b}
+		\includegraphics[width=3.5cm]{img/f1c}
+		\includegraphics[width=3.5cm]{img/f1d}
 		\caption{\label{fig:images_types}Different illumination conditions generate different images, because of the absence of a standardized acquisition procedure.
-From left to right: acquisition of the same smear with four microscope's brightness levels.}
+From left to right: acquisition of the same smear with four microscope's brightness levels.\newline
+			Courtesy of CHUV, Lausanne.}
 	\end{figure}
@@
 		\centering
-	\includegraphics[width=4.5cm]{img/f3_Pfalciparum}
-	\includegraphics[width=4.5cm]{img/f3_Pvivax}
-	\includegraphics[width=4.5cm]{img/f3_Povale}
-	%	\includegraphics[width=2.75cm]{img/f3_Pmalariae}
+		\includegraphics[width=3.5cm]{img/f2_Pfalciparum}
+		\includegraphics[width=3.5cm]{img/f2_Pvivax}
+		\includegraphics[width=3.5cm]{img/f2_Povale}
+		\includegraphics[width=3.5cm]{img/f2_Pmalariae}
 		\caption{\label{fig:malaria_types}Types of human malaria: from left to right, P.
@@
 Vivax in two gametocytes specimens and one ring stage, P.
-Ovale in its ring stage.}
+Ovale in its ring stage, P.
+Malariae in its schizont stage.\newline
+			Courtesy of CHUV, Lausanne.}
 	\end{figure}
@@
 		\centering
-	\includegraphics[width=2.5cm, height=2.5cm]{img/falciparum_1_ring}
-	\includegraphics[width=2.5cm, height=2.5cm]{img/falciparum_2_trophozoiteAge}
-	\includegraphics[width=2.5cm, height=2.5cm]{img/falciparum_3_schizont}
-	\includegraphics[width=2.5cm, height=2.5cm]{img/falciparum_4_gametocyte}
-	\includegraphics[width=2.5cm, height=2.5cm]{img/ovale_1_ring}
-	\includegraphics[width=2.5cm, height=2.5cm]{img/ovale_2_trophozoite}
-	\includegraphics[width=2.5cm, height=2.5cm]{img/ovale_3_schizont}
-	\includegraphics[width=2.5cm, height=2.5cm]{img/ovale_4_gametocyte}
-	\includegraphics[width=2.5cm, height=2.5cm]{img/vivax_1_ring}
-	\includegraphics[width=2.5cm, height=2.5cm]{img/vivax_2c_trophozoiteDeveloped}
-	\includegraphics[width=2.5cm, height=2.5cm]{img/vivax_4_gametocyte}
-	\caption{\label{fig:malaria_stages}Examples of malaria parasite stages. From top left: P.falciparum ring, trophozoite, schizont, gametocyte;
-		P.ovale ring, trophozoite, schizont, gametocyte; P.vivax ring, developed trophozoite, gametocyte. \cite{Tek2010}}
+		\includegraphics[width=3.5cm, height=3.5cm]{img/falciparum_1_ring}
+		\includegraphics[width=3.5cm, height=3.5cm]{img/falciparum_2_trophozoiteAge}
+		\includegraphics[width=3.5cm, height=3.5cm]{img/falciparum_3_schizont}
+		\includegraphics[width=3.5cm, height=3.5cm]{img/falciparum_4_gametocyte}
+		\includegraphics[width=3.5cm, height=3.5cm]{img/ovale_1_ring}
+		\includegraphics[width=3.5cm, height=3.5cm]{img/ovale_2_trophozoite}
+		\includegraphics[width=3.5cm, height=3.5cm]{img/ovale_3_schizont}
+		\includegraphics[width=3.5cm, height=3.5cm]{img/ovale_4_gametocyte}
+		\includegraphics[width=3.5cm, height=3.5cm]{img/malariae_1_ring}
+		\includegraphics[width=3.5cm, height=3.5cm]{img/malariae_2_trophozoite}
+		\includegraphics[width=3.5cm, height=3.5cm]{img/malariae_3_schizont}
+		\includegraphics[width=3.5cm, height=3.5cm]{img/malariae_4_gametocyte}
+		\includegraphics[width=3.5cm, height=3.5cm]{img/vivax_1_ring}
+		\includegraphics[width=3.5cm, height=3.5cm]{img/vivax_2c_trophozoiteDeveloped}
+		\includegraphics[width=3.5cm, height=3.5cm]{img/vivax_4_gametocyte}
+		\caption{\label{fig:malaria_stages}Examples of malaria parasite stages. 
+			First row, from left to right: P.falciparum ring, trophozoite, schizont, gametocyte;
+			second row, from left to right: P.ovale ring, trophozoite, schizont, gametocyte; 
+			third row, from left to right: P.malariae ring, trophozoite, schizont, gametocyte; 
+			last row, from left to right: P.vivax ring, developed trophozoite, gametocyte.
+			\newline Courtesy of CHUV, Lausanne.}
 	\end{figure}
 	\fi
-
 	\subsection{Mathematical morphology}
@@
 From these two basic operators, others have been derived (opening, closing, hit-or-miss).
-They can be applied to extract image components useful in the representation and descriptions of region shapes, such as boundaries, skeleton, or convex hull.
+They can be applied to extract image components useful in the representation and descriptions of region shapes, such as area granulometry, boundaries, skeleton, or convex hull.
 Also morphological operators can be used for image preprocessing and postprocessing, such as morphological filtering, thinning, and especially for segmentation.
@@
 The system was able to detect the parasites by using an automatic thresholding and morphological operators.
-A morphological approach to cell segmentation which is more efficient than watershed algorithm was proposed.
+A morphological approach to cell segmentation which is more efficient than watershed algorithm \cite{Soille2004} was proposed.
 Finally, the parasites classification was still based on morphological operators.
@@
 	The rest of the paper is organised as follows.
-Section 2 presents a review of the considered works, according to a typical pipeline of a computer-aided image analysis process: preprocessing, segmentation, feature extraction.
-
+Section 3 presents a review of the considered works, according to a typical pipeline of a computer-aided image analysis process: preprocessing, segmentation, feature extraction.
 	All the considered works make use of morphological operators in at least one of the phases of image analysis.
-Section 3 contains an overall discussion about the methods and the conclusions are expressed in section 4.
+Section 4 contains an overall discussion about the methods and the conclusions are expressed in section 5.
 
@@
 At the same time the objects which are not eliminated remain unchanged.
-Also, a morphological closing is performed with a disc structuring element having radius smaller than half the average of the RBCs' radii, in order to connect the possible (more than one) markers that can appear on a single cell.
+Also, a morphological closing is performed with a disk structuring element having radius smaller than half the average of the RBCs radii, in order to connect the possible (more than one) markers that can appear on a single cell.
 
@@
 A dilation is performed by a concentric ring structuring element and erosion by disk shaped structuring element.
-The radius of the structuring element depends on the radius of the RBC, so that all the components smaller than the RBCs can be removed.
+The radius of the structuring element depends on the radius of the RBCs, so that all the components smaller than the RBCs can be removed.
 
@@
 	Di Ruberto \emph{et al.} \cite{DiRuberto2002} aimed to detect the parasites by means of an automatic thresholding based on a morphological approach applied to cell image segmentation, that is more accurate than the classical watershed-based algorithm.
-They have applied grey scale granulometries based on opening with disk-shaped elements, flat and hemispherical.
-They have used a hemispherical disk-shaped structuring element to enhance the roundness and the compactness of the red cells improving the accuracy of the classical watershed algorithm, while they have used a disk-shaped flat structuring element to separate overlapping cells.
+They applied grey scale granulometries based on opening with disk-shaped elements, flat and hemispherical.
+They used a hemispherical disk-shaped structuring element to enhance the roundness and the compactness of the red cells improving the accuracy of the classical watershed algorithm, while they have used a disk-shaped flat structuring element to separate overlapping cells.
 These methods make use of the red blood cell structure knowledge, that is not used in existing watershed-based algorithms.
@@
 
-Several authors used marker controlled watershed with morphological approach, as following described.
+	Several authors used marker controlled watershed \cite{Soille2004} with morphological approach, as following described.
 
@@
 Markers are computed as a combination of the binary mask of the red blood cells and centres of the cells which are computed using a similar algorithm that was utilized for the evaluation of the average cell radius.
-The binary mask is obtained by thresholding the grey-scale image with an automatically estimated threshold using Otsu method\cite{Otsu1975}.
+The binary mask is obtained by thresholding the grey-scale image with an automatically estimated threshold using Otsu method \cite{Otsu1975}.
 
-In \cite{Sulist2015} Sulistyawati \emph{et al.} combined morphological operations (erosion, dilation, opening and closing) and blob analysis to segment and identify malaria parasites with a high degree of accuracy.\\
+	In \cite{Sulist2015} Sulistyawati \emph{et al.} combined morphological operations (erosion, dilation, opening and closing) and blob analysis to segment and identify malaria parasites with a high degree of accuracy.
 
@@
 
-To conclude many systems for computer aided diagnosis of malaria disease made use of mathematical operations in order to smoothen the boundary of the regions obtained from the segmentation process.
-
-%%BOUNDARY BASED METHODS \subsubsection{Boundary based methods}
-Zou et al.
+	%Zou et al.
 \cite{Zou2010} based their method on the consideration that RBCs have a circular shape.
@@
 However, some RBCs may not have circular shape and it could affect the detection results.
+
+	To conclude, many systems for computer aided diagnosis of malaria disease made use of mathematical operations in order to smoothen the boundary of the regions obtained from the segmentation process.
+
+	%%BOUNDARY BASED METHODS \subsubsection{Boundary based methods}
+
 	%
@@
 
+	A summary of analysed methods is shown in Table 1.
+
 	\section{Conclusions}
@@
 	Finally, malaria diagnosis might be just one element of a suite of diagnostic software tests running on this type of system.
-Several other tests could simultaneously be carried out using the same images, for instance cell counting or detection of other hemoparasites like microfilaria or trypanosoma \cite{Rosado2016}.
+Several other tests could simultaneously be carried out using the same images, for instance cell counting or detection of other hemoparasites, like microfilaria or trypanosoma \cite{Rosado2016}.
 
@@
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
+	%\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
 The following statements should be used ``X.X.
@@
 		\begin{tabular}{@{}ll}
+			CBC &Complete Blood Count \\
 			WBC &White Blood Cell \\
-RBC & Red Blood Cell \\
-MP & Malaria Parasite \\
+			RBC &Red Blood Cell \\
+			MM &Mathematical Morphology \\
+			MP &Malaria Parasite \\
+			SE &Structuring Element
 	\end{tabular}}
@@
 	%% optional
-\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
-\appendixsections{multiple} %Leave argument "multiple" if there are multiple sections.
+	%\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A").
+If the appendix sections contain a heading then change the argument to "yes".
+	%\appendixsections{multiple} %Leave argument "multiple" if there are multiple sections.
 Then a counter is printed ("Appendix A").
 If there is only one appendix section then change the argument to "one" and no counter is printed ("Appendix").
-\appendix
-\section{}
-\subsection{}
-The appendix is an optional section that can contain details and data supplemental to the main text.
+	%\appendix
+	%\section{}
+	%\subsection{}
+	%The appendix is an optional section that can contain details and data supplemental to the main text.
 For example, explanations of experimental details that would disrupt the flow of the main text, but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data is shown in the main text can be added here if brief, or as Supplementary data.
 Mathematical proofs of results not central to the paper can be added as an appendix.
-
-\section{}
-All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.
+	%
+	%\section{}
+	%All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.
 
@@
 	%% optional
-\sampleavailability{Samples of the compounds ...... are available from the authors.}
+	%\sampleavailability{Samples of the compounds ...... are available from the authors.}
 

### 2018-01-25 MDPI_template/Sensors_DiRuberto.tex
Review Jan

--- 2017-12-07 MDPI_template/paper.tex
@@
 The parasites are spread to people through the bites of infected female Anopheles mosquitoes, called "malaria vectors".
-	There are five parasite species that cause malaria in humans and two of these species, Plasmodium falciparum and Plasmodium vivax, constitute the greatest threat.
-Plasmodium ovale, Plasmodium malariae and Plasmodium knowlesi are the three remaining species which are less dangerous in human \cite{WHO_dec_2016}, as shown in fig.\ref{fig:malaria_types}.
+There are five parasite species that cause malaria in humans and two of these species, Plasmodium Falciparum and Plasmodium Vivax, constitute the greatest threat.
+Plasmodium Ovale, Plasmodium Malariae and Plasmodium Knowlesi are the three remaining species which are less dangerous in human \cite{WHO_dec_2016}, as shown in fig.\ref{fig:malaria_types}.
 All five species may appear in four different life-cycle stages during the infection phase in peripheral blood: ring, trophozoite, schizont and gametocyte.
@@
 MM is not only a theory, but also a very powerful image analysis technique \cite{Soille2004}.
-	It was introduced by Matheron as a technique for analysing geometric structure of metallic and geologic samples.
+It was introduced by Matheron in 1964 as a technique for analysing geometric structure of metallic and geologic samples.
 It refers to a branch of non-linear image processing and analysis that concentrates on the geometric structure within an image.
@@
 They can be applied to extract image components useful in the representation and descriptions of region shapes, such as area granulometry, boundaries, skeleton, or convex hull.
-Also morphological operators can be used for image preprocessing and postprocessing, such as morphological filtering, thinning, and especially for segmentation.
+Also, morphological operators can be used for image preprocessing and postprocessing, such as morphological filtering, thinning, and especially for segmentation.
 
@@
 \begin{enumerate}
-		\item Pre-processing.
+	\item Preprocessing.
 	\item Segmentation.
@@
 \end{enumerate}
-	Since morphological techniques have been used in the first three phases, the reviewed works have been divided into the following sub-sections: pre-processing, segmentation and feature extraction.
+Since morphological techniques have been used in the first three phases, the reviewed works have been divided into the following sub-sections: preprocessing, segmentation and feature extraction.
 Each sub-section contains description about methods that cope with malaria parasites (MP) stained components analysis, both on thin and thick blood smears, without distinction.
 
-	%\subsection{Search of related literatures}
 Extensive search of articles has been made in PubMed and Google Scholar search engines based on the keywords: "malaria, mathematical morphology, automated malaria diagnosis" up to October  2017.
@@
 Thereafter, the full texts of these studies were evaluated as per the exclusion criteria.
-
-	%TODO inserire descrizione tabella "una sintesi dei lavori presi in considerazione è mostrata nella tabella X, bla bla bla ... Some works are missing because...
 
@@
 Giemsa stained blood smear is considered in most of the analysed literatures whereas Leishman stain is considered in few studies. %, as showed in table \ref{tab:summary_table}.
-
 It is reported that Leishman stain has bigger sensitivity for parasite detection than Giemsa \cite{Khan2014} and is superior for visualization of red and white blood cell morphology \cite{Sathpathi2014}.
@@
 \subsection{Preprocessing}
-	In image analysis field, especially when we refer to complex computer-aided pipelines, pre-processing methods are particularly used in order to improve the image data by suppressing unwanted noise or enhancing some image features for further processing.
-	It is worth to mention pre-processing methods because they are an important step regarding image analysis field but, for what concerns the malaria-affected blood image analysis, in our review we particularly found methods which operate for illumination correction and noise filtering purposes.
+In image analysis field, especially when we refer to complex computer-aided pipelines, preprocessing methods are particularly used in order to improve the image data by suppressing unwanted noise or enhancing some image features for further processing.
+It is worth to mention preprocessing methods because they are an important step regarding image analysis field but, for what concerns the malaria-affected blood image analysis, in our review we particularly found methods which operate for illumination correction and noise filtering purposes.
 Generally speaking, digital microscopy images can be acquired in different lighting conditions, with several types of acquisition devices or from blood smears stained with various staining protocols and, consequently, the features of similar images could differ a lot.
@@
 
-	In \cite{Kareem2012} Kareen \emph{et al.} illustrated a morphological approach for blood cell identification and use the image features such as intensity, histogram, relative size and geometry for further analysis.
+In \cite{Kareem2012} Kareem \emph{et al.} illustrated a morphological approach for blood cell identification and use the image features such as intensity, histogram, relative size and geometry for further analysis.
 Before the identification of blood cells, the authors  propose a novel morphological filtering based on the size of RBC for platelets and/or artifacts elimination.
-A dilation is performed by a concentric ring structuring element and erosion by disk shaped structuring element.
+A dilation is performed by a concentric ring structuring element and erosion by a disk-shaped structuring element.
 The radius of the structuring element depends on the radius of the RBCs, so that all the components smaller than the RBCs can be removed.
@@
 
-	Romero \emph{et al.} in \cite{Romero2016} presented an algorithm that uses morphological operations, the watershed method, the Hough transform and the clustering method of k-means to detect overlapped RBCs.
-In the pre-processing stage white blood cells and platelets are removed before the segmentation task.
+Romero-Rondon \emph{et al.} in \cite{Romero2016} presented an algorithm that uses morphological operations, the watershed method, the Hough transform and the clustering method of k-means to detect overlapped RBCs.
+In the preprocessing stage white blood cells and platelets are removed before the segmentation task.
 During this step, some noise, the WBC cytoplasm and platelets still remain on the image.
@@
 
-	Reni \emph{et al.} in \cite{Reni2015} described a new algorithm for morphological filtering of the blood images as a pre-processing tool for segmentation.
+Reni \emph{et al.} in \cite{Reni2015} described a new algorithm for morphological filtering of the blood images as a preprocessing tool for segmentation.
 Conventional morphological closing on blood images removes the unwanted components but also useful information.
@@
 As for many problems of image processing, there is no standard solution valid in general, so different segmentation techniques can be applied, according to the characteristics of the images to process and of the objects to segment.
-	Medical images segmentation is typically performed using two main strategies: the first level aims to separate whole cells or tissues from the background and the second one aims to separate the tissue structure in different regions or the cell in their components as the nucleus from the cytoplasm or intracellular parasites.
+Medical images segmentation is typically performed using two main strategies: the first level aims to separate whole cells or tissues from the background and the second one aims to separate the tissue structure in different regions or the cell in their components, as the nucleus from the cytoplasm or intracellular parasites.
 The latter case is commonly used in applications in which the cell class depends on the morphological characteristics of its components.
@@
 The shape of the objects (circular erythrocytes) is known a priori, but the image must be analysed to determine the size distribution of objects in the image and to find the average eccentricity of erythrocytes present.
-	Gray-scale granulometries based on opening with disk shape elements are then used.
-Non flat disk shaped structural element are applied to enhance the roundness and compactness of the red blood cells and flat disk shaped structural element applied to segment overlapping cells.
+Gray-scale granulometries based on opening with disk-shaped elements are then used.
+Non flat disk-shaped structural element are applied to enhance the roundness and compactness of the red blood cells and flat disk-shaped structural element applied to segment overlapping cells.
 The object to be segmented differs greatly in contrast from the background image.
@@
 The gradient image can be computed and a threshold can be applied to create a binary mask containing the segmented cell.
-The binary gradient mask is dilated using the vertical structuring element followed by the horizontal structuring element.
+The binary gradient mask is dilated using a vertical structuring element followed by a horizontal structuring element.
 The cell of interest has been successfully segmented, but it is not the only object that has been found.
@@
 The RBCs are detected as follows: firstly, a pixel classification allowed to label each image pixel as either background or foreground, based on its color features.
-Afterward, an inclusion-tree structure represented the hierarchical object relations between background and foreground so that a filtering process allows to remove irrelevant structures such as artifacts generated at the staining or digitization processes.
+Afterward, an inclusion-tree structure is used to represent the hierarchical object relations between background and foreground so that a filtering process allows to remove irrelevant structures such as artifacts generated at the staining or digitization processes.
 
@@
 In the algorithm described by Romero-Rondon \emph{et al.} in \cite{Romero2016} the detection of overlapped RBCs is still based on marker-controlled watershed transform.
-To define the suitable markers in watershed transform they used three different approaches, based on a morphological erosion operation, on Hough transform and on clustering method of k-means.
+To define the suitable markers in watershed transform they used three different approaches, based on a morphological erosion operation, on Hough transform and on clustering method of K-means.
 
-	Savkare \emph{et al.} in \cite{Savkare2015} segmented cells using k-mean clustering and global threshold.
+Savkare \emph{et al.} in \cite{Savkare2015} segmented cells using K-mean clustering and global threshold.
 Overlapping cells are separated using Sobel edge detector and watershed transform.
 Watershed transform is applied on each cluster separately.
-Over-segmentation is minimized by series of morphological operations, like erosion and dilation utilizing disk shaped structuring element.
+Over-segmentation is minimized by series of morphological operations, like erosion and dilation utilizing disk-shaped structuring elements.
 
@@
 It is the basis for the characterization of the concept of size.
-Some authors used area granulometry for pre-processing purposes in malaria characterization \cite{Tek2010} even though it is certainly effective for extracting cells size features information \cite{Tek2006},\cite{Malihi2013},\cite{Springl2009}.
+Some authors used area granulometry for preprocessing purposes in malaria characterization \cite{Tek2010} even though it is certainly effective for extracting cells size features information \cite{Tek2006},\cite{Malihi2013},\cite{Springl2009}.
 In \cite{Tek2010} local area granulometry combined with colour histogram are used as features.
@@
 %% optional
-	%\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A").
-If the appendix sections contain a heading then change the argument to "yes".
+%\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
 %\appendixsections{multiple} %Leave argument "multiple" if there are multiple sections.

### 2018-01-25 MDPI_template/Sensors_DiRuberto_rev.tex
Review Jan

--- 2018-01-25 MDPI_template/Sensors_DiRuberto.tex
@@
 The scope of this paper is to review and analyze the recent works of different researchers in the area of malaria parasite recognition using computer vision which benefit from mathematical morphology.
+Only few reviews exist in literature about microscopic image processing for malaria parasites.
+However, they are not focused on MM techniques as they analyze generic computer vision systems for malaria diagnosis, as in \cite{Rosado2017}, \cite{Das2015} or \cite{Tek2009}.
+Also, newer and promising approaches, addressing for example the problem of malaria diagnosis in remote areas \cite{Oliveira2017}, \cite{Rosado2017}, or improving significantly both the detection and the classification performances \cite{Dave2017}, \cite{Devi2017}, \cite{Somasekar2017} and \cite{Sulist2015}, have not been considered in the previous reviews.
 
@@
 
-The proposed automated method in \cite{Elter2011} by Elter \emph{et al.} for parasite detection and identification worked on thin blood film acquired with Giemsa stain.
+The automated method proposed in \cite{Elter2011} by Elter \emph{et al.} for parasite detection and identification worked on thin blood film acquired with Giemsa stain.
 The authors found that the G and B channels of the RGB colour are very good features to identify objects containing chromatin in Giemsa stained blood films, being not only considered highly discriminative but also almost independent of differences in illumination and staining intensity.
@@
 In \cite{Ross2006} Ross \emph{et al.} proposed a method which provides a positive or negative diagnosis of malaria and differentiates parasites by species.
-The segmentation step relies on a thresholding strategy which aims to identify and segment potential parasites and erythrocytes from the image background after a six steps threshold selection.
-Mathematical morphology has been used for parasite size estimation, erythrocytes reconstruction and cells bigger than erythrocytes removal.
+The segmentation step relies on a six steps thresholding selection strategy.
+It aims to identify and segment potential parasites and erythrocytes from background.
+Mathematical morphology has been used in several key steps of the procedure.
+Hole filling is used in the first step in order to fill RBCs' binary masks obtained from a first thresholding.
+Afterwards, step 4 employs RBCs' morphological reconstruction with parasites' mask, found in step 2, for identifying infected cells.
+In step 5, a morphological opening filter, using a disk-shaped SE with radius equal to the mean erythrocyte radius less the standard deviation, is applied to the grey-scale, morphologically filtered, green component in order to remove any objects smaller than an erythrocyte.
+The morphological gradient - difference between a dilation and erosion of the image - is then calculated using a diamond-shaped SE with unity length.
+Finally, in step 6, the intersection of morphological gradient image and the dilated cell cluster is calculated.
+This image is then transformed to a binary image by thresholding any value greater than zero.
+A series of morphological operations, namely a closing operation, thinning, and spur-removal are then applied to generate a contour of the segmented erythrocytes.
+Contours are filled, and the segmented mask is again reconstructed with the valid parasite marker image to result in a segmented mask of infected cells.
+RBCs and parasites masks are consequently ready for next generation step.
 
@@
 Also in the method proposed in \cite{Somasekar2017} by Somasekar \emph{et al.} the segmentation of the infected parasites is based on thresholding.
-The segmentation is achieved in two stages by maximizing between-class variance of an original image and consequently by an iterative threshold selection from a stage-one threshold image with suitable stopping criteria.
-The segmented results are post processed to improve the accuracy of the detection of malaria parasites by morphological operators (erosion and closing).
-
-%Makkapati and Rao proposed two different methodologies.
-The first one \cite{Makkapati2009} is based on HSV colour space with the purpose of performing a red blood cells (RBCs) segmentation and an identification of RBCs affected from MP with a chromatin dot identification by thresholding the Saturation channel.
-The RBCs segmentation is achieved by dividing the entire hue range of $360^{\circ}$ into six segments and finding the dominant colour type to be the representative of the background.
-On the other hand, the second work \cite{Makkapati2011} is devoted to the detection of the parasite stage and identification of the species.
-It is based on ontological description and uses the morphology of the two basic components of a parasite, namely, chromatin dot(s) and cytoplasm.
-The authors state that the green component of an image offers maximum separation between the background and RBC pixels.
-The green component is first segmented to detect the RBCs by choosing an optimal threshold [10].
-The pixels belonging to chromatin dots have higher saturation values and are segmented by thresholding the saturation component of the image.
-However, several artefacts may also be segmented as chromatin
-%dots, so a further evaluation is performed to establish if a chromatin dot is valid by identifying whether it is inside a RBC or not.
-
-%The morphology of P.falciparum at gametocyte stage is the key parameter in the segmentation approach proposed in \cite{Kareem2012}.
-It is based on a modification of the ARR transform.
-The method can be used as an efficient tool for gametocyte detection for post-treatment malaria diagnosis.
-It also identifies the presence of any White Blood Cells (WBCs) in the image, and discards other artefacts and non infected cells.
-It utilizes the information based on structure, colour and geometry of the cells and does not require any segmentation or non-illumination correction techniques that are commonly used for cell detection.
-
-%Fang et al.
-\cite{Fang2011} used a threshold to distinguish the unhealthy cells from the healthy cells after a malaria detection phase performed with a phase spectrum methodology.
-This method uses a the Quaternion Fourier Transform (QFT) to obtain the amplitude and phase spectrum for blood smear images.
-Then, the reconstructed image is obtained using the Inverse Quaternion Fourier transform (IQFT) on a constant amplitude spectrum and the original phase spectrum.
-The authors concluded that B and G channels from the RGB colour space and I channel from the HSI colour space are the most sensitive channels for MP detection.
-
-%Malihi et al.
-\cite{Malihi2013} proposed a method for malaria parasite detection in Giemsa-stained blood samples.
-They used both binary image masks obtained with Otsu method and the edge detection images from Canny method for RBC segmentation, followed by a hole filling process.
-
-%The following strategy has been used:
-%\begin{enumerate}
-%	\item an algorithm based on run-length encoding has been used for analysing the connected components of the binary masks obtained in segmentation phase;
-%	\item preliminary tagging, keeping the label equivalences or connected regions with different labels in a table of local matches. Resolution of the equivalences is, then, performed;
-%	\item measurements on the remaining connected components: WBCs and MP region areas are checked and WBCs connected components are removed with a closing operation.
-%\end{enumerate}
-%
+It is achieved in two stages by maximizing between-class variance of an original image and consequently by an iterative threshold selection from a stage-one threshold image with suitable stopping criteria.
+The segmented results are post processed to improve the accuracy of malaria parasites detection by morphological operators (erosion and closing).
 
@@
 
-Airwhar \emph{et al.} \cite{Ahirwar2012} based their approach on thresholding and granulometry.
+%Airwhar \emph{et al.} \cite{Ahirwar2012} based their approach on thresholding and granulometry.
 The histogram of the complemented, green component has been used and it is said to be a bimodal distribution in all the considered images.
@@
 Those objects that do not exceed the area of a circle with radius equal to the mean erythrocyte radius plus the standard deviation are regarded as being single cells, and are unmodified.
-On the other hand, the clumped cells are segmented as follows.
+%On the other hand, the clumped cells are segmented as follows.
 First, the intersection of the morphological gradient image and the dilated cell cluster is taken.
@@
 
-Soni \emph{et al.} \cite{Soni2011} performed segmentation of erythrocytes by using granulometry as well.
+%Soni \emph{et al.} \cite{Soni2011} performed segmentation of erythrocytes by using granulometry as well.
 The size and eccentricity of the erythrocytes are also required for the calculation of some feature values (as these can be indicative of infection).
 The shape of the objects (circular erythrocytes) is known a priori, but the image must be analysed to determine the size distribution of objects in the image and to find the average eccentricity of erythrocytes present.
-Gray-scale granulometries based on opening with disk-shaped elements are then used.
+%Gray-scale granulometries based on opening with disk-shaped elements are then used.
 Non flat disk-shaped structural element are applied to enhance the roundness and compactness of the red blood cells and flat disk-shaped structural element applied to segment overlapping cells.
@@
 
-%Zou et al.
-\cite{Zou2010} based their method on the consideration that RBCs have a circular shape.
-They propose an improved version of the Circle Hough Transform (CHT) for RBCs detection.
-The authors state that overlapping and oval cells can be detected as well as non-clumped RBCs and affirm that the green channel of the RGB colour space permits to better distinguish RBCs.
-However, some RBCs may not have circular shape and it could affect the detection results.
+To sum up, the analysed systems use mathematical morphology methods during or immediately ensuing segmentation step for the following purposes:
 
-To conclude, many systems for computer aided diagnosis of malaria disease made use of mathematical operations in order to smoothen the boundary of the regions obtained from the segmentation process.
+\begin{itemize}
+	\item size evaluation of regions obtained from segmentation
+	\item image cleaning and artefacts removal
+	\item RBCs and parasites segmentation or separation
+	\item improving parasites detection
+	\item minimization of watershed over-segmentation.
+\end{itemize}
 
 %%BOUNDARY BASED METHODS \subsubsection{Boundary based methods}
-
-%
 
@@
 In particular, a lot of authors affirm that Otsu thresholding suffers from limitations when textural variation is high, while histogram thresholding can not deal sufficiently good in identifying valley regions in case of unimodal histograms.
-However, such a simple and fast approach can greatly benefit from mathematical morphology as recent studies demonstrate (\cite{Ahirwar2012}, \cite{Anggraini2011}, \cite{Arco2014}, \cite{DiRuberto2002}, \cite{Elter2011}, \cite{Ghosh2011},  \cite{Kareem2011}, \cite{Mushabe2013},  \cite{Rosado2017}, \cite{Ross2006}, \cite{Savkare2011b}, \cite{Savkare2015}, \cite{Somasekar2017}, \cite{Tek2010}).
+However, such a simple and fast approach can greatly benefit from mathematical morphology as recent studies demonstrate, \cite{Anggraini2011}, \cite{Arco2014}, \cite{DiRuberto2002}, \cite{Elter2011}, \cite{Ghosh2011},  \cite{Kareem2011}, \cite{Mushabe2013},  \cite{Rosado2017}, \cite{Ross2006}, \cite{Savkare2011b}, \cite{Savkare2015}, \cite{Somasekar2017}, \cite{Tek2010}).
 
@@
 
-Other authors (\cite{Ahirwar2012}, \cite{DiRuberto2002}, \cite{Khan2011}, \cite{Mushabe2013}, \cite{Ross2006}, \cite{Soni2011}, \cite{Tek2010}) employed granulometry and stated that it is very effective to segment cells with regular size.
+Other authors, \cite{DiRuberto2002}, \cite{Khan2011}, \cite{Mushabe2013}, \cite{Ross2006}, \cite{Tek2010}) employed granulometry and stated that it is very effective to segment cells with regular size.
 
 The analysed works performed classification phase for different purposes.
-The majority of them aimed to distinguish among two classes only, malaria infected and noninfected RBCs, or to detect and count parasites in a malaria blood image (\cite{Anggraini2011}, \cite{Arco2014}, \cite{Das2011}, \cite{Dave2017}, \cite{Das2015}, \cite{DiRuberto2002}, \cite{Elter2011}, \cite{Ghosh2011}, \cite{Kareem2012}, \cite{Khan2014}, \cite{Malihi2013}, \cite{Mushabe2013}, \cite{Oliveira2017}, \cite{Savkare2011b}, \cite{Savkare2011a}, \cite{Somasekar2011}, \cite{Somasekar2015}, \cite{Somasekar2017}, \cite{Soni2011}, \cite{Sulist2015}, \cite{Tek2006}).
+The majority of them aimed to distinguish among two classes only, malaria infected and noninfected RBCs, or to detect and count parasites in a malaria blood image (\cite{Anggraini2011}, \cite{Arco2014}, \cite{Das2011}, \cite{Dave2017}, \cite{Das2015}, \cite{DiRuberto2002}, \cite{Elter2011}, \cite{Ghosh2011}, \cite{Kareem2012}, \cite{Khan2014}, \cite{Malihi2013}, \cite{Mushabe2013}, \cite{Oliveira2017}, \cite{Savkare2011b}, \cite{Savkare2011a}, \cite{Somasekar2011}, \cite{Somasekar2015}, \cite{Somasekar2017}, \cite{Sulist2015}, \cite{Tek2006}).
 
-More complex classification strategies aimed to classify parasites into different classes, i.e.
-different human parasites species (\cite{Ahirwar2012}, \cite{Das2013}, \cite{Das2014}, \cite{Khan2011}, \cite{Tek2010}),
+More complex classification strategies aimed to classify parasites into different classes, i.e. different human parasites species \cite{Das2013}, \cite{Das2014}, \cite{Khan2011}, \cite{Tek2010}),
 and/or different parasites life stages (\cite{Anggraini2011}, \cite{Das2013}, \cite{Das2014}, \cite{DiRuberto2002}, \cite{Diaz2009}, \cite{Tek2010}).
@@
 A summary of analysed methods is shown in Table 1.
+For each approach we report the processing phase where MM is used (preprocessing and/or segmentation), which operators are applied, which type of morphology (grayscale or binary), and shape and size of the used SE, if described.
+We detail the kind of classification, i.e.
+if a method addresses the detection problem (infected and non infected) only, if it faces the labelling of different parasites (\emph{P.falciparum}, \emph{P.ovale}, \emph{P.malariae} or \emph{P.vivax}) and, in some cases, the different stages of life (ring, trophozoite, schizont, gametocyte).
+Finally, the measures used for evaluating the system performance are reported, if present.
 
@@
 The main efforts towards the realization of a fully automatic blood cells segmentation and classification system cannot leave this aspect out.
+Usually, any  pattern recognition system is evaluated by performing tests using real set of samples.
+In case of diagnosis this is a key aspect.
+Any system for diagnosis should be tested on different images, acquired by different sources, and different specimens.
+This is crucial to guarantee the diagnostic capability and usefulness of a computer vision system.
 
 Mathematical morphology techniques have been widely used for image processing purposes.
-Among the application fields, it has been applied for fingerprint feature extraction, recognition of handwritten digits, license plate detection, border extraction, denoising using morphological filters, text extraction and detection of imperfection in printed circuit boards \cite{Kaur2013}.
+Among the application fields, it has been applied for fingerprint feature extraction, recognition of handwritten digits, license plate detection, border extraction, denoising using morphological filters, text extraction and so on.
 Apart from this kind of fields, mathematical morphology has been employed successfully in biomedical image analysis, especially in preprocessing and segmentation techniques.
@@
 %=====================================
+\reftitle{References}
 \externalbibliography{yes}

### 2018-01-25 MDPI_template/Table.tex
Added malariae imgs and courtesy
Review Jan

@@
 \newcolumntype{H}{>{\setbox0=\hbox\bgroup}c<{\egroup}@{}}
-    \begin{longtable}{C{2.1cm} C{2.1cm} C{2.3cm} H C{2.7cm} C{2.2cm} }
+    \begin{longtable}{C{2.1cm} C{2.7cm} C{2.2cm} C{2.0cm} C{2.7cm} C{2.0cm} }
     %\begin{longtable}{|C{2cm}|c|C{2.5cm}|C{3cm}|C{2.4cm}|c|}
@@
 		\hline
-    	\textbf{Authors} & \textbf{Preprocessing} & \textbf{Segmentation} & \textbf{Features} & \textbf{Classification} & \textbf{Performance}  \\[1pt] \hline
+    	\textbf{Authors} & \textbf{Preprocessing} & \textbf{Segmentation} & \textbf{MM/SE} & \textbf{Classification} & \textbf{Performance}  \\[1pt] \hline
 
-    Ahirwar \emph{et al.}, 2012  &
-    	- &
-    	thresholding + granulometry, opening, morphological gradient, dilation, closing, thinning, spur removal &
-    	- &
-    	five (\emph{P.falciparum}, \emph{P.vivax}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) &
-    	- \vspace{0.6cm} \\
-    	%\hline
-
+%    Ahirwar \emph{et al.}, 2012  &
+%    	- &
+%    	thresholding + granulometry, opening, morphological gradient, dilation, closing, thinning, spur removal &
+%    	gray, bin/disk of size depending on RBCs, diamond of size=1 &
+%    	five (\emph{P.falciparum}, \emph{P.vivax}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) &
+%    	-  \\
+%    	%\hline
+%&   &  &  &  & \\
     Anggraini \emph{et al.}, 2011 &
@@
 		thresholding + hole filling &	
-		- &
+		gray, bin &
 		two (\emph{P.falciparum} infected and noninfected) + two life-cycle-stages &
 		SE=93\% SP=99\%
-        \vspace{0.6cm}
+        %\vspace{0.6cm}
 		\\
+&   &  &  &  & \\
     Arco \emph{et al.}, 2014 &
@@
 		adaptive thresholding + hole filling, closing, regional minima &	
-		- &
+		gray, bin/disk &
 		two (infected and noninfected) &
 		Acc=96.46\%
-		\vspace{0.6cm}
+		%\vspace{0.6cm}
         \\
+        &   &  &  &  & \\
     Das \emph{et al.}, 2011 &
@@
 		marker controlled watershed + opening, closing &	
-		opening, closing  &
+		gray, bin  &
 		two (infected and noninfected) &
 		Acc=88.77\%
-		\vspace{0.6cm}
+		%\vspace{0.6cm}
         \\
-
+        &   &  &  &  & \\
     Das \emph{et al.}, 2013 &
@@
 		marker controlled watershed &	
-		- &
+		gray &
 	three (\emph{P.falciparum}, \emph{P.vivax} infected and noninfected) + three life-cycle-stages per species &
 		Acc=84\%
-		\vspace{0.6cm}
+%		\vspace{0.6cm}
         \\
-
+&   &  &  &  & \\
     Das \emph{et al.}, 2014 &
@@
 		marker controlled watershed &	
-		- &
+		gray &
 	three (\emph{P.falciparum}, \emph{P.vivax} infected and noninfected) + three life-cycle-stages per species &
 		SE=99.72\% SP=84.39\%
-		\vspace{0.6cm}
+		%\vspace{0.6cm}
         \\
-
+&   &  &  &  & \\
     Dave \emph{et al.}, 2017 &
@@
 		adaptive thresholding + erosion, dilation &	
-		- &
+		bin &
 		two (infected and noninfected) &
 		Acc=97.83\% thin films, Acc=89.88\% thick films
-		\vspace{0.6cm}
+		%\vspace{0.6cm}
         \\
-
+&   &  &  &  & \\
     Devi \emph{et al.}, 2017 &
@@
 		marker controlled watershed &	
-		- &
+		gray &
 		two (infected and noninfected)  &
 		Acc=98.02\%
-		\vspace{0.6cm}
+		%\vspace{0.6cm}
         \\
-
+&   &  &  &  & \\
     Diaz \emph{et al.}, 2009 &
@@
 		inclusion tree &	
-		- &
+		gray &
 		two (\emph{P.falciparum} infected and noninfected) + three life-cycle-stages &
 		SE=94\% SP=99.7\% for detection, SE=78.8\% SP=91.2\% for life-stages
-\vspace{0.6cm}
+%\vspace{0.6cm}
 		\\
-
+&   &  &  &  & \\
     Di Ruberto \emph{et al.}, 2002 &
@@
 		thresholding + granulometry, watershed  transform + skeleton &	
-		skeleton &
+		gray, bin/disk, flat and nonflat, with size depending on RBCs &
 		two (\emph{P.falciparum} infected and noninfected) + three life-cycle-stages &
 		-
-\vspace{0.6cm}
+%\vspace{0.6cm}
 		\\
-
+&   &  &  &  & \\
     Elter \emph{et al.}, 2011 &
@@
 		thresholding + black top-hat, dilation &	
-		- &
+		gray/disk, nonflat with size=9 &
 		two (infected and noninfected) &
 		SE=97\%
-		\vspace{0.6cm} \\
-
+%		\vspace{0.6cm}
+\\
+&   &  &  &  & \\
     Gonzalez-Betancourt \emph{et al.}, 2016  &
@@
     	watershed  transform  &
+    	gray/disk with size depending on RBCs &
     	- &
-    	- &
-    	- \vspace{0.6cm} \\
-
+    	-
+    %\vspace{0.6cm}
+    \\
+    &   &  &  &  & \\	
     Ghosh \emph{et al.}, 2011  &
@@
     	thresholding + opening, closing  &
-    	- &
+    	bin &
     	two (\emph{P.vivax} infected and noninfected) &
-    	- \vspace{0.6cm} \\
-
-    Kareem \emph{et al.}, 2011, 2012  &
+    	-
+    %\vspace{0.6cm}
+    \\
+&   &  &  &  & \\
+    Kamreem \emph{et al.}, 2011, 2012  &
     	dilation, erosion &
     	-  &
-    	- &
+    	gray/concentric ring, disk with size depending on RBCs  &
     	two (infected and noninfected) &
-    	Acc=88\% SE=90\% SP=86\% \vspace{0.6cm} \\
-
+    	Acc=88\% SE=90\% SP=86\%
+    %\vspace{0.6cm}
+    \\	
+&   &  &  &  & \\
     Khan \emph{et al.}, 2011  &
@@
     	thresholding + granulometry, opening, morphological reconstruction, gradient, dilation &
-    	- &
+    	gray &
     	five (\emph{P.falciparum}, \emph{P.vivax}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) &
-    	Acc=81\% SE=85.5\% \vspace{0.6cm} \\
-
+    	Acc=81\% SE=85.5\%
+    %\vspace{0.6cm}
+    \\	
+&   &  &  &  & \\
     Malihi \emph{et al.}, 2013  &
@@
     	area granulometry  &
-    	area granulometry &
+    	gray/disk with size depending on RBCs &
     	two (infected and noninfected) &
-    	Acc=91\% SE=80\% SP=95.5\% \vspace{0.6cm} \\
-
+    	Acc=91\% SE=80\% SP=95.5\%
+    %\vspace{0.6cm}
+    \\	
+&   &  &  &  & \\
     Mushabe \emph{et al.}, 2013  &
@@
     	thresholding + granulometry, dilation, erosion  &
+    	gray, bin/disk &
+    	two (infected and noninfected) &
+    	SE=98.5 SP=97.2\%
+    %\vspace{0.6cm}
+    \\	
+&   &  &  &  & \\
+    Oliveira \emph{et al.}, 2017  &
+    	erosion &
+    	gray, bin  &
     	- &
     	two (infected and noninfected) &
-    	SE=98.5 SP=97.2\% \vspace{0.6cm} \\
-
-    Oliveira \emph{et al.}, 2017  &
-    	erosion &
-    	-  &
-    	- &
-    	two (infected and noninfected) &
-    	Acc=91\% \vspace{0.6cm} \\
-
+    	Acc=91\%
+    %\vspace{0.6cm}
+    \\
+&   &  &  &  & \\
     Reni \emph{et al.}, 2015  &
@@
     	-  &
+    	gray/anular ring, disk with size depending on RBCs &
     	- &
-    	- &
-    	- \vspace{0.6cm} \\
-
+    	-
+    %\vspace{0.6cm}
+    \\
+&   &  &  &  & \\
     Romero-Rondon \emph{et al.}, 2016  &
@@
     	marker controlled watershed, erosion  &
+    	gray, bin/disk with size depending on RBCs &
     	- &
-    	- &
-    	- \vspace{0.6cm} \\
-
+    	-
+    %\vspace{0.6cm}
+    \\
+&   &  &  &  & \\
     Rosado \emph{et al.}, 2017  &
@@
     	adaptive thresholding + closing  &
-    	- &
+    	bin/elliptical with size=3 &
     	four (\emph{P.falciparum}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) +
     three life-cycle-stages for species &
-    	SE=73.9-96.2\% SP=92.6-99.3\% \vspace{0.6cm} \\
-
+    	SE=73.9-96.2\% SP=92.6-99.3\%
+    %\vspace{0.6cm}
+    \\
+&   &  &  &  & \\
     Ross \emph{et al.}, 2006  &
@@
     	thresholding + granulometry, opening, reconstruction, morphological gradient, closing, thinning  &
-        - &
+        gray, bin/disk with size=6 and depending on RBCs, diamond with size=1 &
     	five (\emph{P.falciparum}, \emph{P.vivax}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) &
     	SE=85\% for detection, Acc=73\% for classification
-    	 \vspace{0.6cm} \\
-
+    	 %\vspace{0.6cm}
+    \\
+&   &  &  &  & \\
     Savkare \emph{et al.}, 2011a  &
@@
     	thresholding + hole filling, watershed transform  &
-    	- &
+    	bin &
     	two (infected and noninfected) &
-    	- \vspace{0.6cm} \\
-
+    	-
+%    \vspace{0.6cm}
+    \\
+&   &  &  &  & \\
     Savkare \emph{et al.}, 2011b  &
@@
     	thresholding + hole filling, watershed transform  &
-    	- &
+    	bin &
     	two (infected and noninfected) &
-    	SE=93.12\% SP=93.17\% \vspace{0.6cm} \\
-
+    	SE=93.12\% SP=93.17\%
+    %\vspace{0.6cm}
+    \\
+&   &  &  &  & \\
     Savkare \emph{et al.}, 2015  &
@@
     	thresholding + watershed transform, erosion, dilation  &
-    	- &
+    	bin/disk with size=2 &
     	two (infected and noninfected) &
-    	Acc=95.5\% \vspace{0.6cm} \\
-
+    	Acc=95.5\%
+    %\vspace{0.6cm}
+    \\
+&   &  &  &  & \\
     Sheikhhosseini \emph{et al.}, 2013  &
@@
     	thresholding + hole filling, opening  &
-    	- &
+    	bin &
     	two (infected and noninfected) &
-    	Acc=97.25\% SE=82.21\% SP=98.02\% \vspace{0.6cm} \\
-
+    	Acc=97.25\% SE=82.21\% SP=98.02\%
+    %\vspace{0.6cm}
+    \\
+&   &  &  &  & \\
     Somasekar \emph{et al.}, 2015  &
@@
     	fuzzy C-means clustering + erosion, hole filling  &
-    	- &
+    	bin/square with size=3 &
     	two (infected and noninfected) &
-    	SE=98\% SP=93.3\% \vspace{0.6cm} \\
-
+    	SE=98\% SP=93.3\%
+    %\vspace{0.6cm}
+    \\
+&   &  &  &  & \\
     Somasekar \emph{et al.}, 2017  &
@@
     	thresholding + erosion, closing, hole filling  &
-    	- &
+    	bin/square with size=3 &
     	two (infected and noninfected) &
-    	average DSC=0.8 \vspace{0.6cm} \\
-
-    Soni \emph{et al.}, 2011  &
-    	- &
-    	thresholding + granulometry, morphological gradient, dilation  &
-    	- &
-    	five (\emph{P.falciparum}, \emph{P.vivax}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) &
-    	SE=98\% for detection  \vspace{0.6cm} \\
-
+    	average DSC=0.8
+    %\vspace{0.6cm}
+    \\
+%&   &  &  &  & \\
+%    Soni \emph{et al.}, 2011  &
+%    	- &
+%    	thresholding + granulometry, morphological gradient, dilation  &
+%    	gray, bin/disk, flat and nonflat, vertical, horizontal &
+%    	five (\emph{P.falciparum}, \emph{P.vivax}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) &
+%    	SE=98\% for detection
+%    %\vspace{0.6cm}
+%    \\
+&   &  &  &  & \\
 {\v{S}}pringl, 2009  &
@@
     	thresholding + marker controlled watershed transform, hole filling, dilation, opening, erosion  &
-    	- &
+    	gray, bin/disk with size depending on RBCs &
     	two (infected and noninfected) &
-    	AUC=0.98 \vspace{0.6cm} \\
-
+    	AUC=0.98
+    %\vspace{0.6cm}
+    \\
+&   &  &  &  & \\
     Sulistyawati \emph{et al.}, 2015  &
@@
     	blob analysis + erosion, dilation, opening, closing, hole filling  &
-    	- &
+    	bin &
     	two (infected and noninfected) &
-    	Acc=99.39\% \vspace{0.6cm} \\
-
+    	Acc=99.39\%
+    %\vspace{0.6cm}
+    \\
+&   &  &  &  & \\
  Tek \emph{et al.}, 2006  &
@@
     	top-hat, infinite reconstruction, area granulometry  &
-    	area granulometry &
+    	gray, bin &
     	two (infected and noninfected) &
-    	SE=74\% SP=98\% \vspace{0.6cm} \\
-
+    	SE=74\% SP=98\%
+    %\vspace{0.6cm}
+    \\
+&   &  &  &  & \\
     Tek \emph{et al.}, 2010  &
@@
     	thresholding + granulometry, area top-hat, closing, area granulometry &
-    	area granulometry &
+    	gray/disk with size depending on RBCs &
     	five (\emph{P.falciparum}, \emph{P.vivax}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) +
     four life-cycle-stages for species &
-    	SE=72\% SP=98\% \vspace{0.6cm} \\
-
+    	SE=72\% SP=98\%
+    %\vspace{0.6cm}
+    \\
+&   &  &  &  & \\
     Yunda \emph{et al.}, 2012  &
@@
     	thresholding + morphological gradient, erosion, dilation  &
-    	- &
+    	bin &
     	three (\emph{P.falciparum}, \emph{P.vivax} infected, and noninfected) + two life-cycle-stages for \emph{P.falciparum}&
@@
 
-	\caption{Summary of analysed methods: morphological operations used in the main phases of analysis, kind of classification and performance measures (Sensitivity, Specificity, Accuracy, if reported).} % needs to go inside longtable environment
+	\caption{Summary of analysed methods: morphological operations used in the main phases of analysis, type of MM (gray or binary)/type and size of SE (if reported), kind of classification and performance measures (Sensitivity, Specificity, Accuracy, if reported).} % needs to go inside longtable environment
 	\label{tab:summary_table}

### 2018-02-13 MDPI_template/sensors-257497-proofreading.tex
Sensors final version

%  LaTeX support: latex@mdpi.com
%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).

% You need to save the "mdpi.cls" and "mdpi.bst" files into the same folder as this template file.

%=================================================================
\documentclass[sensors,review,accept,moreauthors,pdftex,10pt,a4paper]{mdpi}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}

% definition of keywords
\def\keywords{\vspace{.5em}{\bfseries\textit{Keywords}---\,\relax}}
\def\endkeywords{\par}

% Packages for table
\usepackage{longtable}
\usepackage{multirow}
\usepackage{fancyhdr}
\usepackage{booktabs}
\usepackage{array}

\usepackage{float}
\usepackage{filecontents}

\usepackage[tight,footnotesize]{subfigure}

\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\usepackage{booktabs}
\usepackage{multirow}

%
%--------------------
% Class Options:
%--------------------
% journal
%----------
\usepackage{subfigure} 
\makeatletter
\renewcommand{\@thesubfigure}{\normalsize(\textbf{\alph{subfigure}})}
\makeatother
% Choose between the following MDPI journals:
% actuators, addictions, admsci, aerospace, agriculture, agronomy, algorithms, animals, antibiotics, antibodies, antioxidants, applsci, arts, asi, atmosphere, atoms, axioms, batteries, bdcc, behavsci, beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsci, buildings, carbon, cancers, catalysts, cells, ceramics, challenges, chemengineering, chemosensors, children, chromatography, climate, coatings, colloids, computation, computers, condensedmatter, cosmetics, cryptography, crystals, cybersecurity, data, dentistry, designs, diagnostics, diseases, diversity, econometrics, economies, education, electrochemistry, electronics, energies, entropy, environments, epigenomes, est, fermentation, fibers, fire, fishes, fluids, foods, forests, fractalfract, futureinternet, galaxies, games, gastrointestdisord, gels, genealogy, genes, geosciences, geriatrics, hazardousmatters, healthcare, heritage, highthroughput, horticulturae, humanities, hydrology, informatics, information, infrastructures, inorganics, insects, instruments, ijerph, ijfs, ijms, ijgi, ijtpp, inventions, j, jcdd, jcm, jcs, jdb, jfb, jfmk, jimaging, jof, jintelligence, jlpea, jmmp, jmse, jpm, jrfm, jsan, land, languages, laws, life, literature, logistics, lubricants, machines, magnetochemistry, make, marinedrugs, materials, mathematics, mca, mti, medsci, medicines, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, modelling, molbank, molecules, mps, nanomaterials, ncrna, neonatalscreening, nitrogen, nutrients, ohbm, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, plants, plasma, polymers, preprints, proceedings, processes, proteomes, publications, quaternary, qubs, recycling, religions, remotesensing, resources, risks, robotics, safety, scipharm, sensors, separations, sexes, sinusitis, socsci, societies, soils, sports, standards, surgeries, sustainability, symmetry, systems, technologies, toxics, toxins, tropicalmed, universe, urbansci, vaccines, vetsci, vibration, viruses, vision, water, wem
%---------
% article
%---------
% The default type of manuscript is article, but can be replaced by:
% abstract, addendum, article, benchmark, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, correction, conferenceproceedings, conferencereport, expressionofconcern, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimages, letter, meetingreport, newbookreceived, opinion, obituary, projectreport, reply, reprint, retraction, review, perspective, protocol, shortnote, supfile, technicalnote, viewpoint
% supfile = supplementary materials
% protocol: If you are preparing a "Protocol" paper, please refer to http://www.mdpi.com/journal/mps/instructions for details on its expected structure and content.
%----------
\usepackage{booktabs} 
\usepackage{multirow}
\usepackage{soul} 
\usepackage{microtype}
\setitemize{parsep=6pt,itemsep=0pt,leftmargin=*,labelsep=5.5mm}
\setenumerate{parsep=6pt,itemsep=0pt,leftmargin=*,labelsep=5.5mm}
\setlist[description]{itemsep=0mm}   
\usepackage{array}
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g. the logo of the journal will get visible), the headings, and the copyright information. In addition, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.

%=================================================================
\firstpage{1}
\makeatletter
\setcounter{page}{\@firstpage}
\makeatother
\articlenumber{x}
\doinum{10.3390/------}
\pubvolume{18}
\pubyear{2018}
\copyrightyear{2018}
\history{Received: 14 December 2017; Accepted: 5 February 2018; Published: }

%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, calc, indentfirst, fancyhdr, graphicx, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, amsthm, hyphenat, natbib, hyperref, footmisc, geometry, caption, url, mdframed, tabto, soul, multirow, microtype, tikz

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Recent Advances of Malaria Parasites Detection Systems Based on Mathematical Morphology}

% Author Orchid ID: enter ID or remove command
\newcommand{\orcidauthorA}{0000-0002-6571-3816} % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-000-000X} % Add \orcidB{} behind the author's name

% Authors, for the paper (add full first names)
\Author{Andrea Loddo $^{1,\dagger}$\orcidA{}, Cecilia Di Ruberto $^{1,\ddagger}$ and Michel Kocher$^{2,\ddagger}$}

% Authors, for metadata in PDF
\AuthorNames{Andrea Loddo, Cecilia Di Ruberto and Michel Kocher}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
	$^{1}$ \quad Department of Mathematics and Computer Science, University of Cagliari, 09124 Cagliari, Italy; andrea.loddo@unica.it, dirubert@unica.it\\
	$^{2}$ \quad Biomedical Imaging Group, École Polytechnique Fédérale de Lausanne (EPFL), Switzerland; michel.kocher@heig-vd.ch}

% Contact information of the corresponding author
\corres{Correspondence: andrea.loddo@unica.it; Tel.: +39-070-675-8503}

% Current address and/or shared authorship
\firstnote{These authors contributed equally to this work.}
% The commands \thirdnote{} till \eighthnote{} are available for further notes

% Simple summary
\simplesumm{}

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{Malaria is an epidemic health disease and a rapid, accurate diagnosis is necessary for proper intervention. Generally, pathologists visually examine blood stained slides for malaria diagnosis.
Nevertheless, this kind of visual inspection is subjective, error-prone and time-consuming. In order to overcome the issues, numerous methods of automatic malaria diagnosis have been proposed so far. In particular, many researchers have used mathematical morphology as a powerful tool for computer aided malaria detection and classification.
Mathematical morphology is not only a theory for the analysis of spatial structures, but also a very powerful technique widely used for image processing purposes and employed successfully in biomedical image analysis, especially in preprocessing and segmentation tasks.
Microscopic image analysis and particularly malaria detection and classification can greatly benefit from the use of morphological operators.  The aim of this paper is to present a review of recent mathematical morphology based methods for malaria parasite detection and identification in stained blood smears images.
}

% Keywords
\keyword{malaria, red blood cells segmentation, mathematical morphology, medical image analysis}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences:
%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data:
%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately. If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal. In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}

%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%\setcounter{secnumdepth}{4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Only for the journal Gels: Please place the Experimental Section after the Conclusions

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% d

\section{Introduction}\label{s1}

Haematology is the branch of medicine concerned with the study, diagnosis, monitoring, treatment, and prevention of diseases related to the blood and blood-forming organs. Haematology~studies the blood in health and pathological conditions, firstly to identify the patient's health condition and, secondly, to predict how the bone marrow may have contributed to reach that~condition.

Thus, haematology studies the relationship between the bone marrow and the systemic circulation. In fact, there are many diseases, disorders, and deficiencies that can affect the number and type of produced blood cells, their function and their lifespan. Usually, only normal, mature or nearly mature cells are released into the bloodstream, but certain circumstances can induce the bone marrow to release immature and/or abnormal cells into the circulation. One of the most frequently ordered tests to monitor the proportion of the cell components into the blood stream is the Complete Blood Count (CBC), which~offers various haematologic data represented by the numbers and types of cells in the peripheral~circulation. The cells percentage is compared with the reference ranges in order to determine if the cells are present in their expected percentage, if one cell type is increased, decreased~or if immature cells exist. Reference ranges for blood tests are sets of values used to interpret a set of diagnostic test results from blood samples. Since it is difficult to prove that healthy-considered subjects may not have infections, parasitic infection and nutritional deficiency, it is more feasible to talk about reference ranges rather than normal ranges. A reference range is usually defined as the set of values in which 95\% of the normal population falls within. It is determined by collecting data from vast numbers of laboratory tests result from a large number of subjects who are assumed to be representative of the population. With~automatic counters or the flow cytometry, an automated CBC can be performed quickly. \mbox{However, if the} results from an automated cell count indicate the presence of abnormal cells or if there is a reason to suspect that abnormal cells are present, then a blood smear will be collected~\cite{Loddo2016}. A blood smear is often used to categorize and/or identify conditions that affect one or more types of blood cells and to monitor individuals undergoing treatment for these conditions. The results of a blood smear typically include a description of the cells appearance, as~well as any abnormalities that may be seen on the slide. The manual analysis of blood smears is tedious, lengthy, repetitive~and it suffers from the presence of a non-standard precision because it depends on the operator's skill. The~use of image processing techniques can help to analyse, count~the cells in human blood and, at~the same time, to provide useful and precise information about cells morphology.
Peripheral~blood smears analysis is a common and economical diagnosis technique by which expert pathologists may obtain health information about the patients. Although~this procedure requires highly trained experts, it is error-prone and could be affected by inter-observer variations. Moreover, blood cells' images taken from a microscope could vary in their illumination and colouration conditions, as~shown in Figure \ref{fig:images_types}. Typical~blood cells' images contain three main components of interest: the platelets (or~thrombocytes), the red blood cells (or erythrocytes) and the white blood cells (or leukocytes). It~is worth considering that blood cells exist with different shapes, characteristics and colourations, according to their types.
Many tests are designed to determine the number of erythrocytes and leukocytes in the blood, together with the volume, sedimentation~rate, and haemoglobin concentration of the red blood cells (blood~count). In addition, certain tests are used to classify blood according to specific red blood cell antigens, or blood groups. Other tests elucidate the shape and structural details of blood cells and haemoglobin and other blood proteins. Blood~can be analysed to determine the activity of various enzymes, or protein catalysts, that either are associated with the blood cells or are found free in the blood plasma.
Blood also may be analysed on the basis of properties such as total volume, circulation time, viscosity, clotting time and clotting abnormalities, acidity (pH), levels of oxygen and carbon dioxide, and the clearance rate of various substances. There~are special tests based on the presence in the blood of substances characteristic of specific infections, such as the serological tests for syphilis, hepatitis, and human immunodeficiency virus (HIV, the AIDS virus) \mbox{(\url{https://www.britannica.com/topic/blood-analysis})}. %the footnote has been moved to here, please confirm.
%Pls replace the website with a number in brackets from the References list, websites should not be cited in-text
Among the several available blood tests, the~most common are certainly the blood cells counts, \mbox{e.g., a CBC} is a measure of the haematologic parameters of the blood. Included in the CBC is the calculation of the number of red blood cells (red blood cell count) or white blood cells (white blood cell count) in a cubic millimetre (mm$^{3}$) of blood, a differential white blood cell count, a haemoglobin assay, a hematocrit, calculations~of red cell volume, and a platelet count. The differential white blood cell count includes measurements of the different types of white blood cells that constitute the total white blood cell count: the band neutrophils, segmented neutrophils, lymphocytes, monocytes, eosinophils, and basophils. A specific infection can be suspected on the basis of the type of leukocyte that has an abnormal value \cite{DiRuberto2016}.

Human malaria infection is not strongly related to cell count, but it needs different tests in order to be identified. It can only be caused by parasitic protozoans belonging to the \emph{Plasmodium} type. The~parasites are spread to people through the bites of infected female Anopheles mosquitoes, called~``malaria vectors''.
There are five parasite species that cause malaria in humans and two of these species, \emph{Plasmodium falciparum} and \emph{Plasmodium vivax}, constitute the greatest threat. \emph{Plasmodium~ovale}, \emph{Plasmodium malariae} and \emph{Plasmodium knowlesi} are the three remaining species that are less dangerous in humans \cite{WHO_dec_2016}, as shown in Figure \ref{fig:malaria_types}.
All five species may appear in four different life-cycle stages during the infection phase in peripheral blood: ring, trophozoite, schizont and gametocyte. Some~examples are shown in Figure \ref{fig:malaria_stages}.
The life-cycle-stage of the parasite is defined by its morphology, size and the presence or absence of malarial pigment.
The species differ in the changes of infected cell's shape, presence of some characteristic dots and the morphology of the parasite in some of the life-cycle-stages~\cite{Somasekar2011}.

\begin{figure}[H]
	\centering	
	\includegraphics[width=3.8cm]{img/f1a}
	\includegraphics[width=3.8cm]{img/f1b}
	\includegraphics[width=3.8cm]{img/f1c}
	\includegraphics[width=3.8cm]{img/f1d}
	\caption{\label{fig:images_types} Different illumination conditions generate different images because of the absence of a standardized acquisition procedure. From left to right: acquisition of the same smear with four microscope's brightness levels. Courtesy of CHUV, Lausanne.}

\end{figure}

\iftrue
\begin{figure}[H]
	\centering
	\includegraphics[width=3.8cm]{img/f2_Pfalciparum}
	\includegraphics[width=3.8cm]{img/f2_Pvivax}
	\includegraphics[width=3.8cm]{img/f2_Povale}
	\includegraphics[width=3.8cm]{img/f2_Pmalariae}
	\caption{\label{fig:malaria_types}Types of human malaria parasites: from left to right, \emph{P. falciparum} in its schizont stage, \emph{P. vivax} in two gametocytes specimens and one ring stage, \emph{P. ovale} in its ring stage, \emph{P. malariae} in its schizont stage.
	Courtesy of CHUV, Lausanne.}
\end{figure}
\fi

\iftrue
\begin{figure}[H]
	\centering
	\includegraphics[width=3.5cm, height=3.5cm]{img/falciparum_1_ring}
	\includegraphics[width=3.5cm, height=3.5cm]{img/falciparum_2_trophozoiteAge}
	\includegraphics[width=3.5cm, height=3.5cm]{img/falciparum_3_schizont}
	\includegraphics[width=3.5cm, height=3.5cm]{img/falciparum_4_gametocyte}
	\includegraphics[width=3.5cm, height=3.5cm]{img/ovale_1_ring}
	\includegraphics[width=3.5cm, height=3.5cm]{img/ovale_2_trophozoite}
	\includegraphics[width=3.5cm, height=3.5cm]{img/ovale_3_schizont}
	\includegraphics[width=3.5cm, height=3.5cm]{img/ovale_4_gametocyte}
	\includegraphics[width=3.5cm, height=3.5cm]{img/malariae_1_ring}
	\includegraphics[width=3.5cm, height=3.5cm]{img/malariae_2_trophozoite}
	\includegraphics[width=3.5cm, height=3.5cm]{img/malariae_3_schizont}
	\includegraphics[width=3.5cm, height=3.5cm]{img/malariae_4_gametocyte}
	\caption{\textit{Cont.}}
\end{figure}
\begin{figure}[H]\ContinuedFloat
\centering	
	\includegraphics[width=3.5cm, height=3.5cm]{img/vivax_1_ring}
	\includegraphics[width=3.5cm, height=3.5cm]{img/vivax_2c_trophozoiteDeveloped}
	\includegraphics[width=3.5cm, height=3.5cm]{img/vivax_4_gametocyte}
	\caption{\label{fig:malaria_stages}Examples of malaria parasite stages.
		First row, from left to right: \emph{P. falciparum} ring, trophozoite, schizont, gametocyte;
		second row, from left to right: \emph{P. ovale} ring, trophozoite, schizont, gametocyte;
		third row, from left to right: \emph{P. malariae} ring, trophozoite, schizont, gametocyte;
		last row, from left to right: \emph{P. vivax} ring, developed trophozoite, gametocyte.
 Courtesy of CHUV, Lausanne.}
\end{figure}

Computer vision techniques for malaria diagnosis and recognition represent a relatively new area for early malaria detection and, in general, for medical imaging, able to overcome the problems related to manual analysis, which is performed by human visual examination of blood smears. The~whole process requires an ability to differentiate between non-parasitic stained components/bodies (e.g.,~red~blood cells, white blood cells, platelets, and artefacts) and the malarial parasites using visual information. If the blood sample is diagnosed as positive (i.e.,  parasites present), an additional capability of differentiating species and life-stages (i.e.,  identification) is required to specify the infection. Numerous~methods of automatic malaria diagnosis have been proposed so far, in order to overcome the issues before mentioned. The aim of this paper is to review and analyse the works of different researchers who in particular have used mathematical morphology as a powerful tool for computer aided malaria detection and classification.

\fi
\subsection{Mathematical Morphology}
Mathematical morphology (MM) can be defined as a theory for the analysis of spatial structures. It is called morphology because it aims at analysing the shape and form of objects. It is mathematical in the sense that the analysis is based on set theory, integral geometry, and lattice algebra. MM is not only a theory, but also a very powerful image analysis technique \cite{Soille2004}.
It was introduced by Matheron in 1964 as a technique for analysing geometric structure of metallic and geologic samples. It refers to a branch of nonlinear image processing and analysis that concentrates on the geometric structure within an image.
The morphological filters, which can be constructed on the basis of the underlying morphological operations, are more suitable for shape analysis than the standard linear filters since the latter sometimes distort the underlying geometric form of the image. Some of the salient points regarding the morphological approach are as follows \cite{Giardina1988}:
\begin{itemize}[leftmargin=*,labelsep=5.8mm]
	\item Morphological operations provide for the systematic alteration of the geometric content of an image while maintaining the stability of the important geometric characteristics.
	\item There exists a well-developed morphological algebra that can be employed for representation and optimization.
	\item It is possible to express digital algorithms in terms of a very small class of primitive morphological~operations.
	\item There exist rigorous representation theorems by means of which one can obtain the expression of morphological filters in terms of the primitive morphological operations.
\end{itemize}

MM was initially developed for binary images and later on generalized to grayscale images \cite{Soille2004,Serra1984}, considered as a sampled function of $\mathbb{R}^{2}$ in $\mathbb{R}$, or in general of any function of $\mathbb{R}^{n}$ in $\mathbb{R}$.
More recently, several researchers have extended morphological operators to colour (or in general multispectral) images, considered as sampled functions of $\mathbb{R}^{n}$ in $\mathbb{R}^{m}$, with $m$ being equal to three in the case of the usual colour images or to the number of bands otherwise \cite{Benavent2012}.  Moreover, several approaches for fuzzifying MM have been proposed, extending the ordinary  morphological operations by using fuzzy sets \cite{Kerre2000}.

Dilation and erosion are the basic morphological processing operations. They are defined in terms of more elementary set operations, but are employed as the basic elements of many algorithms. Both~dilation and erosion are produced by the interaction of a set called structuring element (SE) with a set of pixels of interest in the image. The structuring element has both a shape and an origin. From these two basic operators, others have been derived (opening, closing, hit-or-miss). They can be applied to extract image components useful in the representation and descriptions of region shapes, such as area granulometry, boundaries, skeleton, or convex hull. In addition, morphological operators can be used for image preprocessing and postprocessing, such as morphological filtering, thinning, and~especially for segmentation.

\section{Scope of This Review}\label{s2}
In this paper, we present a review of computer-aided methods oriented to malaria parasite detection and segmentation by MM based techniques.
Most of the studies have followed Di Ruberto's work \cite{DiRuberto2002}, which first proposed a system to evaluate parasitaemia in the blood. The system was able to detect the parasites by using an automatic thresholding and morphological operators. A~morphological approach to cell segmentation, which is more efficient than the watershed algorithm \cite{Soille2004}, was proposed. Finally, the parasites' classification was still based on morphological operators. Since then, many systems for computer aided diagnosis of malaria have been proposed. Most of them make use of MM to process and analyse malaria-infected peripheral blood cells images. The scope of this paper is to review and analyze the recent works of different researchers in the area of malaria parasite recognition using computer vision that benefits from MM. Only few reviews exist in literature about microscopic image processing for malaria parasites. However, they are not focused on MM techniques as they analyze generic computer vision systems for malaria diagnosis, as in \cite{Rosado2017,Das2015,Tek2009}. Furthermore, newer~and promising approaches, addressing for example the problem of malaria diagnosis in remote areas~\cite{Rosado2017,Oliveira2017}, or improving significantly both the detection and the classification performances~\cite{Dave2017,Devi2017,Somasekar2017,Sulist2015}, have not been considered in the previous reviews.

The rest of the paper is organised as follows. Section \ref{s3} presents a review of the considered works, according to a typical pipeline of a computer-aided image analysis process: preprocessing, segmentation, feature extraction.
All of the considered works make use of morphological operators in at least one of the phases of image analysis. Section \ref{s4} contains an overall discussion about the methods and the conclusions are expressed in Section \ref{s5}.

\section{Computer Aided Diagnosis of Malaria by Using Mathematical Morphology}\label{s3}
This section presents a review of some of the main recent studies existing in literature regarding the analysis of malaria infected blood smears using MM.

Extensive search of articles has been made in PubMed and Google Scholar search engines based on the keywords: ``malaria'', ``mathematical morphology'', ``automated malaria diagnosis'' up to October 2017. The search includes papers published in English language. The following criteria were used to evaluate the relevance and suitability of an article in our study. \mbox{Firstly, every single} article has been chosen if it had a JCR
rank value, according to Thomson Reuters website \mbox{(\url{https://jcr.incites.thomsonreuters.com/})}. %the footnote has been moved to here, please confirm.
%Pls replace the website with a number in brackets from the References list, websites should not be cited in-text
 In that case, the work has been considered suitable for our purposes. However, JCR value was not available for all the considered articles, so that we selected the ones with an SJR rank value 
(available from Scimago website \mbox{(\url{http://www.scimagojr.com/})}). The~remaining articles, more precisely \cite{Somasekar2011,Oliveira2017,Somasekar2017,Khan2011,Savkare2011a,Savkare2011b}, have been considered and included because of their innovativeness or because the results were particularly interesting and promising. Furthermore,~some well known books, like \cite{Soille2004,Serra1984,Kerre2000,Das2014}, were included, as well as Schulze's Ph.D thesis \cite{Schulze1994} and {\v{S}}pringl's masters thesis \cite{Springl2009}. Thereafter, a conference paper was included if the related conference owns a SJR impact factor. Finally, the full texts of these studies were analysed and~evaluated.

A typical approach of a computer-aided method usually comprises four different image processing and analysis tasks, as follows:
\begin{enumerate}
	\item Preprocessing.
	\item Segmentation.
	\item Feature extraction.
	\item Classification.
\end{enumerate}

Since morphological techniques have been used in the first three phases, the reviewed works have been divided into the following sub-sections: preprocessing, segmentation and feature extraction.
Each sub-section contains description about methods that cope with malaria parasites (MP) stained components analysis, both on thin and thick blood smears, without distinction.

%\subsection{Blood smear staining techniques}
Two main factors are generally considered if we refer to staining techniques: the type of colouration, in which Giemsa and Leishman are the most common, and the thickness of blood slide, which may be thick or thin.
The majority of studies have been employed on thin blood smear images (over two-third of the total count) while only a few have used thick blood smear images.
\mbox{Typically, thin smears} permit the identification of specific parasitic stage and quantification of malaria parasite; on the other hand, thick smears are better if the target is to perform an initial identification of malaria infection using blood pathology. Some examples are shown in Figure  \ref{fig:thick_thin_film}.
Giemsa stained blood smear is considered in most of the analysed literatures, whereas the Leishman stain is considered in few studies. %, as showed in table \ref{tab:summary_table}.
It is reported that the Leishman stain has more sensitivity for parasite detection than Giemsa \cite{Khan2014} and is superior for visualization of red and white blood cell morphology \cite{Sathpathi2014}. On the contrary, Giemsa~stain highlights both malaria parasites and white blood cells and, therefore, it is an additional issue to deal with.
The Giemsa stain is a more costly and also time-consuming procedure than Leishman.
Moreover, magnification of 100X by using an oil immersion objective is used for capturing microscopic images of thin blood smear for identification of specific parasites and their infected stage.

\begin{figure}[H]
	\centering
	\includegraphics[height=0.17\textheight]{img/f3_ThickGiemsa}
	\includegraphics[height=0.17\textheight]{img/f3_ThinGiemsa}
	\includegraphics[height=0.17\textheight]{img/f3_ThickLeishman}
	\includegraphics[height=0.17\textheight]{img/f3_ThinLeishman}
	\caption{\label{fig:thick_thin_film}Malaria infected blood smears types. This image shows a comparison between staining colouration procedures and smears thickness.
	On top, left: thick smear with Giemsa stain \cite{Sathpathi2014}, right: thin smear with Giemsa stain (courtesy of CHUV). On bottom, left: thick smear with Leishman stain \cite{Sathpathi2014}, right: thin smear with Leishman stain (courtesy of CHUV). Dots in thick smears and rings in thin smears are P. Falciparum ring stages, while elongated erythrocytes (in images on the right) are affected from P. Falciparum in its trophozoite schizont stage. The difference between thick and thin smears is clearly evident by observing cells and parasite shapes. Thin smears typically offer a better shape representation, while thick ones contain smaller and less clear region shapes. Furthermore, Giemsa stain shows a better contrast between cells, parasites and background respect to Leishman stain.}
\end{figure}

\subsection{Preprocessing}
In an image analysis field, especially when we refer to complex computer-aided pipelines, preprocessing methods are particularly used in order to improve the image data by suppressing unwanted noise or enhancing some image features for further processing.
It is worth mentioning preprocessing methods because they are an important step regarding the image analysis field, but, for what concerns the malaria-affected blood image analysis, in our review, we particularly found methods that operate for illumination correction and noise filtering purposes.
Generally speaking, digital~microscopy images can be acquired in different lighting conditions, with several types of acquisition devices or from blood smears stained with various staining protocols, and, consequently, the features of similar images could differ a lot.
Different techniques for illumination correction have been suggested to reduce such variation, e.g., a lot of authors work with grayscale-converted images as an illumination correction method.
On the other hand, noise filtering aims to remove the noise introduced by mishandling the slides and/or the camera settings.
Morphological operators have been extensively used as preprocessing for image enhancement in major studies.
Erosion and dilation operations on raw smear images allow for discarding undesired patterns and help in the selection of required cells or regions of interest. Morphological operators are useful for removal of unwanted
objects, holes filling, splitting, thinning and thickening. Different researchers during automated diagnosis of malaria used morphological operations in the preprocessing phase, and the most recent are listed below.

In \cite{Gonzalez2016}, Gonzalez-Betancourt et al. proposed a system to determine markers for watershed segmentation based on the Radon transform and mathematical operators. In the first step of the process, small irrelevant structures and part of the noise are eliminated by a morphological filter, in~order to ensure the preservation of the cells edges. Image smoothing is performed by a morphological erosion-reconstruction and dilation--reconstruction filter with a disk structuring element of a radius equal to 20 pixels, which is $0.274$ times smaller than the average radius of the RBCs.

In this way, the influences of the size and the shape of the structures can be separated in the smoothing process. At the same time, the objects that are not eliminated remain unchanged. In addition, a morphological closing is performed with a disk structuring element having a radius smaller than half the average of the RBCs radii, in order to connect the possible (more than one) markers that can appear on a single cell.

In \cite{Kareem2012}, Kareem et al. illustrated a morphological approach for blood cell identification and use the image features such as intensity, histogram, relative size and geometry for further analysis. Before~the identification of blood cells, the authors  propose a novel morphological filtering based on the size of RBCs for platelets and/or artifacts elimination. A dilation is performed by a concentric ring structuring element and erosion by a disk-shaped structuring element. The radius of the structuring element depends on the radius of the RBCs, so that all the components smaller than the RBCs can be removed.

The system proposed in \cite{Oliveira2017} by Oliveira et al. is based on image processing, artificial intelligence techniques and an adapted face detection algorithm to identify Plasmodium parasites. The latter uses the integral image and haar-like features concepts, and weak classifiers with adaptive boosting learning. The search scope of the learning algorithm is reduced in the preprocessing step by removing the background around blood cells by means of morphological erosions both for training and for~testing.

Romero-Rondon et al. in \cite{Romero2016} presented an algorithm that uses morphological operations, the watershed method, the Hough transform and the clustering method of k-means to detect overlapped RBCs. In the preprocessing stage, white blood cells and platelets are removed before the segmentation task. During this step, some noise, the WBC cytoplasm and platelets still remain on the image.

Therefore, the small objects are removed using a morphological opening and then the image is dilated with a disk-shaped structuring element.

Reni et al. in \cite{Reni2015} described a new algorithm for morphological filtering of the blood images as a preprocessing tool for segmentation. Conventional morphological closing on blood images removes the unwanted components but also useful information. On the contrary, the proposed method preserves the necessary information of foreground components while removing noise and artefacts.

In the method proposed in \cite{Sheik2013} by Sheikhhosseini et al., the first phase is the stained object extraction that detects candidates' objects that can be infected by malaria parasites using intensity and colour. Before detecting the stained objects, the method firstly extracts the foreground. The foreground image is a binary image that is produced after applying morphological hole filling on such pixels that have lower intensity value than average intensity value of the green layer. After the stained objects' extraction process, a series of morphological operations is also employed in order to eliminate small components and complete the final stained objects.

An edge-based segmentation of erythrocytes infected with malaria parasites using microscopic images is proposed by Somasekar et al. in \cite{Somasekar2015}.  A fuzzy C-means clustering is applied to extract infected erythrocytes, which is further processed for the final segmentation. A morphological erosion is used to erase some small noises and spots before the segmentation and holes inside the infected erythrocytes are filled using a morphological hole filling operation for the final segmentation.

In \cite{Tek2010}, Tek et al. presented a complete framework to detect and identify malaria parasites in images of Giemsa stained thin blood film specimens. In addition, the system is able to identify the infecting species and life-cycle stages.
The preprocessing step of the proposed method is applied to reduce the variations in the observed size, intensity, and colour of the cells and stained objects before the detection and classification steps. The aim is to correct the non-uniform illumination in the images. The estimation is based on a morphological closing operation using a sufficiently large structuring element. The sufficiently large size for an input image is determined automatically with respect to its average cell size computed from the area granulometry distribution.

%Median filter is often used for reducing impulse noise. Several studies have used it to enhance microscopic images of peripheral blood smear towards characterization of malaria followed by adaptive or local histogram equalization. Local low pass filter and local adaptive histogram equalization techniques have also been applied to enhance the pathological image quality. Das et al. \cite{Das2015} showed that geometric mean filter provides better performance towards enhancing peripheral blood smear images. Another preprocessing method like space-time manifolds-based image mosaicing has been used to obtain large field of view \cite{Zou2010} towards identifying malaria parasites, while Di Ruberto et al. \cite{Put16c} proposed an approach to overcome the problem of uneven illumination conditions in image acquisition. For this purpose they designed an illumination pattern that simulates the classic visual defects introduced by the digital microscope lenses, that
%is the vignetting effect. Starting from the smallest radius, they applied different illumination patterns, created modifying the radius of the Gaussian curve, to the original images. Then, a similarity value measures the difference in terms of pixels between the original images and the corrupted ones.

\subsection{Segmentation of RBCs and Parasites}
Segmentation is a key step in image analysis because it permits the identification and separation of the regions that compose an image, according to certain criteria of homogeneity and separation. Its~main target is to divide the image into parts that have a strong correlation with objects or areas of the real world contained in the image.
The commonly used segmentation methods essentially operate considering characteristics such as the brightness value, colour and reflection of the individual pixels, identifying groups of pixels that correspond to spatially connected regions. As for many problems of image processing, there is no standard solution valid in general, so different segmentation techniques can be applied, according to the characteristics of the images to process and of the objects to segment.
Medical images segmentation is typically performed using two main strategies: the first level aims to separate whole cells or tissues from the background and the second one aims to separate the tissue structure in different regions or the cell in their components, as the nucleus from the cytoplasm or intracellular parasites. The latter case is commonly used in applications in which the cell class depends on the morphological characteristics of its components.

%%% THRESHOLDING METHODS
Several other authors attempted to use thresholding combined with morphological operation as a segmentation method in their computer-aided systems, and they are described as follows.

Arco et al. in \cite{Arco2014} worked on thick blood films and proposed a method that uses an adaptive thresholding based scheme, which also allows an effective classification of pixels. This means that the election of whether a pixel belongs to the background or to the signal (parasites and white blood cells) is only established by the pixels around it, that is its neighbourhood. Then, morphological methods are applied to evaluate the area of connected components, labelling those belonging to parasites and counting their number.

Anggraini et al. \cite{Anggraini2011} proposed a method for separating blood cells, parasites and other components from the background in a microscopic field of a thin blood smear. They applied several global thresholding methods and visually compared the results to qualitatively determine which technique yields the best result. The binary image was then subjected to a hole filling morphological operator and applied as a marker to label blood cells. From each identified cell (RBC and WBC), constituents of the parasite (nucleus and cytoplasm) were extracted using multiple thresholds.

Dave et al. in \cite{Dave2017} performed image segmentation using histogram based adaptive thresholding followed by mathematical morphological operations (erosion and dilation). The detection of infected RBCs is based on an unsupervised learning technique.

The automated method proposed in \cite{Elter2011} by Elter et al. for parasite detection and identification worked on thin blood film acquired with Giemsa stain. The authors found that the G and B channels of the RGB colour are very good features to identify objects containing chromatin in Giemsa stained blood films, being not only considered highly discriminative but also almost independent of differences in illumination and staining intensity. They transformed the colour input image into a monochrome image \textit{I(x,y)}, which highlights objects containing chromatin: $ I(x,y) = arctan \frac{I_{green}(x,y)}{I_{blue}(x,y)} $. In this work, mathematical morphology has been used with a black top-hat operator to separate MP from both leukocytes and platelets, with a non-flat paraboloid structuring element with a radius of 9 and a slope of one~pixel. It should be taken into account that these fixed parameters might not be suitable for images with different pixel resolutions. The black top-hat operator is followed by a thresholding operation with a fixed threshold, which, according to the authors, is reliable given the independence of the G and B channels with regard to illumination and staining intensity. However, the authors do not define the value of this fixed threshold on the publication.

Kareem et al. in \cite{Kareem2011} used the Annular Ring Ratio transform method. Before applying it, a~preprocessing phase for removing platelets, parasites and other artefacts in the image has been performed. In the proposed method, the image after being converted to grayscale undergoes a morphological opening similar to closing. Unlike conventional closing (dilation followed by erosion), which uses the same structuring element, two different structuring elements are used: a concentric ring for dilation and a disk for erosion. The inner and outer diameter of the dilation ring is set to 35\% and 70\% of RBCs size, respectively, and the erosion disk has the same diameter. Therefore, considering that fixed manually defined parameters are used for this strategy, the results may substantially differ depending on the image resolution. This approach results in locating only the stained components in the image instead of all the cells and hence will not only speed up the operation but reduce the~complexity.

Mushabe et al.~\cite{Mushabe2013} used morphological and statistical classification to detect malaria in blood smears by identifying and counting red blood cells and Plasmodium parasites. Morphological~operations and histogram-based thresholding are used to extract RBCs
and boundary curvature calculations and Delaunay triangulation are used for splitting clumped RBCs. They worked on Giemsa-stained thin blood smears.

In \cite{Ross2006}, Ross et al. proposed a method that provides a positive or negative diagnosis of malaria and differentiates parasites by species. The segmentation step relies on a six step thresholding selection strategy. It aims to identify and segment potential parasites and erythrocytes from background. Mathematical morphology has been used in several key steps of the procedure. Hole~filling is used in the first step in order to fill RBCs' binary masks obtained from a first thresholding. \mbox{Afterwards, step 4} employs RBCs' morphological reconstruction with parasites' mask, found in step 2, for identifying infected cells. In step 5, a morphological opening filter, using a disk-shaped SE with a radius equal to the mean erythrocyte radius less the standard deviation, is applied to the grayscale, morphologically filtered, green component in order to remove any objects smaller than an erythrocyte. The morphological gradient (difference between a dilation and erosion of the image) is then calculated using a diamond-shaped SE with unity length.
Finally, in step 6, the intersection of morphological gradient image and the dilated cell cluster is calculated. This image is then transformed to a binary image by thresholding any value greater than zero. A series of morphological operations, namely a closing operation, thinning, and spur-removal are then applied to generate a contour of the segmented erythrocytes. Contours are filled, and the segmented mask is again reconstructed with the valid parasite marker image to result in a segmented mask of infected cells. RBCs and parasites masks are consequently ready for the next generation step.

Savkare et al. \cite{Savkare2011b} worked on thin blood films with Giemsa staining and used a global threshold and Otsu threshold \cite{Otsu1975} on grayscale enhanced image (green channel) for separating foreground from background. Hole filling has been performed on identified cells, and morphological operators have been used to identify overlapping cells. Then, a watershed transform has been applied for separating overlapped cells.

In addition, in the method proposed in \cite{Somasekar2017} by Somasekar et al., the segmentation of the infected parasites is based on thresholding. It is achieved in two stages by maximizing between-class variance of an original image and consequently by an iterative threshold selection from a stage-one threshold image with suitable stopping criteria. The segmented results are postprocessed to improve the accuracy of malaria parasites detection by morphological operators (erosion and closing).

% Morphology and/or granulometry in Segmentation stage:
On the other hand, a lot of works have been realized by means of mathematical morphology and/or granulometry in the segmentation stages, even in combination with thresholding strategies. They are briefly analysed below.

Ahirwar et al. \cite{Ahirwar2012} based their approach on thresholding and granulometry. The histogram of the complemented green component has been used, and it is said to be a bimodal distribution in all the considered images. Then, both local and global thresholds are used, and the union of the two binary images is chosen as the parasite marker image. A morphological opening filter, using a disk-shaped SE with radius equal to the mean erythrocyte radius less the standard deviation, is applied to the grayscale morphologically filtered green component of the image to remove any objects smaller than an erythrocyte. The morphological gradient is then calculated using a diamond-shaped SE with unity length. The segmentation method is applied to each object in the reconstructed binary image of erythrocytes individually. Those objects that do not exceed the area of a circle with a radius equal to the mean erythrocyte radius plus the standard deviation are regarded as being single cells, and are unmodified. On the other hand, the clumped cells are segmented as follows. First, the intersection of the morphological gradient image and the dilated cell cluster is taken. This image is then transformed to a binary image by thresholding any value greater than zero. A series of morphological operations, namely a closing operation, thinning, and spur removal are then applied to generate a contour of the segmented erythrocytes. The contours are filled, and the segmented mask is again reconstructed with the valid parasite marker image to result in a segmented mask of infected cells.

Di Ruberto et al. \cite{DiRuberto2002} aimed to detect the parasites by means of an automatic thresholding based on a morphological approach applied to cell image segmentation, which is more accurate than the classical watershed-based algorithm. They applied grey scale granulometries based on opening with disk-shaped elements, flat and hemispherical. They used a hemispherical disk-shaped structuring element to enhance the roundness and the compactness of the red cells improving the accuracy of the classical watershed algorithm, while they have used a disk-shaped flat structuring element to separate overlapping cells. These methods make use of the red blood cell structure knowledge, which is not used in existing watershed-based algorithms.

Khan et al. in \cite{Khan2011} presented a novel threshold selection technique used to identify erythrocytes and possible parasites present on microscopic slides that greatly benefits from morphological operations, such as granulometry and morphological reconstruction.

In \cite{Rosado2017}, Rosado et al. proposed a system using supervised classification to assess the presence of malaria parasites and determine the species and life cycle stage in Giemsa-stained thin blood smears. For the RBCs segmentation, they used an adaptive thresholding approach followed by a closing morphological operation with an elliptical structuring element.

Soni et al. \cite{Soni2011} performed segmentation of erythrocytes by using granulometry as well. The~size and eccentricity of the erythrocytes are also required for the calculation of some feature values (as~these can be indicative of infection). The shape of the objects (circular erythrocytes) is known a priori, but the image must be analysed to determine the size distribution of objects in the image and to find the average eccentricity of erythrocytes present.
Grayscale granulometries based on opening with disk-shaped elements are then used. Non flat disk-shaped structural elements are applied to enhance the roundness and compactness of the red blood cells and flat disk-shaped structural elements applied to segment overlapping cells. The object to be segmented differs greatly in contrast to the background image. Changes in contrast can be detected by operators that calculate the gradient of an image. The gradient image can be computed and a threshold can be applied to create a binary mask containing the segmented cell. The binary gradient mask is dilated using a vertical structuring element followed by a horizontal structuring element. The cell of interest has been successfully segmented, but~it is not the only object that has been found. Any objects that are connected to the border of the image can be removed.

In Tek et al. \cite{Tek2010}, the localisation of the parasites is achieved after a foreground and background segmentation step. Firstly, a rough foreground image using morphological area top-hats (using the average cell area value) is extracted. Then, from these rough foreground and background regions, two different threshold values are determined and used in morphological double thresholding of the input grey level image to produce a refined binary foreground mask. From the foreground image, the stained pixels are detected using again a thresholding approach and finally used as markers to extract the stained objects by morphological area top-hats based on the estimated average area value.

In \cite{Yunda2012}, Yunda et al. proposed a method for \textit{P. vivax} parasite detection. The segmentation phase is a combination of border and region detection that allows rejection of the image background and permits identifying each of the objects. Initially, the morphological gradient method is used to enhance the borders of previously found objects. This is followed by a threshold detection stage using the K-Median method.
Furthermore, a Laplacian operator was used to discriminate the pixels that are interior or exterior in relation to the regions of the images and then erosion operation followed by two dilations were applied to delete the pixels that did not make up part of any object. In the end, Absence~of Gradients and Nernstian Equilibrium Stripping (AGNES) and K-Median techniques were applied to assign the remaining number of pixels to each region, using the image regions previously identified as objects and background as the starting point.

Several authors used marker controlled watershed \cite{Soille2004} with morphological approach, as~described in the following.

Das et al. in \cite{Das2011,Das2013,Das2014,Das2015} segmented erythrocytes as aforesaid and then morphological operators \textls[-20]{are used to eliminate unwanted cells like leukocytes and platelets. To conclude, overlapping~erythrocytes} are segmented by using a marker controlled watershed segmentation technique.

In \cite{Devi2017}, Devi et al. proposed a computer assisted system for quantification of erythrocytes in microscopic images of thin blood smears. The performance of the system in classifying the isolated and clump erythrocytes by geometric features is evaluated for the different classifiers. The clump erythrocytes are segmented using marker controlled watershed with h-minima as internal marker.

In \cite{Dey2015}, Dey et al. presented an automatic system for segmenting platelets, useful for identifying disease as malaria, using a colour based segmentation and mathematical morphology (opening~operations with a disk element of radius 2).

In the study presented in \cite{Diaz2009} by Diaz et al. for quantification and classification of erythrocytes in stained thin blood films infected with \textit{Plasmodium Falciparum}, the authors used connected morphological operators in the segmentation step. The RBCs are detected as follows: firstly, a pixel classification allowed for labeling each image pixel as either background or foreground, based on its colour features. Afterwards, an inclusion-tree structure is used to represent the hierarchical object relations between background and foreground so that a filtering process allows for removing irrelevant structures such as artefacts generated at the staining or digitization processes.

Khan et al. \cite{Khan2011}, among other experimentations, used it in order to try to separate overlapping cells because, according to their statements, watershed transform can separate touching cells, but it is not sufficient for overlapping cells.

In the algorithm described by Romero-Rondon et al. in \cite{Romero2016}, the detection of overlapped RBCs is still based on marker-controlled watershed transform. To define the suitable markers in watershed transform, they used three different approaches, based on a morphological erosion operation, on Hough transform and on a clustering method of K-means.

\textls[-17]{Savkare et al. in \cite{Savkare2015} segmented cells using K-mean clustering and global threshold. Overlapping~cells} are separated using a Sobel edge detector and watershed transform. Watershed transform is applied on each cluster separately. Over-segmentation is minimized by a series of morphological operations, like~erosion and dilation utilizing disk-shaped structuring elements.

In \cite{Savkare2011a}, an approach  to detect red blood cells with consecutive classification into parasite infected and normal cells for further estimation of parasitemia is proposed. For separation of overlapping cells, watershed transform is applied on a distance transform of binary mask of cells having a larger area.

In \cite{Springl2009}, {\v{S}}pringl performed red blood cell segmentation by using marker-controlled watershed transformation based on the image gradient. Markers are computed as a combination of the binary mask of the red blood cells and centres of the cells that are computed using a similar algorithm that was utilized for the evaluation of the average cell radius. The binary mask is obtained by thresholding the grayscale image with an automatically estimated threshold using the Otsu method \cite{Otsu1975}.

In \cite{Sulist2015}, Sulistyawati et al. combined morphological operations (erosion, dilation, opening and closing) and blob analysis to segment and identify malaria parasites with a high degree of accuracy.

Tek et al. in \cite{Tek2006} proposed a classifier-based method for the segmentation stage, which relies on a Bayesian pixel classifier to distinguish between stained and non-stained pixels. In particular, they used a non-parametric method based on histograms in order to produce the probability density functions of stained and non-stained classes. Stained pixels can belong to other components such as WBCs, platelets or artefacts, in addition to the parasites, and so the detection procedure requires a further classification to distinguish among parasite and non-parasite pixels. However, the stained pixels have to be represented as connected sets, representing stained objects, to extract features for the classifier. Furthermore, top-hat extraction and infinite reconstruction were applied to find the regions that include the objects.

To sum up, the analysed systems use mathematical morphology methods during or immediately following the segmentation step for the following purposes:

\begin{itemize}
	\item size evaluation of regions obtained from segmentation
	\item image cleaning and artefacts removal
	\item RBCs and parasites segmentation or separation
	\item improving parasites detection
	\item minimization of watershed over-segmentation.
\end{itemize}

%%BOUNDARY BASED METHODS \subsubsection{Boundary based methods}

\subsection{Feature Extraction}
Feature extraction has the aim of reducing the computational complexity of the subsequent process and facilitating a reliable and accurate recognition for unknown novel data, considering~that the input data to an algorithm could be too large to be processed and it could be redundant (e.g., repetitiveness of pixels patterns in an image). Moreover, the in-depth understanding of the domain-specific knowledge gained by human experts on the problem being addressed can be of extreme importance for the design of a reliable and effective feature extraction engine \cite{Jiang2009}.
It starts from determining a subset of the initial features, and this procedure is called feature selection. The~selected features are expected to contain the relevant information from the input data, so that the desired task can be performed by using this reduced representation instead of the complete initial data.
Malaria~parasite infection causes micro structural changes in erythrocytes. The microscopic features of the RBCs are usually specific to morphology, intensity and texture. They may also represent the differences that occur among healthy and unhealthy cells. Most of the studies have reported both textural and geometric features for describing malaria infection stages \cite{Das2015}.
Generally speaking, features may be distinguished according to the following characteristics: morphological features and textural and intensity features.

It is a well known mathematical morphology approach to compute a size distribution of grains in binary images, using a series of morphological opening operations. It is the basis for the characterization of the concept of size. Some authors used area granulometry for preprocessing purposes in malaria characterization \cite{Tek2010}, even though it is certainly effective for extracting cell size features' information \cite{Springl2009,Tek2006,Malihi2013}. In \cite{Tek2010}, local area granulometry combined with colour histogram are used as features. The area granulometry feature is calculated locally on the binary mask of the stained objects, for the RGB channels, and then concatenated. Morphological features are also used
%in \cite{Ross2006} (erosion dilation),
in \cite{Das2011} (opening, closing) and in \cite{DiRuberto2002} (skeleton) to classify parasites.

%\input{Test.tex}

\section{Discussion}\label{s4}
%A comprehensive synthesis of the analysed works is presented in table \ref{tab:summary_table}.
In the review, we have only considered the methods that employed mathematical morphology in at least one step of the pipelines, and it has been structured by considering the following information: preprocessing, segmentation, features extraction.
Most of the studies are based on \emph{P. vivax} and/or \emph{P. falciparum} characterization.
With regards to the approaches shown and related results, it is clear that malaria parasite detection and segmentation techniques in microscopic images need further experiments and improvements. In general, the analysed works have been tested with a limited number of images and the datasets are not publicly available; therefore, a comparison between different approaches is very difficult. Despite promising results reported during the past few years, the~great majority of the computer-aided methods found on the 
%Should this be "past few" or "past several"?
literature for malaria diagnosis are based on images acquired under well controlled conditions and with proper microscopic equipment. \mbox{However, one should} take into account that 80\% of malaria cases occur in Africa, where this type of equipment is scarce or even nonexistent in common healthcare facilities \cite{Rosado2017}.
\mbox{Moreover, this review} showed that \emph{P. falciparum} is the most analysed if we refer to segmentation and detection, considering~that it is the most widespread among malaria parasite types.
The majority of the works used thin blood smear. It is typically used for identification of malaria infected stages, types~of parasitic infection and percentage of parasitemia, while thick blood smear is used for identification and quantification of malaria parasite count against leukocyte count per microliter of blood.

The preprocessing phase is typically taken on with filters and the most used in the analysed works is certainly the median filter, which permits preserving sharp edges. Apart from the classic histogram equalization and contrast stretching techniques, other filters have been employed, e.g., geometric mean filter to remove Gaussian noise preserving edges, Laplacian filter in order to find edges, and so on.
A median filter has been found to be effective for reducing impulse noises from the microscopic images, even though recent studies have shown that geometric mean filter provides better performance than the median filter \cite{Das2013,Das2015}. However, morphological operators have been greatly used with successful performances, imposing themselves as powerful alternatives to more common and used techniques for image enhancement and noise filtering \cite{DiRuberto2002,Gonzalez2016,Kareem2011,Kareem2012,Malihi2013,Mushabe2013,Oliveira2017,Reni2015,Romero2016,Ross2006,Sheik2013,Somasekar2015,Springl2009,Tek2010}.

Malaria parasites may be discriminated according to two different strategies: by segmenting the whole erythrocyte from the blood smear image on the basis of which malaria infection is detected, or by segmenting chromatin dot or parasite infection regions for characterizing parasite infection stages based on some extracted target features.
In general, a thresholding-based approach is still widely used for segmentation purposes. In particular, a lot of authors affirm that Otsu thresholding suffers from limitations when textural variation is high, while histogram thresholding can not deal sufficiently well in identifying valley regions in the case of unimodal histograms. However, such a simple and fast approach can greatly benefit from mathematical morphology as recent studies demonstrate~\cite{Anggraini2011,Arco2014,DiRuberto2002,Elter2011,Kareem2011,Mushabe2013,Rosado2017,Ross2006,Savkare2011b,Savkare2015,Somasekar2017,Tek2010}.

% start a new page
\newpage
% change it to landscape
\paperwidth=\pdfpageheight
\paperheight=\pdfpagewidth
\pdfpageheight=\paperheight
\pdfpagewidth=\paperwidth
\newgeometry{layoutwidth=297mm,layoutheight=210 mm, left=2.7cm,right=2.7cm,top=1.8cm,bottom=1.5cm, includehead,includefoot}
\fancyheadoffset[LO,RE]{0cm}
\fancyheadoffset[RO,LE]{0cm}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

 \begin{table}[H]
\centering
\tablesize{\footnotesize} 
\caption{Summary of analysed methods: morphological operations used in the main phases of analysis, type of MM (gray or binary)/type and size of SE (if reported), kind of classification and performance measures (Sensitivity, Specificity, Accuracy, if reported).}

\label{t1}
\begin{tabular}{m{3cm}<{\centering}m{3cm}<{\centering}m{5cm}<{\centering}m{3cm}<{\centering}m{5cm}<{\centering}m{2cm}<{\centering}}
		\toprule
    	\textbf{Authors} & \textbf{Preprocessing} & \textbf{Segmentation} & \textbf{MM/SE} & \textbf{Classification} & \textbf{Performance}\\  \midrule
    Ahirwar et al., 2012  &
    	- &
    	thresholding + granulometry, opening, morphological gradient, dilation, closing, thinning, spur removal &
    	gray, bin/disk of size depending on RBCs, diamond of size = 1 &
    	five (\emph{P.falciparum}, \emph{P.vivax}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) &
    	-  \\
\midrule
    Anggraini et al., 2011 &
		- &
		thresholding + hole filling &	
		bin &
		two (\emph{P.falciparum} infected and noninfected) + two life-cycle-stages &
		SE = 93\% SP = 99\%
        %\vspace{0.6cm}
		\\
		\midrule
    Arco et al., 2014 &
		- &
		adaptive thresholding + hole filling, closing, regional minima &	
		gray, bin/disk &
		two (infected and noninfected) &
		Acc = 96.46\%
		%\vspace{0.6cm}
        \\
        \midrule

    Das et al., 2011 &
		- &
		marker controlled watershed + opening, closing &	
		gray, bin  &
		two (infected and noninfected) &
		Acc = 88.77\%
		%\vspace{0.6cm}
        \\
        \midrule

    Das et al., 2013 &
		- &
		marker controlled watershed &	
		gray &
	three (\emph{P.falciparum}, \emph{P.vivax} infected and noninfected) + three life-cycle-stages per species &
		Acc = 84\%
%		\vspace{0.6cm}
        \\
        \midrule
    Das et al., 2014 &
		- &
		marker controlled watershed &	
		gray &
	three (\emph{P.falciparum}, \emph{P.vivax} infected and noninfected) + three life-cycle-stages per species &
		SE = 99.72\% SP = 84.39\%
		%\vspace{0.6cm}
        \\
        \midrule
    Dave et al., 2017 &
		- &
		adaptive thresholding + erosion, dilation &	
		bin &
		two (infected and noninfected) &
		Acc = 97.83\% thin films, Acc = 89.88\% thick films
		%\vspace{0.6cm}
        \\
        \midrule
    Devi et al., 2017 &
		- &
		marker controlled watershed &	
		gray &
		two (infected and noninfected)  &
		Acc = 98.02\%
		%\vspace{0.6cm}
        \\
        \midrule
    Diaz et al., 2009 &
		- &
		inclusion tree &	
		gray &
		two (\emph{P.falciparum} infected and noninfected) + three life-cycle-stages &
		SE = 94\% SP = 99.7\% for detection, SE = 78.8\% SP = 91.2\% for life-stages
%\vspace{0.6cm}
		\\
		        \midrule	
		  Di Ruberto et al., 2002 &
		area closing, opening &
		thresholding + granulometry, watershed  transform + skeleton &	
		gray, bin/disk, flat and nonflat, with size depending on RBCs &
		two (\emph{P.falciparum} infected and noninfected) + three life-cycle-stages &
		-
%\vspace{0.6cm}
		\\
	\bottomrule
\end{tabular}  \end{table}
\begin{table}[H]\ContinuedFloat
\centering
\tablesize{\footnotesize} 
\caption{{\em Cont.}} \label{xx}
\begin{tabular}{m{3cm}<{\centering}m{3cm}<{\centering}m{5cm}<{\centering}m{3cm}<{\centering}m{5cm}<{\centering}m{2cm}<{\centering}}
		\toprule
    	\textbf{Authors} & \textbf{Preprocessing} & \textbf{Segmentation} & \textbf{MM/SE} & \textbf{Classification} & \textbf{Performance}\\  \midrule

    Elter et al., 2011 &
		- &
		thresholding + black top-hat, dilation &	
		gray/disk, nonflat with size = 9 &
		two (infected and noninfected) &
		SE = 97\%
%		\vspace{0.6cm}
\\
\midrule
    Gonzalez-Betancourt et al., 2016  &
    	morphological filter, erosion-reconstruction, dilation--reconstruction, closing &
    	watershed  transform  &
    	gray/disk with size depending on RBCs &
    	- &
    	-
    %\vspace{0.6cm}
    \\
    %&   &  &  &  & \\	
%    Ghosh \emph{et al.}, 2011  &
%    	- &
%    	thresholding + opening, closing  &
%    	bin &
%    	two (\emph{P.vivax} infected and noninfected) &
%    	-
%    %\vspace{0.6cm}
%    \\
\midrule
    Kareem et al., 2011, 2012  &
    	dilation, erosion &
    	-  &
    	gray/concentric ring, disk with size depending on RBCs  &
    	two (infected and noninfected) &
    	Acc = 88\% SE = 90\% SP = 86\%
    %\vspace{0.6cm}
    \\	
    \midrule
    Khan et al., 2011  &
    	area closing &
    	thresholding + granulometry, opening, morphological reconstruction, gradient, dilation &
    	gray &
    	five (\emph{P.falciparum}, \emph{P.vivax}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) &
    	Acc = 81\% SE = 85.5\%
    %\vspace{0.6cm}
    \\	
    \midrule
    Malihi et al., 2013  &
    	closing &
    	area granulometry  &
    	gray/disk with size depending on RBCs &
    	two (infected and noninfected) &
    	Acc = 91\% SE = 80\% SP = 95.5\%
    %\vspace{0.6cm}
    \\	
    \midrule
    Mushabe et al., 2013  &
    	closing &
    	thresholding + granulometry, dilation, erosion  &
    	gray, bin/disk &
    	two (infected and noninfected) &
    	SE = 98.5 SP = 97.2\%
    %\vspace{0.6cm}
    \\	
    \midrule
    Oliveira et al., 2017  &
    	erosion &
    	-  &
    	gray, bin &
    	two (infected and noninfected) &
    	Acc = 91\%
    %\vspace{0.6cm}
    \\
    \midrule
    Reni et al., 2015  &
    	new morphological filtering &
    	-  &
    	gray/anular ring, disk with size depending on RBCs &
    	- &
    	-
    %\vspace{0.6cm}
    \\
    \midrule
    Romero-Rondon et al., 2016  &
    	dilation, opening &
    	marker controlled watershed, erosion  &
    	gray, bin/disk with size depending on RBCs &
    	- &
    	-
    %\vspace{0.6cm}
    \\
        \midrule
      Rosado et al., 2017  &
    	- &
    	adaptive thresholding + closing  &
    	bin/elliptical with size = 3 &
    	four (\emph{P.falciparum}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) +
    three life-cycle-stages for species &
    	SE = 73.9--96.2\% SP = 92.6-99.3\%
    %\vspace{0.6cm}
    \\
    \midrule
    Ross et al., 2006  &
    	area closing &
    	thresholding + granulometry, opening, reconstruction, morphological gradient, closing, thinning  &
        gray, bin/disk with size = 6 and depending on RBCs, diamond with size = 1 &
    	five (\emph{P.falciparum}, \emph{P.vivax}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) &
    	SE = 85\% for detection, Acc = 73\% for classification
    	 %\vspace{0.6cm}
    \\
    	\bottomrule
\end{tabular}  \end{table}
\begin{table}[H]\ContinuedFloat
\centering
\tablesize{\footnotesize} 
\caption{{\em Cont.}} \label{xx}
\begin{tabular}{m{3cm}<{\centering}m{3cm}<{\centering}m{5cm}<{\centering}m{3cm}<{\centering}m{5cm}<{\centering}m{2cm}<{\centering}}
		\toprule
    	\textbf{Authors} & \textbf{Preprocessing} & \textbf{Segmentation} & \textbf{MM/SE} & \textbf{Classification} & \textbf{Performance}\\  \midrule

    Savkare et al., 2011a  &
    	- &
    	thresholding + hole filling, watershed transform  &
    	bin &
    	two (infected and noninfected) &
    	-
%    \vspace{0.6cm}
    \\
    \midrule
    Savkare et al., 2011b  &
    	- &
    	thresholding + hole filling, watershed transform  &
    	bin &
    	two (infected and noninfected) &
    	SE = 93.12\% SP = 93.17\%
    %\vspace{0.6cm}
    \\
    \midrule
    Savkare et al., 2015  &
    	- &
    	thresholding + watershed transform, erosion, dilation  &
    	bin/disk with size = 2 &
    	two (infected and noninfected) &
    	Acc = 95.5\%
    %\vspace{0.6cm}
    \\
    \midrule
    Sheikhhosseini et al., 2013  &
    	hole filling &
    	thresholding + hole filling, opening  &
    	bin &
    	two (infected and noninfected) &
    	Acc = 97.25\% SE = 82.21\% SP = 98.02\%
    %\vspace{0.6cm}
    \\
    \midrule
    Somasekar et al., 2015  &
    	erosion &
    	fuzzy C-means clustering + erosion, hole filling  &
    	bin/square with size = 3 &
    	two (infected and noninfected) &
    	SE = 98\% SP = 93.3\%
    %\vspace{0.6cm}
    \\
    \midrule
    Somasekar et al., 2017  &
    	- &
    	thresholding + erosion, closing, hole filling  &
    	bin/square with size = 3 &
    	two (infected and noninfected) &
    	average DSC = 0.8
    %\vspace{0.6cm}
    \\
    \midrule
    Soni et al., 2011  &
    	- &
    	thresholding + granulometry, morphological gradient, dilation  &
    	gray, bin/disk, flat and nonflat, vertical, horizontal &
    	five (\emph{P.falciparum}, \emph{P.vivax}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) &
    	SE = 98\% for detection
    %\vspace{0.6cm}
    \\
    \midrule
{\v{S}}pringl, 2009  &
    	closing &
    	thresholding + marker controlled watershed transform, hole filling, dilation, opening, erosion  &
    	gray, bin/disk with size depending on RBCs &
    	two (infected and noninfected) &
    	AUC = 0.98
    %\vspace{0.6cm}
    \\
    \midrule
    Sulistyawati et al., 2015  &
    	- &
    	blob analysis + erosion, dilation, opening, closing, hole filling  &
    	bin &
    	two (infected and noninfected) &
    	Acc = 99.39\%
    %\vspace{0.6cm}
    \\
    \midrule
 Tek et al., 2006  &
    	- &
    	top-hat, infinite reconstruction, area granulometry  &
    	gray, bin &
    	two (infected and noninfected) &
    	SE = 74\% SP = 98\%
    %\vspace{0.6cm}
    \\
    \midrule
    Tek et al., 2010  &
    	closing, granulometry &
    	thresholding + granulometry, area top-hat, closing, area granulometry &
    	gray/disk with size depending on RBCs &
    	five (\emph{P.falciparum}, \emph{P.vivax}, \emph{P.ovale}, \emph{P.malariae} infected, and noninfected) +
    four life-cycle-stages for species &
    	SE = 72\% SP = 98\%
    %\vspace{0.6cm}
    \\
    \midrule
    Yunda et al., 2012  &
    	- &
    	thresholding + morphological gradient, erosion, dilation  &
    	gray &
    	three (\emph{P.falciparum}, \emph{P.vivax} infected, and noninfected) + two life-cycle-stages for \emph{P.falciparum}&
    	SE = 77.19\% \\

		 \bottomrule
		\end{tabular}
	\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% change everything back
\newpage
\restoregeometry
\paperwidth=\pdfpageheight
\paperheight=\pdfpagewidth
\pdfpageheight=\paperheight
\pdfpagewidth=\paperwidth
\headwidth=\textwidth

Another often used segmentation approach is clearly the watershed transform. The classic watershed approach is reported to produce over segmentation results \cite{Savkare2011b}, whereas the marker controlled approach does not suffer from this issue, and it is reported to be very effective for overlapping cells' segmentation even though some authors affirm that it may fail to segment highly overlapped cells  \cite{Das2011,Das2013,Das2014,Das2015,Devi2017,Khan2011,Romero2016,Savkare2015,Savkare2011a,Springl2009}.

Other authors \cite{DiRuberto2002,Khan2011,Mushabe2013,Ross2006,Tek2010} employed granulometry and stated that it is very effective to segment cells with regular size.

The analysed works performed classification phase for different purposes. The majority of them aimed to distinguish among two classes only, malaria infected and noninfected RBCs, or to detect and count parasites in a malaria blood image \cite{Anggraini2011,Arco2014,Das2011,Dave2017,Das2015,DiRuberto2002,Elter2011,Kareem2012,Khan2014,Malihi2013,Mushabe2013,Oliveira2017,Savkare2011b,Savkare2011a,Somasekar2011,Somasekar2015,Somasekar2017,Sulist2015,Tek2006}.

More complex classification strategies aimed to classify parasites into different classes, \mbox{i.e.,  different human} parasites species \cite{Das2013,Das2014,Khan2011,Tek2010},
and/or different parasites life \mbox{stages \cite{Anggraini2011,Das2013,Das2014,DiRuberto2002,Diaz2009,Tek2010}}.

A summary of analysed methods is shown in Table \ref{t1}. For each approach, we report the processing phase where MM is used (preprocessing and/or segmentation), which operators are applied, which~type of morphology (grayscale or binary), and shape and size of the used SE, if~described. We~detail the kind of classification, i.e., if a method addresses the detection problem (infected and non infected) only, if it faces the labelling of different parasites (\emph{P. falciparum}, \emph{P. ovale}, \emph{P. malariae} or \emph{P. vivax}) and, in~some cases, the different stages of life (ring, trophozoite, schizont, gametocyte). Finally, the measures used for evaluating the system performance are reported, if present.

\section{Conclusions}\label{s5}
This work reviewed several computational microscopic imaging techniques oriented to a mathematical morphology approach, proposed in literature for malaria parasite detection and segmentation in blood smear microscopic images.

The computer vision methodologies reported in the literature are based on light microscopic images of human peripheral blood smears for computer-aided detection of malaria parasites and their different life stages.
Image preprocessing, segmentation of erythrocytes and parasites, malaria parasite feature extraction, and malaria detection techniques have been discussed here.

It is worth noticing that cell colour and the colour contrast between cells and background can vary so often according to the different, existing staining techniques, thickness of smear, microscope illumination and microscope's image acquisition procedure, as shown in Figure  \ref{fig:images_types}. A standardization of the procedure should be really useful to avoid superfluous differences in similar images' features and to have fair comparisons among the several proposed methods.  Moreover, independence of sample cannot be overlooked. Usually, any  pattern recognition system is evaluated by performing tests using a real set of samples. In the case of diagnosis, this is a key aspect. Any system for diagnosis should be tested on different images, acquired by different sources, and different specimens. This is crucial to guarantee the diagnostic capability and usefulness of a computer vision system. The main efforts towards the realization of a fully automatic blood cells segmentation and classification system cannot leave these aspects out.

Mathematical morphology techniques have been widely used for image processing purposes. Among the application fields, it has been applied for fingerprint feature extraction, recognition of handwritten digits, license plate detection, border extraction, denoising using morphological filters, text extraction and so on. Apart from these kinds of fields, mathematical morphology has been employed successfully in biomedical image analysis, especially in preprocessing and segmentation techniques.

Morphological cell analysis is used to face off abnormality identification and classification, early~cancer detection. It has been integrated in new methods for biomedical applications, such~as automatic segmentation and analysis of histological tumour sections, boundary detection of cervical cell nuclei considering overlapping and clustering, the granules segmentation and spatial distribution analysis, morphological characteristics analysis of specific biomedical cells, understanding the chemotactic response and drug influences, or identifying cell morphogenesis in different cell cycle progression.
Morphological feature quantification for grading cancerous or precancerous cells is especially widely researched in the literature, such as nuclei segmentation based on marker-controlled watershed transform and snake model for hepatocellular carcinoma feature extraction and classification, which is important for prognosis and treatment planning, nuclei feature quantification for cancer cell cycle analysis, and using feature extraction including image morphological analysis, wavelet analysis, and texture analysis for automated classification of renal cells \cite{Chen2012}.

Moreover, nonlinear filtering has become increasingly important in many image processing applications. Initially, the attraction to nonlinear filters was mostly limited to the impulse-removing and edge-preserving qualities of the median filter. However, as the number and sophistication of nonlinear filters have increased, so has the variety of applications for these filters. The shape-based methods of mathematical morphology, in particular, are now used in a wide variety of medical applications, including electrocardiography, ultrasound imaging, radiology, and histological image analysis \cite{Schulze1994}.

Furthermore, microscopic image analysis and, in particular, malaria detection and classification can greatly benefit from the use of mathematical morphology. The interest in this approach to image processing ad analysis is proved by the increasing number of works proposing methods for malaria image analysis based on mathematical morphology techniques.
Mainly, morphological techniques have been used in the preprocessing and segmentation phases, both in grayscale and in binary. The~most used operators have been dilation, erosion, opening and closing in order to improve the quality of the image in both of the phases, marker controlled watershed and granulometry to segment and detect RBCs and parasites.
No proposed system for malaria diagnosis is based on colour morphological operations. Since such techniques have been already applied for shape analysis, border detection through morphological gradient, segmentation, and texture analysis, a morphological approach using colour information could be extremely useful in blood smear processing and analysis.

In the end, it is worth considering that the development of new mobility-aware microscopic devices (and ideally low cost) is an area that can greatly improve the chances of the successful deployment of computer vision CAD solutions for malaria diagnosis in the field.
The mobile phone is currently Africa's most important digital technology, and is boosting African health as it emerges as a platform for diagnosis and treatment.
Considering the recent significant improvements of the new generation of mobile devices in terms of image acquisition and processing power, if~a reliable automatic diagnostic performance is ensured through the usage of those devices, one~would dramatically reduce the effort in the exhaustive and time-consuming activity of microscopic examination.
\mbox{Moreover, the lack} of highly trained microscopists on malaria diagnosis in rural areas could then be complemented by a significantly less specialized technician that knows how to operate the system and prepare blood smears. The usage of mobile devices in the system architecture can also bring significant improvements in terms of portability and data transmission, like the systems proposed by \cite{Oliveira2017,Rosado2017}.
Finally, malaria diagnosis might be just one element of a suite of diagnostic software tests running on this type of system. Several other tests could simultaneously be carried out using the same images, for instance cell counting or detection of other hemoparasites, like microfilaria or trypanosoma \cite{Rosado2016}.

%\input{Table_sum.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\supplementary{The following are available online at www.mdpi.com/link, Figure S1: title, Table S1: title, Video S1: title.}

% Only for the journal Methods and Protocols:
% If you wish to submit a video article, please do so with any other supplementary material.
% \supplementary{The following are available at www.mdpi.com/link: Figure S1: title, Table S1: title, Video S1: title. A supporting video article is available at doi: link.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{The authors would like to thank Centre Hospitalier Universitaire Vaudois (CHUV) of Lausanne for all the material provided and, in particular, Guy Prod'Hom for his availability and collaboration.}
%Pls define

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{A.L. and C.D.R. designed the study, structured the review and wrote the paper. M.K. reviewed the submitted manuscript and contributed to the writing of the paper. All authors read and approved the final version.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\conflictsofinterest{The authors declare no conflict of interest.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\abbreviations{The following abbreviations are used in this manuscript:\\

\noindent
\begin{tabular}{@{}ll}
CBC & Complete Blood Count \\
SJR & Scientific Journal Rank \\
JCR & Journal Citation Reports \\
WBC &White Blood Cell \\
RBC &Red Blood Cell \\
MM &Mathematical Morphology \\
MP &Malaria Parasite \\
SE &Structuring Element \\
CHUV & Centre Hospitalier Universitaire Vaudois
\end{tabular}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
%\appendixsections{multiple} %Leave argument "multiple" if there are multiple sections. Then a counter is printed ("Appendix A"). If there is only one appendix section then change the argument to "one" and no counter is printed ("Appendix").
%\appendix
%\section{}
%\subsection{}
%The appendix is an optional section that can contain details and data supplemental to the main text. For example, explanations of experimental details that would disrupt the flow of the main text, but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data is shown in the main text can be added here if brief, or as Supplementary data. Mathematical proofs of results not central to the paper can be added as an appendix.
%
%\section{}
%All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.

%=====================================
% References, variant A: internal bibliography
%=====================================
%\reftitle{References}
%\begin{thebibliography}{999}
%% Reference 1
%\bibitem[Author1(year)]{ref-journal}
%Author1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142-149, DOI.
%% Reference 2
%\bibitem[Author2(year)]{ref-book}
%Author2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32-58, ISBN.
%\end{thebibliography}

% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences. For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p. 328; \citeyear{ref-journal-3b}, p.475). This produces: Wong (1999, p. 328; 2000, p. 475)

%=====================================
% References, variant B: external bibliography
%=====================================
\reftitle{References}
\externalbibliography{yes}
\begin{thebibliography}{999}
\providecommand{\natexlab}[1]{#1}

\bibitem[Loddo \em{et~al.}(2016)Loddo, Putzu, Di~Ruberto, and Fenu]{Loddo2016}
Loddo, A.; Putzu, L.; Di~Ruberto, C.; Fenu, G.
\newblock A Computer-Aided System for Differential Count from Peripheral Blood
  Cell Images.
\newblock  In Proceedings of the 2016 12th International Conference on
  Signal-Image Technology Internet-Based Systems (SITIS),  Naples, Italy, 28 November--1 December 2016; pp. 112--118.

\bibitem[Di~Ruberto \em{et~al.}(2016)Di~Ruberto, Loddo, and
  Putzu]{DiRuberto2016}
Di~Ruberto, C.; Loddo, A.; Putzu, L.
\newblock A leucocytes count system from blood smear images: Segmentation~and
  counting of white blood cells based on learning by sampling.
\newblock {\em Mach. Vis.  Appl.} {\bf 2016}, {\em
  27},~1151--1160.

\bibitem[WHO(2016)]{WHO_dec_2016}
World Health Organization (WHO).
\newblock Malaria Fact Sheet December 2016. Available online: 
\newblock \url{http://www.who.int/mediacentre/factsheets/fs094/en/} 
\newblock (accessed on 6 March 2017).

\bibitem[Somasekar(2011)]{Somasekar2011}
Somasekar, J.
\newblock Computer vision for malaria parasite classification in erythrocytes.
\newblock {\em Int. J. Comput. Sci. Eng.} {\bf
  2011}, {\em 3},~2251--2256.

\bibitem[Soille(2004)]{Soille2004}
Soille, P.
\newblock {\em Morphological Image Analysis: Principles and Applications};
  Springer: Berlin, Germany,  2004; p. 392.

\bibitem[Giardina and Dougherty(1988)]{Giardina1988}
Giardina, C.; Dougherty, E.
\newblock {\em Morphological Methods in Image and Signal Processing};
  Prentice-Hall, Inc.: Upper~Saddle River, NJ, USA,  1988.

\bibitem[Serra(1984)]{Serra1984}
Serra, J.
\newblock {\em Image Analysis and Mathematical Morphology}; Academic Press: Cambridge, MA, USA,
  1984.

\bibitem[Benavent \em{et~al.}(2012)Benavent, Dura, Vegara, and
  Domingo]{Benavent2012}
Benavent, X.; Dura, E.; Vegara, F.; Domingo, J.
\newblock Mathematical Morphology for Color Images: An~Image-Dependent
  Approach.
\newblock {\em Math. Probl. Eng.} {\bf 2012}, {\em 2012},~1--19.

\bibitem[Nachtegael and Kerre(2000)]{Kerre2000}
Nachtegael, M.; Kerre, E. Classical and Fuzzy Approaches towards Mathematical
  Morphology.
\newblock In {\em Fuzzy Techniques in Image Processing}; Studies in Fuzziness and
  Soft Computing; Kerre, E., Nachtegael,
  M., Eds.; Physica-Verlag: Heidelberg,  Germany, 2000; Volume~52, Chapter~1, pp. 3--57.

\bibitem[Di~Ruberto \em{et~al.}(2002)Di~Ruberto, Dempster, Khan, and
  Jarra]{DiRuberto2002}
Di~Ruberto, C.; Dempster, A.; Khan, S.; Jarra, B.
\newblock Analysis of infected blood cell images using morphological operators.
\newblock {\em Image Vis. Comput.} {\bf 2002}, {\em 20},~133--146.

\bibitem[Rosado \em{et~al.}(2017)Rosado, da~Costa, Elias, and
  Cardoso]{Rosado2017}
Rosado, L.; da~Costa, J.M.C.; Elias, D.; Cardoso, J.S.
\newblock Mobile-Based Analysis of Malaria-Infected Thin Blood Smears:
  Automated Species and Life Cycle Stage Determination.
\newblock {\em Sensors} {\bf 2017}, {\em 17},~1--22.

\bibitem[Das \em{et~al.}(2015)Das, Mukherjee, and Chakraborty]{Das2015}
Das, D.; Mukherjee, R.; Chakraborty, C.
\newblock Computational microscopic imaging for malaria parasite detection: A
  systematic review.
\newblock {\em J. Microsc.} {\bf 2015}, {\em 260},~1--19.

\bibitem[Tek \em{et~al.}(2009)Tek, Dempster, and Kale]{Tek2009}
Tek, F.; Dempster, A.; Kale, I.
\newblock Computer vision for microscopy diagnosis of malaria.
\newblock {\em Malar. J.} {\bf 2009}, {\em 8},~1--14.

\bibitem[Oliveira \em{et~al.}(2017)Oliveira, Prats, Espasa, and
  et~al.]{Oliveira2017}
Oliveira, A.; Prats, C.; Espasa, M.; Serrat, F.Z.; Sales, C.M.; Silgado, A.; Codina, D.L.; Arruda, M.E.; i Prat, J.G.; Albuquerque, J.
\newblock The Malaria System MicroApp: A New, Mobile Device-Based Tool for
  Malaria Diagnosis.
\newblock {\em JMIR Res. Protoc.} {\bf 2017}, {\em 6},~1--11.

\bibitem[Dave and Upla(2017)]{Dave2017}
Dave, I.; Upla, K.
\newblock Computer Aided Diagnosis of Malaria Disease for Thin and Thick Blood
  Smear Microscopic Images.
\newblock  In Proceedings of the 2017 4th International Conference on Signal
  Processing and Integrated Networks (SPIN), Noida, India, 2--3 February 2017; pp. 561--561.

\bibitem[Devi \em{et~al.}(2017)Devi, Singha, Sharma, and Laskar]{Devi2017}
Devi, S.; Singha, J.; Sharma, M.; Laskar, R.
\newblock Erythrocyte segmentation for quantification in microscopic images of
  thin blood smears.
\newblock {\em J. Intell. Fuzzy Syst.} {\bf 2017}, {\em
  32},~2847--2856.

\bibitem[Somasekar and Reddy(2017)]{Somasekar2017}
Somasekar, J.; Reddy, B.
\newblock A Novel Two-Stage Thresholding Method for Segmentation of Malaria
  Parasites in Microscopic Blood Images.
\newblock {\em J. Biomed. Eng. Med. Imaging} {\bf
  2017}, {\em 4},~31--42.

\bibitem[Sulistyawati \em{et~al.}(2015)Sulistyawati, Rahmanti, Purnama, and
  Purnomo]{Sulist2015}
Sulistyawati, D.; Rahmanti, F.; Purnama, I.; Purnomo, M.
\newblock Automatic Segmentation of Malaria Parasites on Thick Blood Film using
  Blob Analysis.
\newblock  In Proceedings of the 2015 International Seminar on Intelligent
  Technology and Its Applications, Surabaya, Indonesia,  20--21 May 2015; pp. 137--142.

\bibitem[Khan \em{et~al.}(2011)Khan, Acharya, Singh, and Soni]{Khan2011}
Khan, M.; Acharya, B.; Singh, B.; Soni, J.
\newblock Content based image retrieval approaches for detection of malarial
  parasite in blood images.
\newblock {\em Int. J. Biometr. Bioinform. (IJBB)}
  {\bf 2011}, {\em 5},~97--110.

\bibitem[Savkare and Narote(2011{\natexlab{a}})]{Savkare2011a}
Savkare, S.; Narote, S.
\newblock Automatic Classification of Normal and Infected Blood Cells for
  Parasitemia Detection.
\newblock {\em Int. J. Comput. Sci. Netw. Secur.
  (IJCSS)} {\bf 2011}, {\em 11},~94--97.

\bibitem[Savkare and Narote(2011{\natexlab{b}})]{Savkare2011b}
Savkare, S.; Narote, S.
\newblock Automatic detection of malaria parasites for estimating parasitemia.
\newblock {\em Int. J. Comput. Sci. Netw. Secur.
  (IJCSS)}
  {\bf 2011}, {\em 5},~310--315.

\bibitem[Das \em{et~al.}(2014)Das, Maiti, and Chakraborty]{Das2014}
Das, D.; Maiti, A.; Chakraborty, C. Textural pattern classification of
  microscopic images for malaria screening.
\newblock In {\em Advances in Therapeutic Engineering}; Yu, W., Chattopadhyay,
  S., Lim, T., Acharya, U., Eds.; CRC Press: Boca~Raton, FL, USA, 2014; Chapter~19, pp. 419--446.

\bibitem[Schulze(1994)]{Schulze1994}
Schulze, M.
\newblock Biomedical Image Processing with Morphology-Based Nonlinear Filters.~Ph.D. Thesis, University~of Texas at Austin, Austin, TX, USA, 1994.

\bibitem[{\v{S}}pringl(2009)]{Springl2009}
{\v{S}}pringl, V.
\newblock Automatic Malaria Diagnosis Through Microscopy Imaging.
\newblock Master's Thesis, Czech Technical University in Prague, Faculty of
  Elecrical Engineering, Prague, Czech Republic,  2009.

\bibitem[Khan \em{et~al.}(2014)Khan, Pervaz, Latif, Musharraf, and
  Saniya]{Khan2014}
Khan, N.; Pervaz, H.; Latif, A.; Musharraf, A.
\newblock Unsupervised identification of malaria parasites using computer
  vision.
\newblock  In Proceedings of the 2014 11th International Joint Conference on
  Computer Science and Software Engineering (JCSSE), Chonburi, Thailand, 14--16 May 2014; pp. 263--267.

\bibitem[Sathpathi \em{et~al.}(2014)Sathpathi, Mohanty, Satpathi, Mishra,
  Behera, Patel, and Dondorp]{Sathpathi2014}
Sathpathi, S.; Mohanty, A.; Satpathi, P.; Mishra, S.; Behera, P.; Patel, G.;
  Dondorp, A.
\newblock Comparing Leishman and Giemsa staining for the assessment of
  peripheral blood smear preparations in a malaria-endemic region in India.
\newblock {\em Malar. J.} {\bf 2014}, {\em 13},~512--516.

\bibitem[Gonzalez-Betancourt and et~alt.(2016)]{Gonzalez2016}
\textls[-15]{González-Betancourt, A.; Rodríguez-Ribalta, P.; Meneses-Marcel, A.; Sifontes-Rodríguez, S.; Lorenzo-Ginori,~J.V.;} Orozco-Morales, R. 
\newblock Automated marker identification using the Radon transform for
  watershed segmentation.
\newblock {\em IET Image Process.} {\bf 2016}, {\em 11},~183--189.

\bibitem[Kareem \em{et~al.}(2012)Kareem, Kale, and Morling]{Kareem2012}
Kareem, S.; Kale, I.; Morling, R.
\newblock Automated Malaria Parasite Detection in Thin Blood Films:- A Hybrid
  Illumination and Color Constancy Insensitive, Morphological Approach.
\newblock  In Proceedings of the 2012 IEEE Asia Pacific Conference on Circuits
  and Systems (APCCAS), Kaohsiung, Taiwan, 2--5 December 2012; pp.~240--243.

\bibitem[Romero-Rondon \em{et~al.}(2016)Romero-Rondon, Sanabria-Rosas,
  Bautista-Rozo, and Mendoza-Castellanos]{Romero2016}
Romero-Rondon, M.; Sanabria-Rosas, L.; Bautista-Rozo, L.; Mendoza-Castellanos,
  A.
\newblock Algorithm for detection of overlapped red blood cells in microscopic
  images of blood smears.
\newblock {\em DYNA} {\bf 2016}, {\em 83},~187--194.

\bibitem[Reni \em{et~al.}(2015)Reni, Kale, and Morling]{Reni2015}
Reni, S.K.; Kale, I.; Morling, R. Analysis of thin blood images for automated malaria diagnosis.  In~Proceedings of the 2015 E-Health and Bioengineering Conference
  (EHB),  Iasi, Romania, 19--21 November 2015; pp. 1--4.

\bibitem[Sheikhhosseini \em{et~al.}(2013)Sheikhhosseini, Rabbani, Zekri, and
  Talebi]{Sheik2013}
Sheikhhosseini, M.; Rabbani, H.; Zekri, M.; Talebi, A.
\newblock Automatic diagnosis of malaria based on complete circle ellipse
  fitting search algorithm.
\newblock {\em J. Microsc.} {\bf 2013}, {\em 252},~189--203.

\bibitem[Somasekar and Reddy(2015)]{Somasekar2015}
Somasekar, J.; Reddy, B.
\newblock Segmentation of erythrocytes infected with malaria parasites for the
  diagnosis using microscopy imaging.
\newblock {\em Comput. Electr. Eng.} {\bf 2015}, {\em
  45},~336--351.

\bibitem[Tek \em{et~al.}(2010)Tek, Dempster, and Kale]{Tek2010}
Tek, F.; Dempster, A.; Kale, I.
\newblock Parasite detection and identification for automated thin blood film
  malaria diagnosis.
\newblock {\em Comput. Vis. Image Underst.} {\bf 2010}, {\em
  114},~21--32.

\bibitem[Arco \em{et~al.}(2014)Arco, Gorriz, Ramirez, Alvarez, and
  Puntonet]{Arco2014}
Arco, J.; Gorriz, J.; Ramirez, J.; Alvarez, I.; Puntonet, C.
\newblock Digital image analysis for automatic enumeration of malaria parasites
  using morphological operations.
\newblock {\em Expert Syst. Appl.} {\bf 2014}, {\em 42},~3041--3047.

\bibitem[Anggraini \em{et~al.}(2011)Anggraini, Nugroho, Pratama, Rozi,
  Iskandar, and Hartono]{Anggraini2011}
Anggraini, D.; Nugroho, A.; Pratama, C.; Rozi, I.; Iskandar, A.; Hartono, R.
\newblock Automated status identification of microscopic images obtained from
  malaria thin blood smears.
\newblock  In Proceedings of the 2011 International Conference on Electrical
  Engineering and Informatics,  Bandung, Indonesia, 17--19 July 2011.

\bibitem[Elter \em{et~al.}(2011)Elter, Ha{\ss}lmeyer, and
  Zerfa{\ss}]{Elter2011}
Elter, M.; Ha{\ss}lmeyer, E.; Zerfa{\ss}, T.
\newblock Detection of malaria parasites in thick blood films.
\newblock  In Proceedings of the 2011 Annual International Conference of the
  IEEE Engineering in Medicine and Biology Society, Boston, MA,~USA, 30 August--3 September 2011; pp. 5140--5144.

\bibitem[Kareem \em{et~al.}(2011)Kareem, Morling, and Kale]{Kareem2011}
Kareem, S.; Morling, R.; Kale, I.
\newblock A novel method to count the red blood cells in thin blood films.
\newblock  In Proceedings of the 2011 IEEE International Symposium on Circuits
  and Systems (ISCAS), Rio de Janeiro, Brazil, 15--18 May 2011; pp. 1021--1024.

\bibitem[Mushabe \em{et~al.}(2013)Mushabe, Dendere, and Douglas]{Mushabe2013}
Mushabe, M.; Dendere, R.; Douglas, T.
\newblock Automated detection of malaria in Giemsa-stained thin blood smears.
\newblock  In Proceedings of the 2013 35th Annual International Conference of
  the IEEE Engineering in Medicine and Biology Society (EMBC), Osaka, Japan, 3--7 July 2013; pp. 3698--3701.

\bibitem[Ross \em{et~al.}(2006)Ross, Pritchard, Rubin, and Dus{\'e}]{Ross2006}
Ross, N.; Pritchard, C.; Rubin, D.; Dus{\'e}, A.
\newblock Automated image processing method for the diagnosis and
  classification of malaria on thin blood smears.
\newblock {\em Med. Biol. Eng. Comput.} {\bf 2006},
  {\em 44},~427--436.

\bibitem[Otsu(1975)]{Otsu1975}
Otsu, N.
\newblock A threshold selection method from gray-level histograms.
\newblock {\em Automatica} {\bf 1975}, {\em 11},~23--27.

\bibitem[Ahirwar \em{et~al.}(2012)Ahirwar, Pattnaik, and Acharya]{Ahirwar2012}
Ahirwar, N.; Pattnaik, S.; Acharya, B.
\newblock Advanced image analysis based system for automatic detection and
  classification of malarial parasite in blood images.
\newblock {\em Int. J. Inf. Technol. Knowl.
  Manag.} {\bf 2012}, {\em 5},~59--64.

\bibitem[Soni \em{et~al.}(2011)Soni, Mishra, and Kamargaonkar]{Soni2011}
Soni, J.; Mishra, N.; Kamargaonkar, C.
\newblock Automatic difference between RBC and Malaria parasites based on
  morphology with first order features using image processing.
\newblock {\em Int. J. Adv. Eng. Technol.}
  {\bf 2011}, {\em 1},~290--297.

\bibitem[Yunda \em{et~al.}(2012)Yunda, Alarc{\'o}n, and Mill{\'a}n]{Yunda2012}
Yunda, L.; Alarc{\'o}n, A.; Mill{\'a}n, J.
\newblock Automated image analysis method for p-vivax malaria parasite
  detection in thick film blood images.
\newblock {\em Sistemas \& Telem{\'a}tica} {\bf 2012}, {\em 10},~9--25.

\bibitem[Das \em{et~al.}(2011)Das, Ghosh, Chakraborty, Maiti, and Pal]{Das2011}
Das, D.; Ghosh, M.; Chakraborty, C.; Maiti, A.; Pal, M.
\newblock Probabilistic prediction of malaria using morphological and textural
  information.
\newblock  In Proceedings of the 2011 International Conference on Image
  Information Processing (ICIIP), Shimla, India, 3--5 November 2011; pp. 1--6.

\bibitem[Das \em{et~al.}(2013)Das, Ghosh, Pal, Maiti, and Chakraborty]{Das2013}
Das, D.; Ghosh, M.; Pal, M.; Maiti, A.; Chakraborty, C.
\newblock Machine learning approach for automated screening of malaria parasite
  using light microscopic images.
\newblock {\em Micron} {\bf 2013}, {\em 45},~97--106.

\bibitem[Dey \em{et~al.}(2015)Dey, Roy, Bhattacharjee, Nasipuri, and
  Ghosh]{Dey2015}
Dey, R.; Roy, K.; Bhattacharjee, D.; Nasipuri, M.; Ghosh, P.
\newblock An Automated system for Segmenting platelets from Microscopic images
  of Blood Cells.
\newblock  In Proceedings of the 2015 International Symposium on Advanced
  Computing and Communication (ISACC), Silchar, India, 14--15 September 2015; pp. 230--237.

\bibitem[Diaz \em{et~al.}(2009)Diaz, Gonzalez, and Romero]{Diaz2009}
Diaz, G.; Gonzalez, F.; Romero, E.
\newblock A semi-automatic method for quantification and classification of
  erythrocytes infected with malaria parasites in microscopic images.
\newblock {\em J. Biomed. Inform.} {\bf 2009}, {\em
  42},~296--307.

\bibitem[Savkare and Narote(2015)]{Savkare2015}
Savkare, S.; Narote, S.
\newblock Blood Cell Segmentation from Microscopic Blood Images.
\newblock  In Proceedings of the 2015 International Conference on Information
  Processing (ICIP), Shimla, India, 21--24 December 2015; pp.~502--505.

\bibitem[Tek \em{et~al.}(2006)Tek, Dempster, and Kale]{Tek2006}
Tek, F.; Dempster, A.; Kale, I.
\newblock Malaria parasite detection in peripheral blood images.
\newblock  In Proceedings of the 2006 British Machine Vision Conference (BMVC), Edinburgh, UK, 4--7 September  2006;  pp. 347--356.

\bibitem[Jiang(2009)]{Jiang2009}
Jiang, X.
\newblock Feature extraction for image recognition and computer vision.
\newblock  In Proceedings of the 2009 2nd IEEE International Conference on
  Computer Science and Information Technology (ICCSIT), Beijing, China, \mbox{8--11 August 2009}; pp. 1--15.

\bibitem[Malihi \em{et~al.}(2013)Malihi, Ansari-Asl, and Behbahani]{Malihi2013}
Malihi, L.; Ansari-Asl, K.; Behbahani, A.
\newblock Malaria parasite detection in giemsa-stained blood cell images.
\newblock  In~Proceedings of the 2013 8th Iranian Conference on Machine Vision
  and Image Processing (MVIP), Zanjan, Iran, 10--12 September 2013; pp. 360--365.

\bibitem[Chen \em{et~al.}(2012)Chen, Zhao, Wu, Yao, and Zhang]{Chen2012}
Chen, S.; Zhao, M.; Wu, G.; Yao, C.; Zhang, J.
\newblock Recent Advances in Morphological Cell Image Analysis.
\newblock {\em Comput. Math. Methods  Med.} {\bf 2012},
  {\em 2012},~1--10.

\bibitem[Rosado \em{et~al.}(2016)Rosado, da~Costa, Elias, and
  Cardoso]{Rosado2016}
Rosado, L.; da~Costa, J.M.C.; Elias, D.; Cardoso, J.S.
\newblock A review of automatic malaria parasites detection and segmentation in
  microscopic images.
\newblock {\em Anti-Infect. Agents} {\bf 2016}, {\em 14},~11--22.

\end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\sampleavailability{Samples of the compounds ...... are available from the authors.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}
